### 2025-05-15 manuscript.tex
ediittt Prof comment

\documentclass[lettersize,journal]{IEEEtran}

\def\pub{false} % true for publication, false for draft
\newcommand*{\template}{template} 
\input{\template/preamble/preamble_journal.tex}

\newcommand*{\figSizeTwoCol}{.49}
\newcommand*{\figSizeOneCol}{0.98}

\addbibresource{\template/refs.bib}

\begin{document}

\title{
    Constrained Optimization-Based Neuro-Adaptive Control (CONAC) for Euler-Lagrange Systems Under Weight and Input Constraints
} %% Article title

\author{
  Myeongseok Ryu, Donghwa Hong, and Kyunghwan Choi
\thanks{
    Myeongseok Ryu and Kyunghwan Choi are with the Cho Chun Shik Graduate School of Mobility, Korean Advanced Institute of Science and Technology (KAIST), Daejeon 34051, Korea (e-mail: dding\_98@kaist.ac.kr and kh.choi@kaist.ac.kr).

    Donghwa Hong is with the Department of Mechanical and Robotics Engineering, Gwangju Institute of Science and Technology (GIST), Gwangju 61005, Korea (e-mail: fairytale@gm.gist.ac.kr).

    This work was supported in part by a Korea Research Institute for Defense Technology planning and advancement (KRIT) grant funded by Korea government DAPA (Defense Acquisition Program Administration) (No. KRIT-CT-22-087, Synthetic Battlefield Environment based Integrated Combat Training Platform) and in part by the National Research Foundation of Korea (NRF) grant funded by the Korea government (MSIT) (RS-2025- 00554087). (Corresponding author: Kyunghwan Choi)
    }}

\markboth{Journal of \LaTeX\ Class Files,~Vol.~18, No.~9, September~2020}%
{How to Use the IEEEtran \LaTeX \ Templates}

\maketitle

%% Abstract
\begin{abstract}
  This study presents a constrained optimization-based neuro-adaptive controller (CONAC) for uncertain Euler-Lagrange systems subject to weight norm and convex input constraints. 
  A deep neural network (DNN) is employed to approximate the ideal stabilizing control law while simultaneously learning the unknown system dynamics and addressing both types of constraints within a constrained optimization framework.
  The weight adaptation laws are formulated to solve the defined constrained optimization problem, ensuring satisfaction of first-order optimality conditions at steady state.
  The controller's stability is rigorously analyzed using Lyapunov theory, guaranteeing bounded tracking errors and bounded DNN weights. 
  The proposed controller is validated through a real-time implementation on a 2-DOF robotic manipulator, demonstrating its effectiveness in achieving accurate trajectory tracking while satisfying all imposed constraints.

  via a real-time implementation conducted on a 2-DOF robotic manipulator, demonstrating the controller's effectiveness in achieving desired trajectory tracking while satisfying constraints.
\end{abstract}

\begin{IEEEkeywords}
    Neuro-adaptive control, constrained optimization, deep neural network, input constraint.
\end{IEEEkeywords}

\section*{Notation}
In this study, the following notation is used:

\begin{itemize}
    \item $\otimes$ denotes the Kronecker product \cite[Chap. 7 Def. 7.1.2]{Bernstein:2009aa}.
    \item A vector and a matrix are denoted by $\mv{x}=[x_i]_{i\in\{1,\cdots,n\}}\in\R^n$ and $
        \mm A
        := 
        [a_{ij}]
        _{
            i\in\{1,\cdots,n\},j\in\{1,\cdots ,m\}
        }\allowbreak\in\R^{n\times m}
        $, respectively.
    \item $\myrow_i(\mm A)$ denotes the $i\textsuperscript{th}$ row of the matrix $\mm A\in\R^{n\times m}$. 
    \item For $\mm{A}\in\R^{n\times m}$, denotes the vectorization of $\mm A$, $\myvec(\mm A):=(\myrow_1(\mm A^\top),\cdots,\myrow_m(\mm A^\top)  )^\top$.
    % \item $\mycol_i(A)$ and $\myrow_j(A)$ denote the $i\textsuperscript{th}$ column and $j\textsuperscript{th}$ row of $A\in\R^{n\times m}$, respectively.
    % \item $\myvec(A):= [\mycol_1(A)^\top  ,\cdots,\mycol_m(A)^\top  ]^\top   $ for $A\in\R^{n\times m}$.
    \item $\lambda_{\min}(\mm A)$ denotes the minimum eigenvalue of the matrix $\mm A\in\R^{n\times n}$.
    \item $\mm I_n$ denotes the $n\times n$ identity matrix, and $\mm 0_{n\times m}$ denotes the $n\times m$ zero matrix.
\end{itemize}

%  SECTION INTRODUCTION ===================================
\section{Introduction}

\subsection{Background}

\IEEEPARstart{M}{any} engineering systems, including those in aerospace, robotics, and automotive applications, can be modeled using Euler-Lagrange systems. 
These systems are governed by dynamic equations derived from energy principles and describe the motion of mechanical systems with constraints. 
In practice, however, such systems often exhibit uncertainties due to unmodeled dynamics, parameter variations, or external disturbances. 
These uncertainties can significantly degrade control performance and, in some cases, lead to instability. 
To address these challenges, adaptive control methods have been widely employed to ensure robust performance in the presence of system uncertainties \cite{Ioannou:2006aa, Tao:2003aa}.

More recently, neuro-adaptive control approaches have been introduced to approximate unknown system dynamics or entire control laws using neural networks (NNs) \cite{Farrell:2006aa}. 
NNs are well-known for their universal approximation property, which allows them to approximate any smooth function over a compact set with minimal error. 
Various types of NNs have been utilized in neuro-adaptive control, including simpler architectures like single-hidden layer (SHL) NNs \cite{Ge:2010aa, Yesildirek:1995aa} and radial basis function (RBF) neural networks \cite{Liu:2013ab,Ge:2002aa}, as well as more complex models like deep neural networks (DNNs) \cite{Patil:2022aa} and their variations. 
SHL and RBF NNs are often employed to approximate uncertain system dynamics or controllers due to their simplicity \cite{Esfandiari:2014aa,Esfandiari:2015aa,Yesildirek:1995aa,Gao:2006aa}, while DNNs offer greater expressive power, making them more effective for complex system approximations \cite{Rolnick:2018aa}. 
Additionally, variations of DNNs, such as long short-term memory (LSTM) networks for time-varying dynamics \cite{Liu:2013ab} and physics-informed neural networks (PINNs) for leveraging physical system knowledge \cite{Hart:2024aa}, have further extended the capabilities of neuro-adaptive control systems.

A critical aspect of neuro-adaptive control is the weight adaptation law, which governs how NN parameters are updated. 
Most studies derived these laws using Lyapunov-based methods, ensuring the boundedness of the tracking error and weight estimation error, thus maintaining system stability under uncertainty.

However, two significant challenges persist in using NNs for adaptive control. 
First, the boundedness of NN weights is not inherently guaranteed, which can result in unbounded outputs. 
When NN outputs are used directly in the control law, this may lead to excessive control inputs, violating input constraints. 
Such constraints are commonly encountered in industrial systems, where actuators are limited by physical and safety requirements in terms of amplitude, rate, or energy \cite{Esfandiari:2021aa}. 
Failing to address these constraints can degrade control performance or even destabilize the system.

Therefore, addressing these two key issues—ensuring weight boundedness and satisfying input constraints—is essential for the reliable design of neuro-adaptive controllers. 
The following section provides a detailed review of existing solutions to these challenges.

\subsection{Literature Review}

\subsubsection{Ensuring Weight Norm Boundedness}

A common challenge in neuro-adaptive control is maintaining the boundedness of the NN weights to ensure stability. 
In many studies, projection operators were employed to enforce upper bounds on the weight norms, ensuring that the weights do not grow unboundedly. 
For example, in \cite{Zhou:2023aa,Griffis:2023aa,Patil:2022aa}, projection operators were used to constrain the weight norms to remain below predetermined constants. 
However, these constants were often selected as large as possible due to the lack of theoretical guarantees regarding the global optimality of the weight values. 
While this approach ensured that the NN remained stable, it did not necessarily result in optimal performance.

In addition to projection operators, some studies utilized modification techniques like $\sigma$-modification \cite{Ge:2002aa} and $\epsilon$-modification \cite{Esfandiari:2015aa,Gao:2006aa}. 
These methods ensured that the NN weights remained within an invariant set by incorporating stabilizing functions into the adaptation law. 
Although these techniques were effective in ensuring boundedness and avoiding weight divergence, they similarly lacked a formal analysis of the optimality of the adapted weights, leaving room for improvement in terms of performance optimization.

\subsubsection{Satisfying Input Constraints}

The second major issue is satisfying input constraints, particularly in systems where actuators are subject to physical limitations. 
The unpredictable outputs of NNs can sometimes lead to excessively large control inputs, violating these constraints. 
This problem is exacerbated in neuro-adaptive controllers that attempt to cancel out system dynamics using conventional methods like feedback linearization or backstepping. 
In such cases, controllers may produce overly aggressive control inputs, even when the system's natural dynamics are stabilizing, leading to unnecessary saturation of the control inputs.

To address input saturation, many studies introduced auxiliary systems. 
These systems mitigated the effects of control input saturation by modifying the control strategy when saturation occurred. 
For instance, in \cite{Esfandiari:2014aa,Karason:1994aa,Esfandiari:2015aa}, auxiliary states were generated whenever input saturation was detected, and these states were incorporated into the adaptation law to adjust the NN weights accordingly. 
This approach helped the controller reduce input saturation by indirectly regulating the auxiliary states.

Alternatively, auxiliary states can also be used as feedback terms in the control law to directly compensate for the effects of input saturation constraints, as demonstrated in \cite{Arefinia:2020aa,He:2016aa,Peng:2020aa}. 
In \cite{Gao:2006aa}, the NN was used to approximate the desired control input, compensating for input saturation. 
However, these approaches typically handle input bound constraints on a per-input basis (\ie applying bounds to each scalar control variable individually), and may not account for more complex, nonlinear constraints, like input norm constraints, which are commonly found in physical systems such as robotic actuators or motor systems.

\subsubsection{Limitations of Existing Approaches and Potential of Constrained Optimization}

Although both the projection operator methods for weight norm boundedness and the auxiliary system approach for input constraints have shown effectiveness, they come with significant limitations. 
Projection operators and modification techniques do not guarantee the optimality of the adapted weights. 
Moreover, auxiliary systems typically handle only simple forms of input constraints, such as input bound constraints, limiting their ability to address more complex, nonlinear constraints like input norm constraints.

To overcome these limitations, constrained optimization offers a promising approach. 
By formulating the neuro-adaptive control problem as an optimization problem with constraints, it is possible to adapt the NN weights while minimizing an objective function (\eg tracking error) subject to both weight norm and input constraints. 
Constrained optimization provides a theoretical framework for defining optimality and presents numerical methods for finding solutions that satisfy the constraints \cite{Nocedal:2006aa}.

In existing literature, constrained optimization techniques, such as the Augmented Lagrangian Method (ALM) \cite{Evens:2021aa} and the Alternating Direction Method of Multipliers (ADMM) \cite{Wang:2019aa,Taylor:2016aa}, have been used to train NNs offline. 
These methods impose constraints to address issues like gradient vanishing in backpropagation. 
However, to the best of the authors' knowledge, no prior work has applied constrained optimization to adaptive control systems with real-time weight adaptation. 
This gap suggests that constrained optimization could be key to addressing both weight norm boundedness and input constraints in a unified, theoretically grounded framework, particularly in real-time neuro-adaptive control.

\subsection{Contributions}

\KHCH{Abstract에 있는 것처럼 이 내용도 하나 추가해두죠.

The controller’s stability is rigorously
analyzed using Lyapunov theory, guaranteeing bounded
tracking errors and DNN weights.}

The main contributions of this study are listed as follows:
\begin{itemize}
    \item A constrained optimization-based neuro-adaptive controller (CONAC) is proposed, where the design of a deep neural network (DNN) is formulated as a constrained optimization problem. In this formulation, weight norm and convex input constraints are incorporated as inequality constraints, while minimizing a given objective function.
    % \item A constrained optimization-based neuro-adaptive controller (CONAC) is developed using a DNN, where input constraints and the boundedness of DNN weights are formulated as inequality constraints within the adaptation process.
    \item The adaptation laws for DNN weights in CONAC are systematically derived to solve the defined problem, using constrained optimization theory. These laws guarantee convergence to the first-order optimality conditions, specifically the Karush–Kuhn–Tucker (KKT) conditions, as described in \cite[Chap. 12 Thm. 12.1]{Nocedal:2006aa}.
    % \item The weight adaptation laws in CONAC are derived using constrained optimization theory to minimize the objective function while satisfying the inequality constraints. The adaptation laws ensure convergence of the weights to the first-order optimality conditions, specifically the Karush-Kuhn-Tucker (KKT) conditions defined in \cite[Chap. 12 Thm. 12.1]{Nocedal:2006aa}. 
    % \item The forward sensitivity method is applied in CONAC to accurately calculate the gradient of the objective function. \MSRY{WILL BE DELETED}
    %, enabling more precise real-time adaptation.
\end{itemize}

\subsection{Organization}

\MSRY{Started from here}

The remainder of this paper is organized as follows. 
Section \ref{sec:Problem Formulation} presents the target system and control objective.
Section \ref{sec:ctrl design} introduces the proposed controller and the architecture of DNN in the controller. 
In Section \ref{sec:adap_laws}, the weight adaptation laws are developed.
Section \ref{sec:stability} analyzes the stability of the proposed controller.
A real-time implementation of the proposed controller is presented in Section \ref{sec:sim}, where the proposed controller is applied to a 2-DOF robotic manipulator.
Finally, Section \ref{sec:conclusion} concludes the paper and discusses potential future work.
The candidates of input constraints are presented in Appendix \ref{sec:appen:cstr}.

% ====================================
%  SECTION PROBLEM FORMULATION 
% ====================================
\section{Problem Formulation}\label{sec:Problem Formulation}

\subsection{Model Dynamics and Control Objective}

We consider an uncertain Euler-Lagrange system modeled as
\begin{equation}
    \rbM\ddq + \rbVm + \rbF + \rbG + \rbd
    =
    \mysat(\rbu)
    \label{eq:sys1}
\end{equation}
where $\q\in \R^n$ denotes the generalized coordinate, $\rbd$ represents the disturbance and $\rbu\in\R^n$ denotes the generalized control input. 
The terms $\rbM:=\rbM(\q)\in\R^{n\times n}$, $\rbVm:=\rbVm(\q,\dq)\in\R^{n\times n}$, $\rbF:=\rbF(\dq)\in\R^{n}$, and $\rbG:=\rbG(\q)$ represent the unknown inertia matrix, Coriolis/centripetal matrix, friction vector, and gravity vector, respectively.
The function $\mysat(\cdot):\R^n\to\R^n$ represents the inherent physical limitations of the actuators such that $\norm{\mysat(\rbu)}\le \overline\tau$, where $\overline\tau$ denotes the maximum norm of control input.
To account for these limitations, it is essential to incorporate physically motivated constraints into the controller design.
Appendix \ref{sec:appen:cstr} introduces candidate constraints that can be applied to ensure compliance with these physical limitations.

The Euler-Lagrange system \eqref{eq:sys1} satisfies several important physical properties, as presented in \cite[Chap. 3 Tab. 3.2.1]{Lewis:1998aa}. 
The key properties are introduced below:
\begin{prop} 
    The inertia matrix $\rbM$ is symmetric, positive definite, and bounded.
    \label{prop:M}
\end{prop}

\begin{prop} 
    The Coriolis/centripetal\MSRY{인용 책 그대로 적은 것입니다.} matrix $\rbVm$ can always be selected\MSRY{상동.} so that the matrix $\ddtt{\rbM}-2\rbVm$ is skew-symmetric, \ie $\mv{x}^\top(\ddtt{\rbM}-2\rbVm)\mv{x}=0,\ \forall \mv{x}\in\R^n$.
    \label{prop:skew}
\end{prop}

\begin{prop}
    The disturbance $\rbd$ is bounded, \ie $\norm{\rbd}\le \overline\tau_d$.
    \label{prop:dis_bound}
\end{prop}

The control objective is to develop a neuro-adaptive controller that enables $\q$ to track a continuously differentiable desired trajectory $\qd:=\qd(t): \R_{\ge0} \to \R^n$, compensating for the unknown system dynamics while addressing the imposed constraints.
The desired trajectory $\qd(t)$ is supposed to satisfy the following assumption:
\begin{assum}
    The desired trajectory $\qd(t)$ is assumed to be bounded, i.e., $\norm{\qd(t)}\le \overline{q}_d$ , and available for designing a bounded control input.
    \label{assum:feasible}
\end{assum}

%  SECTION CONTROLLER DESIGN ==============================
\section{Control Law Development}\label{sec:ctrl design}

The architecture of the proposed CONAC is illustrated in Fig.~\ref{fig:ctrl:diagram}, and consists of a DNN that functions as a neuro-adaptive controller, and a weight optimizer for the DNN.
Before proceeding to the controller design, some mathematical preliminaries are revisited in Section \ref{sec:sub:math preliminaries}. 
Section \ref{sec:sub:NAC} introduces the neuro-adaptive controller, followed by the DNN model in Section \ref{sec:sub:NN definition}. 
The constrained optimization-based weight adaptation laws are presented in Section \ref{sec:adap_laws}.

\begin{figure}[!t]
    \centering
    \includegraphics[width=.99\linewidth]{src/figures/Controller.drawio.pdf}
    % \includesvg[width=0.75\linewidth]{Controller.drawio.svg}
    \caption{Architecture of the constrained optimization-based neuro-adaptive controller (CONAC). \MSRY{SAT은 좀더 명확히 하기 위해서 사용하였음; 그렇지 않으면 아래 실험 피규어를 통해 포화를 넘어선 제어가 이ㅔㅂ력된다고 생각할까봐 그렇습니다; ddt의 경우 수식과 통일하기 위함이 맞습니다.}}
    \label{fig:ctrl:diagram}
\end{figure}

\subsection{Mathematical Preliminaries}\label{sec:sub:math preliminaries}

The following proposition and lemma will be used in subsequent sections:

\begin{propsit}[see, {\cite[Chap. 7 Prop. 7.1.9]{Bernstein:2009aa}}] \label{propsit:kron}
	For matrix $\mm A\in\R^{n\times m}$ and vector $\mv b\in\mathbb R^n$, the following property holds:
	\begin{equation}
		\mm A^\top \mv b 
		= 
		\myvec(\mm A^\top\mv b)
		=
		\myvec(\mv b^\top\mm A)
		= 
		(\mm{I}_{m}\otimes \mv b^\top) \myvec(\mm A)
		.
	\end{equation}
\end{propsit}

\begin{lem} \label{lem:stable:set}
    For Lyapunov function $V:=V(\mv{x})\ge0$ with $\mv{x}\in\R^n$, if the time derivative of $V$ is given by $\ddtt V\le -a_1 \mv{x}^\top \mv{x} + a_2 \mv{x}$ for some positive constants $a_1\in\R$ and $a_2\in\R$, then $\mv{x}$ is bounded as $\mv{x}\in\{\mv{x}\mid\norm{\mv{x}}\le\tfrac{a_2}{a_1}\}$.
\end{lem}

\begin{proof}
    The time derivative of $V$ can be rewritten as
    \begin{equation}
        \ddtt V
        \le
        -a_1 \norm{\mv{x}}^2 + a_2 \norm{\mv{x}}
        =
        -a_1
        \left(
            \norm{\mv{x}}-\tfrac{a_2}{2a_1}
        \right)^2
        -
        \tfrac{a_2^2}{4a_1}
        .
        \label{eq:lem:lya:dot}
    \end{equation}
    According to \eqref{eq:lem:lya:dot}, one can conclude that $\ddtt V$ is negative definite if $\norm{\mv{x}}\le\tfrac{a_2}{a_1}$ holds.
    In other words, $\mv{x}$ remains within the bounded set $\{\mv{x}\mid\norm{\mv{x}}\le\tfrac{a_2}{a_1}\}$.
\end{proof}

\subsection{Neuro-Adaptive Controller Design}\label{sec:sub:NAC}

First, the filtered tracking error $\fe\in\R^n$ is defined as 
\begin{equation}
    \fe := \ddtt{\mv e}-\mm\Lambda \mv e
    ,
    \label{eq:err:filtered}
\end{equation}
where $\mv e := \q - \qd$ denotes the tracking error, $\ddtt{\mv e} := \dq - \dqd$ represents the time derivative of the tracking error, and $\mm\Lambda\in\R^{n\times n}_{>0}$ is a user-designed filtering matrix.
Since \eqref{eq:err:filtered} is stable system, it implies that $\mv e$ is bounded if $\fe$ is bounded.

Using $\fe$, the system dynamics \eqref{eq:sys1} can be rewritten as
\begin{equation}
    \rbM \ddtt\fe
    =
    -\rbVm \fe
    -\mm K \fe
    + \mv f
    -\rbd + \mysat(\rbu)
    ,
    \label{eq:sys:filtered}
\end{equation}
where $\mm{K}=\mm{K}^\top\in\R^{n\times n}_{>0}$ denotes an arbitrary unknown matrix and $
    \mv f:= \mv f(\q,\dq,\qd,\dqd,\ddqd)
    =
    \mm K \fe
    +\rbM(-\ddqd+\mm\Lambda\ddtt{\mv e})
    +
    \rbVm(-\dqd+\mm\Lambda\mv e)
    -
    \rbF
    -
    \rbG
    \in\R^n
$ denotes the lumped system uncertainty.
% \color{red}
% The unknown matrix $\mm{K}$ is not available to design in practice, 
% \color{black}

Consider the Lyapunov function $V_1:= \tfrac{1}{2}\fe^\top \rbM\fe$. 
Invoking Property \ref{prop:skew}, the time derivative of $V_1$ is
\begin{equation}
    \begin{aligned}
        \ddtt{V}_1
        =&
        \fe^\top \rbM \ddtt \fe
        +
        \tfrac{1}{2}\fe^\top \ddtt{\rbM}\fe
        \\
        =&
        \fe^\top (
            -\rbVm \fe -\mm K \fe + \mv f
            -\rbd + \mysat(\rbu)
        )
        \\
        &
        +
        \tfrac{1}{2}
        \fe^\top \ddtt{\rbM}\fe
        \\
        =&
        -
        \fe^\top \mm K \fe 
        +
        \fe^\top (
            \mv f+\mysat(\rbu)
            -\rbd
        )
        \\
        &
        +
        \tfrac{1}{2}
        \fe^\top(\ddtt{\rbM}- 2\rbVm)\fe
        \\
        \le&
        -\lambda_{\min}(\mm K)\norm{\fe}^2
        +
        \overline\tau_d\norm{\fe}
        +
        \fe^\top (\mv f+\mysat(\rbu))
        % \\
        % \le&
        % -\lambda_{\min}(\mm K)\norm{\fe}^2
        % +
        % \overline\tau_d\norm{\fe}
        % +
        % \fe^\top (\mysat(\rbu)-\rbu^*)
        .
    \end{aligned}
    \label{eq:lya1:dot1}
\end{equation} 
Therefore, invoking Lemma \ref{lem:stable:set}, it can be concluded that the filtered error $\fe$ is exponentially stable, with $\lim_{t\to\infty}\norm{\fe}=\tfrac{\overline\tau_d}{\lambda_{\min}(\mm K)}$, provided that the lumped system uncertainty $\mv{f}$ can be perfectly cancelled by the control input $\rbu$, neglecting the effect of input saturation.
Hence, the ideal control input $\rbu^*$ is defined as $-\mv f$.
However, $\rbu^*$ is not available in practice, since system dynamics encapsulated in $\mv f$ are unknown.

To overcome this issue, a DNN is employed to approximate $\rbu^*$.
Let $\NN:=\NN({\q}_n;\wth): \R^{l_0+1}\times\R^{\Xi}\to\R^{n}$ represent the DNN, where ${\q}_n\in\R^{l_0+1}$ is the DNN input vector, and $\wth\in\R^{\Xi}$ is the vector of trainable weights.
The architecture of $\NN({\q}_n;\wth)$ will be presented in Section \ref{sec:sub:NN definition}.
According to the universal approximation theorem for DNNs \cite{Kidger:2020aa}, $\NN({\q}_n;\wth)$ can approximate a nonlinear function, denoted as $\mv g(\cdot)$, with an ideal weight vector $\idealwth$ over a compact subset $\Omega_{NN}\in\R^{l}$ to within $\epsilon$-accuracy, such that $\sup_{{\q}_n\in\Omega_{NN}}\norm{ \NN({\q}_n;\idealwth) - \mv g(\cdot)} = \epsilon < \infty$.
% Furthermore, the theorem states that the norm of $\idealwth$ is bounded, \ie $\norm{\idealwth}\le \overline\wth<\infty$.
In this study, $\idealwth$ is considered a locally optimal solution, rather than a global one.

Accordingly, the ideal control law $\rbu^*$ can be approximated by the DNN with the ideal weight vector $\idealNN:=\NN({\q}_n;\idealwth)$ as follows:
\begin{align}
    \rbu^*=& \idealNN+\mv\epsilon
    ,
    \label{eq:control:ideal}
\end{align}
where $\mv\epsilon\in\R^n$ is the approximation error vector, bounded by $\norm{\mv\epsilon}\le \bar{\epsilon}$.
The control input $\rbu^*$ is estimated online by
\begin{align}
    \rbu =& \estNN,
    \label{eq:control:est}
\end{align}
where $\hat\NN:=\NN({\q}_n;\estwth)$, and  $\estwth$ is the estimated weight vector for $\idealwth$.

Substituting \eqref{eq:control:ideal} and \eqref{eq:control:est} into \eqref{eq:lya1:dot1}, the time derivative of $V_1$ can be rewritten as
\begin{equation}
    \ddtt{V}_1
    \le 
    -\lambda_{\min}(\mm K)\norm{\fe}^2
    +
    \overline\tau_d\norm{\fe}
    +
    \fe^\top (\mysat(\hat\NN)-\idealNN-\mv\epsilon)
    .
    \label{eq:lya1:dot2}
\end{equation}
Therefore, one can conclude that the result of \eqref{eq:lya1:dot1} can be obtained by adapting $\estwth$ to $\idealwth$ (\ie $\hat\NN\to\idealNN$), neglecting the effect of input saturation.
Note that the control input saturation has not yet been considered in the derivation of the ideal control input; this will be addressed in the subsequent section.

\subsection{Deep Neural Network (DNN) Model}\label{sec:sub:NN definition}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.99\linewidth]{src/figures/DNN.drawio.pdf}
    \caption{
        Architecture of the DNN $\NN({\q}_n;\wth)$ with $k=2,l_0=2,l_1=l_2=3$, and $l_3=2$.
    }
    \label{fig:DNN}
\end{figure}

The DNN architecture $\NN({\q}_n;\wth) := \NN_k$ can be recursively represented as follows:
\begin{equation}
    \NN_i :=
    \begin{cases}
        \wV_i^\top \act_i(\NN_{i-1}), 
        &
        i\in\{1,\dots ,k\},
        \\
        \wV_0^\top {\q}_n,
        &
        i=0
        ,
    \end{cases}
    \label{eq:DNN:architecture}
\end{equation}
where $\wV_i\in\R^{(l_i+1)\times l_{i+1}}$ is the weight matrix of the $i\textsuperscript{th}$ layer, and $\act_i: \R^{l_i}\to\R^{l_i+1}$ represents the activation function of the $i\textsuperscript{th}$ layer. 
The activation function is defined as $\act_i(\mv x)=\left(\sigma(x_1),\sigma(x_2),\cdots, \sigma(x_{l_{i}}), 1\right)^\top,\ \forall\mv x\in\R^{l_i}$, where $\sigma: \R\to\R$ is a nonlinear function, and the augmentation of $1$ is used to account for bias terms in the weight matrices. 
Notice that the output size of $\NN(\cdot)$ is the same as that of the control input $\rbu$ (\ie $l_{k+1}=n$).

One of the widely used activation functions for large DNNs is from the ReLU family \cite{Maas:2013aa}, which effectively avoids the gradient vanishing problem during error backpropagation. 
However, for control applications, where relatively shallow DNNs are typically sufficient, and the gradient vanishing issue is less severe, the sigmoid function or the hyperbolic tangent function is commonly used as the activation function. 
These functions simplify stability analysis, since they are continuously differentiable and their outputs and gradients are bounded. 
In this study, the hyperbolic tangent function $\tanh(\cdot)$ was selected as the activation function (\ie $\sigma(x) = \tanh(x),\ \forall x\in\R$), which provides desirable boundedness with $\norm{\sigma(x)}<1$ and $\norm{\ddtfrac{\sigma(x)}{x}}\le 1$.

For simplicity, each layer's weights are vectorized as $\wth_i:=\myvec(\wV_i)\in\R^{\Xi_i}$, where $\Xi_i:= (l_i+1)l_{i+1}$ is the number of weights in the $i\textsuperscript{th}$ layer. 
The total weight vector $\wth\in\R^{\Xi}$ is defined by augmenting $\wth_i$ for all $i\in \{0,\cdots,k\}$ as 
\begin{equation}
    \wth := 
    \begin{pmatrix}
        \wth_k\\
        \wth_{k-1}\\
        \vdots\\
        \wth_0
    \end{pmatrix}
    =
    \begin{pmatrix}
        \myvec(\wV_k)\\
        \myvec(\wV_{k-1})\\
        \vdots\\
        \myvec(\wV_0)
    \end{pmatrix},
\end{equation}
where $\Xi={\sum_{i=0}^{k} \Xi_i}$ represents the total number of weights. 
The gradient of $ \NN({\q}_n;\wth)$ with respect to $\wth$ can be obtained using Proposition \ref{propsit:kron} and the chain rule as follows:
\begin{equation}
    \pptfrac{\NN}{\wth}=
    \begin{bmatrix}
        \pptfrac{\NN}{\wth_k}&
        \pptfrac{\NN}{\wth_{k-1}}&
    \cdots &
        \pptfrac{\NN}{\wth_0}
    \end{bmatrix}
    \in\R^{n \times \Xi}
    \label{eq:gradient:phi}
\end{equation}
where
\begin{equation}
    \pptfrac{\NN}{\wth_i} = 
    \begin{cases}
        (
            \mm I_{l_{k+1}}
            \otimes 
            \act_{k}^\top  
        )
        , 
        &
        i=k
        ,
        \\
        \wV_k^\top\act_{k}' 
        (
            \mm I_{l_{k}}
            \otimes  
            \act_{k-1}^\top  
        )
        , 
        & 
        i=k-1
        ,
        \\
        &
        \vdots 
        \\
        \wV_k^\top\act'_{k} 
        \cdots 
        \wV_1^\top\act_1' 
        (
            \mm I_{l_1}
            \otimes 
            {\q}_n^\top  
        )
        , 
        &
        i = 0
        ,
    \end{cases}
    \label{eq:NN:grad}
\end{equation}
and $\act_i:= \act_i(\NN_{i-1})$ and $\act_i':= \ddtfrac{\act_i}{\NN_{i-1}}$.

In the following sections, let $\idealNN_i$ represent the output of the $i\textsuperscript{th}$ layer with the ideal weight vector $\idealwth$. 
Additionally, define $\idealact_i:=\act_i(\idealNN_{i-1})$ and ${\idealact}'_i:= \ddtfrac{\idealact_i}{\idealNN_{i-1}}$. 
Similarly, $\estNN_i$ denotes the output of the $i\textsuperscript{th}$ layer with the estimated weight vector $\estwth$, and define $\estact_i:=\act_i(\estNN_{i-1})$ and $\estact_i':= \ddtfrac{\estact_i}{\estNN_{i-1}}$.

%  SECTION ADAPTATION LAW DERIVATION =======================
\section{Weight Adaptation Laws}\label{sec:adap_laws}

% \subsection{Adaptation Law using Lagrangian Function}
\subsection{Weight Optimizer Design}\label{sec:sub:weight optimizer}

The control objective can be represented as follows:
\begin{equation}
    \begin{matrix}
        \min_{\estwth} \ J(\fe;\estwth):= 
        \tfrac{1}{2} \fe^\top \fe
        \\ \\
        \begin{aligned}
        \text{subject to }&c_{j}(\estwth) 
        \le0, \quad \forall j\in\mathcal{I},
        \end{aligned}
    \end{matrix}
    \label{eq:opt:problem1}
\end{equation}
where $J(\fe;\estwth):= \tfrac{1}{2} \fe^\top \fe$ is the objective function.
Inequality constraints $c_j:=c_j(\estwth),\ \forall j\in\mathcal{I}$, are imposed during the weight adaptation process to minimize the objective function, where $\mathcal I$ denotes the set of the imposed inequality constraints. 
Here, $\fe$ is considered a pre-defined data or static parameter for this optimization problem, so that the adaptation process does not consider the dynamics of $\fe$.
The Lagrangian function is defined as
\begin{equation}
    L(
        \fe,\estwth,[\lambda_j]_{j\in\mathcal I}
    ) 
    := 
    J(\fe;\estwth) 
    + 
    \sum_{j\in\mathcal I}
    \lambda_{j}
    c_{j}(\estwth)
    ,
    \label{eq:lagrangian}
\end{equation}
where $\lambda_j$ denotes the Lagrange multiplier for each constraint.

The adaptation laws for $\estwth$ and $[\lambda_j]_{j\in\mathcal I}$ are derived to solve the dual problem of \eqref{eq:opt:problem1} (\ie  $\min_{\estwth} \max_{[\lambda_j]_{j\in\mathcal I}}L(\fe,\estwth,[\lambda_j]_{j\in\mathcal I})$), as follows:
\begin{subequations}
    \begin{align}
            \ddtt{\estwth}
            &
            =
            -\alpha \pptfrac{L}{\estwth}
            =
            -\alpha 
            \left(
                \pptfrac{J}{\estwth}
                +
                \textstyle\sum_{j\in\mathcal{I}}
                \lambda_j 
                \pptfrac{c_j}{\estwth}
            \right),
        \label{eq:adap:th}
            \\
            \ddtt\lambda_j
            & 
            = 
            \beta_j\pptfrac{L}{\lambda_j} 
            = 
            \beta_j c_j ,
            \quad \forall j\in\mathcal I,
        \label{eq:adap:lbd}
            \\
            \lambda_j & \leftarrow \max(\lambda_j,0) ,
        \label{eq:adap:lbd:max}
    \end{align}
    \label{eq:adaptation law}%             %%
\end{subequations}
where $\alpha\in\R_{>0}$ denotes the adaptation gain (learning rate) and $\beta_j\in\R_{>0}$ denotes the update rate of the Lagrange multipliers in $\mathcal I$, and the arguments of $L(\cdot)$ and $J(\cdot)$ are suppressed for brevity. 
% By selecting $\alpha_i\in\R_{>0},\ \forall i\in[1,\cdots,n]$, one can weight the importance of each component of the gradient of the objective function, allowing for a more tailored adaptation process.
The Lagrange multipliers associated with inequality constraints are non-negative, \ie $\lambda_j\ge 0$, due to \eqref{eq:adap:lbd:max}. 
When a constraint $c_j$ becomes active, \ie $c_j > 0$, the corresponding Lagrange multiplier $\lambda_j$ increases to address the violation, \ie see, \eqref{eq:adap:lbd}.
Once the violation is resolved and the constraint is no longer active, \ie $c_j \le 0$, the multiplier decreases gradually until it returns to zero.
Note that this adaption law is similar to the ALM in \cite{Nocedal:2006aa}, where the adaptation law for Lagrange multipliers is given by $\lambda_j\leftarrow \max\left(\lambda_j+\tfrac{c_j}{\mu},0\right)$, with $\mu\in\R_{>0}$ being the penalty parameter.

At steady state, where $\ddtt{\estwth}=0$ and $\ddtt\lambda_j=0$, the KKT conditions are satisfied, \ie $\pptfrac{L}{\estwth}=0$, $c_j \le 0$, $\lambda_j \ge 0$, and $\lambda_j c_j=0$.
In other words, the proposed optimizer updates $\estwth$ and $\lambda_j$ in a way that satisfies the KKT conditions. 
These conditions represent the first-order necessary conditions for optimality, guiding the updates toward candidates for a locally optimal point.
% The KKT condition is satisfied, when $\estwth$ converges to $\idealwth$, since $\partial L/\partial \estwth=0$ (i.e. the first-order necessary condition of the optimality is satisfied.).

\subsection{Approximation of the Gradient of Objective Function}

The adaptation law for $\estwth$ defined in \eqref{eq:adap:th} involves the partial derivative of the filtered error with respect to control input $\pptfrac{\fe}{\rbu}$, \ie 
$
    \pptfrac{J}{\rbu}
    =
    \left(
        (\pptfrac{\fe}{\rbu})
        (\pptfrac{\rbu}{\estwth})
    \right)
    ^\top \fe
    =
    \left(
        (\pptfrac{\fe}{\rbu})
        (\pptfrac{\estNN}{\estwth})
    \right)
    ^\top \fe
$. 
Since the objective function depends on $\fe$ of a dynamic system, obtaining $\pptfrac{\fe}{\rbu}$ is not straightforward. 
The recommended method to calculate the exact value of $\pptfrac{\fe}{\rbu}$ is to use the forward sensitivity method \cite{Sengupta:2014aa} by simulating the sensitivity equation as follows: 
\begin{equation}
    \ddtt 
    \left(
        \pptfrac{\fe}{\rbu}
    \right)
    =
    \pptfrac{}{
        \rbu
    }
    \left(
        \rbM^{-1}(
            -\rbVm \fe
            -\mm K \fe
            + \mv f
            -\rbd + \mysat(\rbu)
        )
    \right)
    .
\end{equation}
However, this method cannot be realized, since we do not have the exact system dynamics, \ie $\rbM,\rbVm, \rbF$ and $ \rbG$.
Additionally, the computational cost of the forward sensitivity method is high, as the number of DNN's weights are generally large.

\KHCH{control direction과 sign of control input이 서로 다른 의미를 가지고 있지만 독자가 헷갈려할 수도 있을 것 같습니다. 이 부분 고민해주세요.}
In \cite{Douratsos:2007aa,Saerens:1991aa}, the authors approximate $\pptfrac{\fe}{\rbu}$ by taking the sign of each entry, \ie
$
    \pptfrac{\fe}{\rbu}
    \approx
    \left[
        \mysign(\pptfrac{r_i}{\tau_j})
    \right]_{i,j\in\{1,\cdots,m\}}
$ 
assuming known control directions.
However, this approach is not applicable to \eqref{eq:sys:filtered}, where control directions are unknown.
Instead, since the sign of the control input is known—owing to the positive definiteness of $\rbM^{-1}$ (see, Property \ref{prop:M})—we approximate $\pptfrac{\fe}{\rbu}\approx \mm I_n$.
Consequently, the adaptation law in \eqref{eq:adap:th} becomes:
\begin{equation}
    \ddtt{\estwth}
    \approx
    -\alpha 
    \left(
        \pptfrac{\estNN}
        {\estwth}^\top
        \fe
        +
        \textstyle\sum_{j\in\mathcal{I}}
        \lambda_j 
        \pptfrac{c_j}{\estwth}
    \right)
    .
    \label{eq:adap:th:approx}
\end{equation}

%% ALGORITHM PART; Now omitted!
% The proposed controller is implemented using Algorithm \ref{alg: alg1}. For implementation in the discrete-time domain, it is recommended to use a sufficiently small sampling time $T_s$. If a large $T_s$ is used, $\alpha$ and $\beta_j$ should satisfy the Armijo condition \cite[Chap.~3 eq.~(3.4)]{Nocedal:2006aa} to ensure that the objective function decreases.

% \begin{algorithm}[!t]
%     \caption{Weight Optimizer Implementation.}\label{alg: alg1}
%     \begin{algorithmic}[1]
%         \renewcommand{\algorithmicrequire}{\textbf{Input:}}
%         \renewcommand{\algorithmicensure}{\textbf{Output:}}
%         \REQUIRE $\xi$, $\estwth$, $\lambda_j$, $\eta$
%         \ENSURE  $\estwth$, $\lambda_j$, $\eta$
%         \STATE Set $\mathcal A \leftarrow \mathcal A\cup \{j\}$ for all $c_j\ge0$;
%         \STATE Determine update matrix $\ddtt\eta$ using \eqref{eq. eta dynamics 1};
%         \STATE Update $\eta\leftarrow \eta +\ddtt\eta\cdot T_s$; 
%         \STATE Determine update directions $\ddtt{\estwth}$, $[\ddtt\lambda_j]_{j\in\mathcal A}$ using \eqref{eq. adaptation law th}, \eqref{eq. adaptation law L};
%         \STATE Update weight vector $\estwth\leftarrow \estwth+\ddtt{\estwth}\cdot T_s$;
%         \STATE Update multipliers $[\lambda_j]_{j\in\mathcal A}\leftarrow [\lambda_j]_{j\in\mathcal A}+[\ddtt\lambda_j]_{j\in\mathcal A}\cdot T_s$;
%         \STATE $[\lambda_j]_{j\in\mathcal A}\leftarrow \max([\lambda_j]_{j\in\mathcal A}, 0)$;
%         \STATE Set $\mathcal A \leftarrow \mathcal A - \{j\}$ for all $\lambda_j=0$;
%         \STATE \textbf{RETURN};
%     \end{algorithmic}
%     \label{alg1}
% \end{algorithm}

\subsection{Essential and Potential Constraint Candidates}\label{sec:sub:cstr}

This section introduces essential constraints required to ensure the stability of the proposed CONAC, as well as candidate conditions for potential control input constraints.

First, weight constraints are essential to prevent the weights from diverging during the adaptation process.
The weight constraints $\mv{c}_{\theta}(\estwth):= [c_{\theta_i}(\estwth)]_{i\in\{0,\cdots ,k\}}\in\mathbb R^{k+1}$ are defined for each layer's weight as follows:
\begin{equation}
    c_{\theta_i}
    =
    \tfrac{1}{2}
    \left(
        \Vert \estwth_i\Vert^2 
        -
        \overline\theta_i^2 
    \right)    
    % \le 0
    \label{eq:cstr:weight:ball}
\end{equation}
with $\overline\theta_i<\infty$ denoting the maximum allowable norm for $\estwth_i$, and the arguments of $c_{\theta_i}(\cdot)$ are omitted for brevity.
The gradient of $\mv{c}_\theta:=\mv c_{\theta}(\estwth)$ with respect to $\estwth$ is given by
\begin{equation}
    \pptfrac{c_{\theta_i}}{\estwth_j} 
    =
    \begin{cases}
        \estwth_i,
        &
        \text{if } i=j,
        \\
        \mv{0},
        &
        \text{if } i\ne j
        .
    \end{cases} 
    \label{eq:cstr:weight:ball:grad}
\end{equation}

Next, we present the following assumptions, which specify the conditions for the potential control input constraints introduced in Appendix \ref{sec:appen:cstr}:

\begin{assum}
    The constraint functions $c_j(\estwth),\ \forall j\in\mathcal I$, are convex in the $\tau$-space and satisfy $c_j(0) \le 0$ \color{red} and $c_j(\idealwth)\le 0$ \color{black}. \MSRY{desired controller가 제약조건 밖으로 나갈 수 있기 때문에 적절하지 않을 수 있음}
    \label{assum:convex}
\end{assum}

\begin{assum}
    The selected constraints satisfy the Linear Independence Constraint Qualification (LICQ) \cite[Chap.~12 Def.~12.1]{Nocedal:2006aa}.
    \label{assum:LICQ}
\end{assum}

\begin{remark}
    Assumption \ref{assum:convex} is not restrictive, since the saturation functions are convex\KHCH{앞에 convex라는 표현이 나오지 않았는데 확인부탁합니다.}\MSRY{컨벡스란 말이 없지만 practical하게 대부분의 포화 함수는 컨벡스라는 말을 의도하였습니다.} in practice including the origin.
    Moreover, Assumption \ref{assum:LICQ} is a standard assumption in optimization problems, which ensures that the gradients of the active constraints are linearly independent.
    This assumption will be used in the stability analysis (see, Lemma \ref{lem:convex:angle}).
\end{remark}

%  SECTION STABILITY ANALYSIS ==============================
\section{Stability Analysis}\label{sec:stability}

Before conducting the stability analysis, let us define the weight estimation error as $\tilde\wth:= [\tilde\wth_i]_{i\in\{0,\cdots,k\}}$, where $\tilde\wth_i:=\estwth_i-\idealwth_i$.
The following lemmas are introduced for the stability analysis.

\begin{lem}
    If Assumptions \ref{assum:convex} and \ref{assum:LICQ} hold, then for each active constraint $c_j$, the angle between $\pptfrac{c_j}{\estwth_k}$ and $\estwth_k$ is positive; that is $\pptfrac{c_j}{\estwth_k}^\top\estwth_k>0$.
    \label{lem:convex:angle}
\end{lem}

\begin{proof}

Since $\rbu = \estNN$, using Proposition \ref{propsit:kron}, a linear mapping $\mm T(\cdot):\estwth_k\to\rbu$ that is independent of $\estwth_k$, can be derived using Proposition \ref{propsit:kron}:
\begin{equation}\label{eq:linear:map}
    \begin{aligned}
    \rbu 
    = 
    &
    \estNN 
    % = 
    % \myvec(\estNN)
    =
    \myvec(
        \estwV_k^\top \estact_k
    ) 
    = 
    (
        \mm I_{l_{k+1}}
        \otimes 
        \estact_k^\top
    )^\top
    \myvec(\estwV_k)
    \\
    = &
    \underbrace{
        (
        \mm I_{l_{k+1}}
        \otimes 
        \estact_k^\top
    )^\top}_{=:\mm T(\estact_k)}
    \estwth_k 
    =
    \mm T(\estact_k) \estwth_k.
    \end{aligned}
\end{equation}
Therefore, the convexity of the input constraints in the $\rbu$-space (see, Assumption \ref{assum:convex}) is preserved in $\estwth_k$-space, implying
that $\pptfrac{c_j}{\estwth_k}^\top\estwth_k>0$.

\end{proof}

\begin{lem} 
    If control input constraint $c_j(\estwth),\ \forall j\in\mathcal I \setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}$ satisfies Assumption \ref{assum:convex}, then $\norm{\pptfrac{c_j}{\estwth_i}},\ \forall i\in\{k-1,\cdots, 0\}$, is bounded, provided the norms of $\estwth_i,\ \forall i\in\{k,\cdots, i+1\}$, remain bounded.
    \label{lem:cstr:grad:bound}
\end{lem}

\begin{proof}

The derivative of $c_j,\ \forall j\in\mathcal I \setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}$, with respect to $\estwth_i$ is represented as
\begin{equation}
    \pptfrac{c_j}{\estwth_i} 
    = 
    \pptfrac{c_j}{\rbu} 
    \pptfrac{\rbu}{\estNN} 
    \pptfrac{\estNN}{\estwth_i}
\end{equation}
where $\pptfrac{\tau}{\estNN}=\mm I_n$, which is bounded. 
From the linear mapping in \eqref{eq:linear:map}, $\rbu$ is bounded as long as $\estwth_k$ is bounded (by the condition of the lemma), and $\norm{\act_k(\cdot)}$ is bounded due to the properties of the activation functions. 
By Assumption \ref{assum:convex}, the function $c_j$ is convex in $\rbu$-space.
The convex function has a bounded derivative with respect to $\rbu$, since $\rbu$ is a bounded variable (\ie $\pptfrac{c_j}{\rbu}$ is bounded). 
Furthermore, $\pptfrac{\estNN}{\estwth_i}$ is bounded, provided that the norms of $\estwth_i,\ \forall i \in[k,\cdots,i+1]$, are bounded. 
This can be verified by using the definition of $\pptfrac{\estNN}{\estwth_i}$ given in \eqref{eq:NN:grad}.
Consequently, $\norm{\pptfrac{c_j}{\estwth_i}},\ \forall j\in\mathcal I \setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}$, is bounded, when $\estwth_i,\ \forall i\in \{k,\cdots,i+1\}$ are bounded.

\end{proof}

The following theorem states that $\estwth$ and $\fe$ are bounded.

\begin{theorem}
    For the dynamical system described in \eqref{eq:sys1}, the neuro-adaptive controller in \eqref{eq:control:est} with the weight adaptation laws in \eqref{eq:adaptation law} ensure the boundedness of the filtered error $\fe$ and the weight estimate $\estwth$, under the control input constraints satisfying Assumption \ref{assum:convex} and \ref{assum:LICQ}.
    This holds with the weight norm constraint \eqref{eq:cstr:weight:ball}.
\end{theorem}

\begin{proof}

% The boundednesses of $\estwth$ and $\fe$ are analyzed recursively from the last $k\textsuperscript{th}$ layer to the first layer of $\estNN$. 
Invoking the fact that the amplitude of the control input depends only on $\estwth_k$, \ie see, \eqref{eq:control:est}, \eqref{eq:DNN:architecture} and $\norm{\act_k}<\infty$, the boundedness of inner layers' weights follows once the boundedness of $\estwth_k$ and $\fe$ is established.
Without loss of generality, weight constraint $c_{\theta_i},\ \forall i\in\{0,\cdots,k\}$ is supposed to be active.
This assumption is justified because, even if $c_{\theta_i}$ is inactive, $\estwth$ is adapted to minimize the objective function $J(·)$ as long as the constraint is not violated.
We consider two cases: (1) the control input is saturated, and (2) the control input is not saturated.

\subsection*{Case 1: Control input is saturated.}

As a\MSRY{a?} result of the time derivative of $V_1$ in \eqref{eq:lya1:dot2}, the invariant set $\Theta_{\fe}^{1}$ of $\fe$ (\ie if $\fe$ leaves $\Theta_{\fe}^{1}$, $\ddtt V_1$ is negative) can be obtained using Lemma \ref{lem:stable:set} as follows:
\begin{equation}
    \Theta_{\fe}^{1} 
    = 
    \left\{ 
        \fe \in \R^n 
        \mid 
        \norm{\fe} 
        \le 
        \tfrac{
            \overline\tau_d
            +
            \overline\tau
            +
            \bar{\theta}_k\sqrt{l_{k}+1}
            +
            \bar{\epsilon}
        }{
            \lambda_{\min}(\mm K)
        }
    \right\}
    .
    \label{eq:invariant:c1:fe}
\end{equation}
It is notable that the ideal DNN $\idealNN_k={\idealwV_k}^\top\idealact$ is bounded by $\bar{\theta}_k\sqrt{l_{k}+1}$, since $\norm{\idealwV_k}_F = \norm{\idealwth}_k \le \bar{\theta}_k$, and $\norm{\idealact_k}\le \sqrt{l_{k}+1}$.

\MSRY{Explain with active set, later}
To investigate the boundedness of $\estwth_k$, consider the Lyapunov function candidate $V_2:=\tfrac{1}{2\alpha}\estwth_k^\top \estwth_k$.
Taking the time derivative of $V_2$ yields:
\begin{equation}
    \begin{aligned}
        \ddtt  V_2 
        =& 
        -\estwth_k^\top 
        \left(
            (\mm I_{l_{k+1}}\otimes \estact_k^\top)^\top
            \fe
            +
            \textstyle\sum_{j\in\mathcal{I}}
            \lambda_j 
            \pptfrac{c_j}{\estwth_k}
        \right)
        \\
        =&
        -\estwth_k^\top 
        \Big(
            (\mm I_{l_{k+1}}\otimes \estact_k^\top)^\top
            \fe
            +
            \lambda_{\theta_k}\estwth_k
        \\
        &
            +
            \textstyle\sum_{
                j\in\mathcal{I}\setminus \{\wth_i\}_{i\in[0,\cdots,k]}
            }
            \lambda_j 
            \pptfrac{c_j}{\estwth_k}
        \Big)
        \\
        \le&
        -
        \lambda_{\theta_k}\norm{\estwth_k}^2
        +
        \underbrace{
            \norm{(\mm I_{l_{k+1}}\otimes {\estact}_k^\top)}
        }_{=:{\gamma}_1\in\R_{>0}}
        \norm{\fe}
        \norm{\estwth_k}
        \\
        &
        \underbrace{
            -
            \textstyle\sum_{
                j\in\mathcal{I}\setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}
            }
            \lambda_j 
            \estwth_k^\top \pptfrac{c_j}{\estwth_k}
        }_{=:{\gamma}_2<0,\ \text{by Lemma \ref{lem:convex:angle} and Assumption \ref{assum:LICQ}}}
        \\
        \le&
        -
        \lambda_{\theta_k}\norm{\estwth_k}^2
        +
        {\gamma}_1     
        \norm{\fe}
        \norm{\estwth_k}
        .
    \end{aligned}
    \label{eq:lya2:dot}
\end{equation}
According to \eqref{eq:lya2:dot} and Lemma \ref{lem:stable:set}, the invariant set of $\estwth_k$ is defined as
\begin{equation}
    \Theta_{\theta_k}^{1} 
    = 
    \left\{ 
        \estwth_k \in \R^{\Xi_{k}} 
        \mid 
        \norm{\estwth_k} 
        \le 
        \tfrac{
            {\gamma}_1
            (
                \overline\tau_d
                +
                \overline\tau
                +
                \bar{\theta}_k\sqrt{l_{k}+1}
                +
                \bar{\epsilon}
            )
        }{
            \lambda_{\theta_k}\lambda_{\min}(\mm K)
        }
    \right\}
    .
    \label{eq:invariant:c1:wth}
\end{equation}

The satisfaction of the constraints can be verified through the boundedness of $\estwth_k$.
For the weight constraint $c_{\theta_k}$, if the corresponding Lagrange multiplier $\lambda_{\theta_k}$ increases infinitely due to the constraint violation, $\estwth_k$ approaches to the origin, as the invariant set $\Theta_{\estwth_k}^{1}$ shrinks to a point.
The remaining control input constraints can be validated implicitly through the term ${\gamma}_2$ in \eqref{eq:lya2:dot}.
Similarly to $c_{\theta_k}$, when a control input constraint $c_j,\ \forall j\in\mathcal{I}\setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}$ is violated, the corresponding Lagrange multiplier $\lambda_j$ increases infinitely, leading to increase in ${\gamma}_2$.
It makes ${\gamma}_2$ dominates the right-hand side of \eqref{eq:lya2:dot}, which leads $\ddtt V_2$ negative definite.
This drives $\estwth_k$ to the origin until the constraint is satisfied.
By Assumption \ref{assum:convex} and Lemma \ref{lem:convex:angle}, the control input constraints can be considered as convex constraint in the $\estwth_k$-space.

Therefore, the boundedness of $\estwth_k$ and $\fe$ is guaranteed by the invariant sets $\Theta_{\fe}^{1}$ and $\Theta_{\theta_k}^{1}$, and the satisfactions of $c_j,\ \forall j\in\mathcal{I}$ is ensured.

\subsection*{Case 2: Control input is not saturated.}

Since, the control input constraints are inactive, the saturation function $\mysat(\cdot)$ in \eqref{eq:lya1:dot2} and the constraint functions $c_j,\ \forall j\in\mathcal I \setminus \{\wth_i\}_{\{i\in[0,\cdots,k\}}$ in \eqref{eq:lya2:dot} can be removed.
Consider the Lyapunov function candidate $V_3:=V_1+V_2$, whose time derivative is given by:
\begin{equation}
    \begin{aligned}
        \ddtt V_3
        =&
        -\lambda_{\min}(\mm K)\norm{\fe}^2
        +\overline\tau_d\norm{\fe}
        +\fe^\top (\estNN - \idealNN - \mv\epsilon)
        \\
        &
        -\estwth_k^\top (
            (\mm I_{l_{k+1}}\otimes \estact_k^\top)^\top
            \fe
            +
            \lambda_{\theta_k}\estwth_k
        )
        \\
        \le&
        -\lambda_{\min}(\mm K)\norm{\fe}^2
        +\fe^\top \estNN
        +(
            \overline\tau_d+\norm{
                \idealNN+\mv\epsilon
            }
        )
        \norm{\fe}
        \\
        &
        -\estNN^\top\fe 
        -\lambda_{\theta_k}\norm{\estwth_k}^2
        \\
        \le&
        -
        \lambda_{\min}(\mm K)
        \norm{\fe}^2
        +
        \left(
            \overline\tau_d
            +
            \bar{\theta}_k\sqrt{l_{k}+1}
            +
            \bar{\epsilon}
        \right)
        \norm{\fe}
        \\
        &
        -
        \lambda_{\theta_k}\norm{\estwth_k}^2
        .
    \end{aligned}
    \label{eq:lya3:dot}
\end{equation}
Following the same reasoning as in Case 1, the boundedness of $\fe$ is ensured by the invariant set:
\begin{equation}
    \Theta_{\fe}^{2}
    =
    \left\{ 
        \fe \in \R^n 
        \mid 
        \norm{\fe} 
        \le 
        \tfrac{
            \overline\tau_d
            +
            \bar{\theta}_k\sqrt{l_{k}+1}
            +
            \bar{\epsilon}
        }{
            \lambda_{\min}(\mm K)
        }
    \right\}
    .
\end{equation}
The boundedness of $\estwth_k$ also can be ensured by the invariant set $\Theta_{\theta_k}^{2}=\{ \estwth_k \in \R^{\Xi_{k}} \mid \norm{\estwth_k} \le \overline\theta_k \}$, as $\lambda_{\theta_k}$ remains positive unless the weight constraint $c_{\theta_k}$ is satisfied.

\hfill

The boundedness of the inner-layer weights and the satisfaction of their corresponding constraints $c_j,\ \forall j\in\{\theta_i\}_{i\in\{k-1,\cdots,0\}}$, can be established recursively using \cite[Chap.~4 Thm.~1.9]{Desoer:2009aa}.
The dynamics of $\estwth_i,\ \forall i\in\{k-1,\cdots,0\}$, are represented as
\begin{equation}
    \ddtt \estwth_i 
    =
    -\alpha
    \left(
        \pptfrac{\estNN}{\estwth_i}^\top
        \fe
        +
        \lambda_{\theta_i}\estwth_i
        +
        \textstyle\sum_{
            j\in\mathcal{I}\setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}
        }
        \lambda_j
        \pptfrac{c_j}{\estwth_i}
    \right)
    .
\end{equation}
By Lemma \ref{lem:cstr:grad:bound} $\estwth_i$ is bounded, provided that $\estwth_i,\ \forall i\in\{k,\cdots,i+1\}$, are bounded, since the system matrix $-\lambda_{\theta_i} \mm I_{\Xi_i}$ is stable and the residual terms are bounded.
Therefore, starting from the $(k-1)\textsuperscript{th}$ layer, the boundedness of $\estwth_i$ can be established recursively down to the input layer ($i=0$).

\hfill

In conclusion, the filtered tracking error $\fe$ and the entire set of estimated weights $\estwth$ are bounded, and all imposed constraints $c_j,\ \forall j\in\mathcal I$ are satisfied.

\end{proof}

%  SECTION SIMULATION ======================================
\section{Real-Time Implementation and Validation}\label{sec:sim}

\subsection{Validation Setup}

% ------------------------------------
% Real-time Implementation Set-up
% ------------------------------------
\begin{figure}[t]
    \centering
        \includegraphics[width=\figSizeOneCol\linewidth]
        {
            src/figures/exp_set.drawio.png
        }%
    \caption{
        Experimental setup of the two-link robotic manipulator used for real-time validation.
    }
    \label{fig:ctrl:exp:set}
\end{figure}

To validate the effectiveness of the proposed CONAC, a real-time experiment was conducted on a two-link manipulator, as shown in Fig.~\ref{fig:ctrl:exp:set}. 
The OpenCR1.0 board \cite{opencr} with a 32-bit ARM Cortex and a 216 MHz clock frequency was used for the real-time implementation.
The two-link manipulator actuated by two Cubemars AK10-90 servo motors. 
The servos were controlled using the OpenCR1.0 board which transmits the reference angle to the corresponding servo via controller area network (CAN) communication.
The system dynamics of the two-link manipulator can follow the Euler-Lagrange equation \eqref{eq:sys1} and are illustrated in Fig.~\ref{fig:robot:model}. 
The estimated system parameters are provided in Table~\ref{table:system:params}.

% ------------------------------------
% Robot Model Figure and Table
% ------------------------------------
\begin{figure}[t]
    \centering
    \subfloat[Two-link manipulator model.\MSRY{ADD inertia}]{
        \includegraphics[width=0.75\linewidth]{src/figures/RobotModel.drawio.pdf}
        \label{fig:robot:model}
    }
    \hfill
    \subfloat[Control input saturation function.]{
        \includegraphics[width=0.55\linewidth]{src/figures/saturation.drawio.pdf}
        \label{fig:robot:sat}
    }
    \caption{Two-link manipulator model and control saturation function.}
    \label{fig:robot}
\end{figure}

% ------------------------------------
% Robot Model Parameters
% -----------------------------------
\begin{table}[t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Two-link manipulator parameters of $p\textsuperscript{th}$ link.}
    \centering
    \begin{tabular}{c m{9.5em} c c c }
    % \begin{tabular}{c c c c c}
    \hline
    \textbf{Symbol} & \textbf{Description} & \textbf{Link 1} & \textbf{Link 2} \\
    \hline
    \hline 
    $m_p$ & Mass & \multicolumn{2}{c}{2.465 (kg)} \\
    \hline
    $l_p$  & Length & \multicolumn{2}{c}{0.2 (m)} \\  
    \hline
    ${l_c}_p$ & Center of mass & \multicolumn{2}{c}{0.139 (m)} \\
    \hline
    % $b_p$   & Viscous friction coef. & \multicolumn{2}{c}{0.1 (Nms)} \\
    % \hline
    % ${f_c}_p$  & Coulomb friction coef. & \multicolumn{2}{c}{0.24 (Nm)} \\
    % \hline
    % ${f_s}_p$  & Static friction coef. &  0.3 (Nm) & 0.2 (Nm) \\
    % \hline
    $I_p$  & Inertia & \multicolumn{2}{c} {0.069 (kgm\textsuperscript{2})} \\
    \hline
    \end{tabular}
    \label{table:system:params}
\end{table}

% \begin{table}[t]
%     \renewcommand{\arraystretch}{1.3}
%     \caption{Two-link manipulator parameters of $p\textsuperscript{th}$ link.}
%     \centering
%     \begin{tabular}{c m{9.5em} c c c }
%     % \begin{tabular}{c c c c c}
%     \hline
%     \textbf{Symbol} & \textbf{Description} & \textbf{Link 1} & \textbf{Link 2} \\
%     \hline
%     \hline 
%     $m_p$ & Mass & 2.465 (kg) & 2.465 (kg) \\
%     \hline
%     $l_p$  & Length & 0.2 (m) & 0.18 (m) \\
%     \hline
%     ${l_c}_p$ & Center of mass & 0.139 (m) & 0.087 (m) \\
%     \hline
%     $b_p$   & Viscous friction coef. &  0.1 (Nms) & 0.1 (Nms) \\
%     \hline
%     ${f_c}_p$  & Coulomb friction coef. &  0.24 (Nm) & 0.24 (Nm) \\
%     \hline
%     ${f_s}_p$  & Static friction coef. &  0.3 (Nm) & 0.2 (Nm) \\
%     \hline
%     $I_p$  & Inertia & 0.069 (kgm\textsuperscript{2}) & 0.015 (kgm\textsuperscript{2}) \\
%     \hline
%     \end{tabular}
%     \label{table:system:params}
% \end{table}

% ------------------------------------
% Reference Trajectory
% ------------------------------------
The desired trajectory $\qd=\left({q_d}_1,{q_d}_2\right)^\top $ was generated using quintic (5$\textsuperscript{th}$-order) polynomial functions $\mathcal{P}^5(t;\mv{x}_1,\mv{x}_2,t_f)$, which satisfy the initial and final joint angles $\mv{x}_1$ and $\mv{x}_2$, as well as zero initial and final joint velocities. 
Here, $t_f$ denotes the duration of each segment.
The desired (reference) trajectory is defined as follows:
\begin{equation}
    \qd(t) 
    =
    \begin{pmatrix}
        {q_d}_1
        \\
        {q_d}_2
    \end{pmatrix}
    =
    \begin{cases}
    \mathcal{P}^5(t;\mv{q}_d^{(0)},\mv{q}_d^{(1)},t_f),
        &
        \text{if } 0 \le t < t_f 
        ,
        \\
    \mathcal{P}^5(t;\mv{q}_d^{(1)},\mv{q}_d^{(2)},t_f),
        &
        \text{if } t_f \le t < 2t_f 
        ,
        \\
    \mathcal{P}^5(t;\mv{q}_d^{(2)},\mv{q}_d^{(1)},t_f),
        & 
        \text{if } 2t_f \le t < 3t_f 
        ,
        \\
    \mathcal{P}^5(t;\mv{q}_d^{(1)},\mv{q}_d^{(0)},t_f),
        &    
        \text{if } 3t_f \le t < 4t_f 
        ,
        \\
    \end{cases}
    \label{eq:desired trajectory}
\end{equation}
where 
$
    \mv{q}_d^{(0)}:=\left(-\tfrac{1}{3}\pi,\tfrac{1}{3}\pi\right)^\top
$, 
$
    \mv{q}_d^{(1)}:=\left(\tfrac{1}{4}\pi,-\tfrac{1}{2}\pi\right)^\top
$, 
$
    \mv{q}_d^{(2)}:=\left(-\tfrac{1}{4}\pi,\tfrac{1}{4}\pi\right)^\top
$, 
and $t_f=3$ s, and the reference trajectory is illustrated in Fig.~\ref{fig:robot:model}.
To demonstrate the learning capability of the DNN, the desired trajectory was applied twice. 
The first and second repetitions of the reference trajectory are referred to as Episode $1$ and Episode $2$, respectively.

However, since the DNN's weights are initialized randomly, large reference trajectories should be avoided to mitigate the effect of initial poor performance of the controller.
Hence, in advance of the reference trajectory, the manipulator was initially positioned at ${\qd}^{\rm{origin}}=\left(-\tfrac{1}{2}\pi,0\right)^\top$ (\ie stable equilibrium point).
Then, a reference trajectory was applied to the manipulator to move from ${\qd}^{\rm{origin}}$ to ${\qd}_0$ in 3 seconds, \ie $\mathcal{P}^5(t;{\qd}^{\rm{origin}},{\qd}_0,3)$, which is the initial and final points of the episode.
After the manipulator reached ${\qd}_0$, the constant reference trajectory $\qd={\qd}_0$ was applied for 8 seconds to be steady and to avoid the effect of the transient response.

% ------------------------------------
% Control Input Saturation
% ------------------------------------
The effectiveness of the proposed CONAC to handle the combination of the convex input constraints was demonstrated by defining the saturation function which projects the control input onto following convex set $\Theta_{\mysat}:=\left\{\Theta_{\mysat, {\rbu}}\cap\Theta_{\mysat, {\tau_2}}\right\}$ as depicted in Fig.~\ref{fig:robot:sat}, where
\begin{subequations}
    \begin{align}
        \Theta_{\mysat, {\rbu}}
        :=&
        \left\{
            \rbu
            \mid
            \norm{\rbu} \le 
            11
            % \overline\tau
        \right\}
        \label{eq:sat:norm}
        ,
        \\
        \Theta_{\mysat, {\tau_2}}
        :=&
        \left\{
            \tau_2
            \mid
            \vert{\tau_2}\vert \le 
            3.5
            % \overline\tau_2
        \right\}
        .
        \label{eq:sat:tau2}
    \end{align}
    \label{eq:sat}% %%
\end{subequations}
These sets \eqref{eq:sat:norm} and \eqref{eq:sat:tau2} can be physically interpreted as the voltage source limitation of the actuators and the torque limitation of the second link actuator, respectively.

\hfill

% ------------------------------------
% Comparative Study
% ------------------------------------
For a comparative study, we implemented 4 controllers.
The properties of the controllers are summarized in Table~\ref{table:controllers}.

% ------------------------------------
% Controllers: CONAC 
% ------------------------------------
\subsubsection*{Controller 1: \normalfont{(C$_1$)} \textit{CONAC with large $\beta_j$}}

To handle the input saturation, the proposed CONAC was implemented with the input bound constraint \eqref{eq:cstr:input:bound} $c_{\overline\tau_2}$ and $c_{\underline\tau_2}$ for second link, and the nonlinear input norm constraint $c_{\rbu}$ \eqref{eq:cstr:input:norm}, as follows:
\begin{equation}
    c_{\overline\tau_2}     
    =
    \tau_2-\overline\tau_2
    % \le 0
    ,
    \;
    c_{\underline\tau_2} 
    =
    \underline\tau_2-\tau_2
    % \le 0
    ,
    \;
    c_{\rbu}
    =
    \tfrac{1}{2}
    \left(
        \norm{\rbu}-\overline\tau
    \right)^2 
    % <0
    ,
    \label{eq:sim:cstr:CONAC}
\end{equation}
where $\overline\tau_2=-\underline{\tau}_2=3.5$ are the absolute bound of the second link actuator's torque, and $\overline\tau=11$ is the maximum feasible norm of the first and second torques.
Hence, (C$_1$) is possible to use full capability of the actuator.
The weight constraints $c_{\theta_i},\ \forall i\in\{0,\cdots,k\}$ were also considered in the CONAC.
As described in \eqref{eq:control:est}, the control input (torques) is equal to the output of the DNN, and the adaptation law is defined in \eqref{eq:adaptation law}.

The update rates of the Lagrange multipliers for weight constraints were set to $\beta_{\theta_i}=10^{-3},\ \forall i\in\{0,\cdots,k\}$, and ,for control input constraints, the update rates were set to large values such that $\beta_{\rbu}=10$ and $\beta_{\overline{\tau}_2}=\beta_{\underline{\tau}_2}=10^{3}$ to investigate the effect of the large values of update rates.

% ------------------------------------
% Controllers: CONAC with small beta
% ------------------------------------
\subsubsection*{Controller 2: \normalfont{(C$_2$)} \textit{CONAC with small $\beta_j$}}

As the second controller, the proposed CONAC was implemented with the same input bound constraint \eqref{eq:sim:cstr:CONAC} as (C$_1$), but with small update rate values such that $\beta_{\rbu}=10$ and $\beta_{\overline{\tau}_2}=\beta_{\underline{\tau}_2}=10^{3}$.
The remaining properties and parameters were the same as (C$_1$).

% ------------------------------------
% Controllers: CONAC without ctrl cstr
% ------------------------------------
\subsubsection*{Controller 3: \normalfont{(C$_3$)} \textit{CONAC without control input constraints}}

The third controller was the same as (C$_1$) and (C$_2$), but without the control input constraints, to emphasize the control input constraint handling capability of the proposed CONAC.
In this case, the update rate of the Lagrange multipliers regarding the control input constraints, \ie $\beta_{\rbu},\beta_{\overline{\tau}_2},\beta_{\underline{\tau}_2}$, was set to $0$, so that $[\lambda_j]_{j\in\{\rbu,\overline{\tau}_2,\underline{\tau}_2\}}$ are not updated (see, \eqref{eq:adap:lbd}).
The remaining properties and parameters were the same as (C$_1$) and (C$_2$).

% ------------------------------------
% Controllers: Aux. system in CONAC-AUX
% ------------------------------------
\subsubsection*{Controller 4: \normalfont{(C$_4$)} \textit{CONAC with auxiliary system instead of control input constraints}}

On the other hand, the existing auxiliary system \cite{Esfandiari:2014aa,Karason:1994aa,Esfandiari:2015aa} was used to compare with the proposed CONAC with control input constraints by replacing them with the auxiliary system.
The auxiliary system is defined as $\ddtt{\mv\zeta} = \mm{A}_\zeta \mv{\zeta} + \mm{B}_\zeta \Delta\rbu,\ \mv{\zeta}\vert_{t=0} = \mv{0}_{2\times1}$, where $\mv{\zeta}\in\R^n$ denotes the auxiliary state variables, and $\mm{A}_\zeta\in\R_{<0}^{n\times n}$, and $\mm{B}_\zeta\in\R^{n\times n}$ are user-designed matrices.
The auxiliary state $\mv{\zeta}$ is affected by the overflow amount of the control input $\Delta\rbu$, whose entries are defined as $\Delta\tau_{i} = 
\tau_{i}-\max(\underline\tau_i,\min(\overline\tau_i,\tau_i)),\ \forall i\in\{1,2\}$.
The auxiliary state are used in the adaptation law \eqref{eq:adap:th:approx} by substituting $\fe$ with $\fe+\mv{\zeta}$.
Consequently, the weights are adapted to reduce not only the tracking error $\fe$ but also the auxiliary state $\mv{\zeta}$, minimizing the control input constraint saturation.

The matrices of the auxiliary system were set to $\mm{A}_\zeta=\mydiag([-10,-10])$ and $\mm{B}_\zeta=\mydiag([10^3,10^3])$, considering the tradeoff between the sensitivity of the control saturation, and the convergence speed of the auxiliary system.
Since, the auxiliary system only considers the saturation of each control input, \ie which is same as the input bound constraint \eqref{eq:cstr:input:bound}, $\overline{\tau}_1$ was set to $10.428$ to approximately match the maximum allowable norm of the control input $\overline\tau$, with respect to $\overline{\tau}_2=3.5$ and $\overline{\tau}=11$, \ie $\overline{\tau}_1^2+\overline{\tau}_2^2=\overline{\tau}^2$).
For simplicity, let say (C$_4$) was imposed with input bound constraint for the first link actuator, \ie $c_{\overline{\tau}_1}$ and $c_{\underline{\tau}_1}$.
\MSRY{사실, 제약조건이 부여된 것은 아니기 떄문에 위와 같이 표현하였음}

\hfill

% ------------------------------------
% Controller Parameters
% ------------------------------------
For all controllers ((C$_1$), (C$_2$), (C$_3$), and (C$_4$)), the DNN input vector ${\q}_n$ was set as ${\q}_n=\left(\q^\top,\dq^\top,\fe^\top,1\right)^\top\in\R^{3n+1}$ with the augmented scalar $1$ included to account for the bias term in the weight matrix. 
The DNNs' weights were initialized randomly with uniform distribution in the range of $[-0.1,0.1]$, and the same random seed.
All DNN architectures have two hidden layers with four nodes, \ie $k=2$, and $l_0=6, l_1=4, l_2=4$, and $l_3=2$, and the adaptation gain was set to $\alpha=0.5$.
The filtering matrix $\mm{\Lambda}$ was set to $\mm{\Lambda}=\mydiag([5,15])$.
The sampling rate of the controller was $250$ Hz considering the real-time implementation.
The weight constraint parameters were selected as $\overline\wth_0=\overline\wth_1=\overline\wth_2=6$.
\MSRY{어차피 내부는 안걸리니까 그림 더 깔끔하게 다 6이라고 함} 
The control results were collected via CAN communication with PCAN USB  and recorded in a computer with a sampling rate of $250$ Hz.

The quantitative comparison of the tracking performance was evaluated using the $L_2$-norm of them, (\ie $\sqrt{\int_0^T\norm{h}\,\der t}$, where $h\in\{e_1,e_2\}$ and $T\in\R_{>0}$ denotes the each episode duration time.

% ------------------------------------
% Controller Table
% ------------------------------------
\begin{table}[t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Properties of the controllers}
    \label{table:controller}
    \centering
    \begin{tabular}{c l l}
    \hline
        &\bf{Description}&\bf{Input Constraint}\\
    \hline
    \hline
        (C$_1$)&large $\beta_j$&$\overline{\tau}=11,\overline{\tau}_2=-\underline{\tau}_2=3.5$ \\
    \hline
        (C$_2$)&small $\beta_j$&$\overline{\tau}=10,\overline{\tau}_2=-\underline{\tau_2}=3.5$ \\
    \hline
        (C$_3$)&no $\rbu$ constraints&$[\beta_j]_{j\in\{\rbu,\overline{\tau}_2,\underline{\tau}_2\}}=\mv{0}$ \\
    \hline
        (C$_4$)&aux. sys.&$\overline{\tau}_1=-\underline{\tau}_1=10.428,\overline{\tau_2}=-\underline{\tau_2}=3.5$\\
    \hline
    \end{tabular}
    \label{table:controllers}
\end{table}

\begin{figure*}[t]
  \centering
    \subfloat[Tracking result of $q_1$.]{
    %   \includegraphics
      \includegraphics[width=\figSizeTwoCol\linewidth]
      {
        % fig/Fig1.pdf
        src/measurement/figures/compare/Fig1.eps
      }%
      \label{fig:ctrl:result:q1}}
    \hfill
    \subfloat[Tracking result of $q_2$.]{
      \includegraphics[width=\figSizeTwoCol\linewidth]
      {
        % src/measurement/figures/compare/Fig11.eps
        src/measurement/figures/compare/Fig2.eps
      }%
      \label{fig:ctrl:result:q2}}
    \vfill
      \subfloat[Control input $\tau_1$.]{
        \includegraphics[width=\figSizeTwoCol\linewidth]
        {
            src/measurement/figures/compare/Fig3.eps
        }%
        \label{fig:ctrl:result:tau:1}}
    \hfill
      \subfloat[Control input $\tau_2$.]{
        \includegraphics[width=\figSizeTwoCol\linewidth]
        {
            src/measurement/figures/compare/Fig4.eps
        }%
        \label{fig:ctrl:result:tau:2}}
        \vfill
    \subfloat[Norm of control input $\rbu$.]{
        \includegraphics[width=\figSizeTwoCol\linewidth]
        {
            src/measurement/figures/compare/Fig5.eps
        }%
        \label{fig:ctrl:result:tau:norm}}
    \hfill
    \subfloat[Norm of weights ${\estwth}_i$.]{
        % \includegraphics
        \includegraphics[width=\figSizeTwoCol\linewidth]
        {
            % fig/Fig7.pdf
            src/measurement/figures/compare/Fig7.eps
        }%
        \label{fig:ctrl:result:weight}}        
    \caption{
    Real-time implementation results of (C$_1$) \protect\colorLine{blue}{solid}, (C$_2$) \protect\colorLine{cyan}{solid}, (C$_3$) \protect\colorLine{gray}{solid}, (C$_4$) \protect\colorLine{magenta}{solid}, and reference signal $\qd$ \protect\colorLine{red}{dashed}. 
  }
\label{fig:ctrl:result}
\end{figure*}

% ------------------------------------
% Norm of error in Episode 1: 
% C1 e1 ep1: 25.013
% C1 e2 ep1: 22.651
% C2 e1 ep1: 21.719
% C2 e2 ep1: 10.530
% C3 e1 ep1: 23.168
% C3 e2 ep1: 13.062
% C4 e1 ep1: 24.195
% C4 e2 ep1: 12.927
% Norm of error in Episode 2: 
% C1 e1 ep2: 7.513
% C1 e2 ep2: 2.617
% C2 e1 ep2: 8.021
% C2 e2 ep2: 2.435
% C3 e1 ep2: 8.408
% C3 e2 ep2: 3.553
% C4 e1 ep2: 19.721
% C4 e2 ep2: 4.256
% Improvement in Episode 2: 
% C1 e1: 0.700
% C1 e2: 0.884
% C2 e1: 0.631
% C2 e2: 0.769
% C3 e1: 0.637
% C3 e2: 0.728
% C4 e1: 0.185
% C4 e2: 0.671
% ------------------------------------
\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Quantitative comparison of performances' $L_2$ norm.}
    \centering
    \begin{tabular}{c c c c c}
    \hline
		& \multicolumn{4}{c}{\textbf{Episode 1}} \\
    \hline
	\hline 
		& (C$_1$) & (C$_2$) & (C$_3$) & (C$_4$) \\
	\hline
		$e_1\times10^{3}$ / rad & $25.013$ & $21.719$ & $23.168$ & $24.195$ \\
	\hline
        $e_2\times10^{3}$ / rad & $22.651$ & $10.530$ & $13.062$ & $12.927$ \\
	\hline
        & \multicolumn{4}{c}{\textbf{Episode 2}} \\
    \hline
    \hline
        & (C$_1$) & (C$_2$) & (C$_3$) & (C$_4$) \\
	\hline
		$e_1\times10^{3}$ / rad & $7.513$ & $8.021$ & $8.408$ & $19.721$ \\
	\hline
        $e_2\times10^{3}$ / rad & $2.617$ & $2.435$ & $3.553$ & $4.256$ \\
    \hline
    \end{tabular}
    \label{tab:sim:L2}
\end{table}

\subsection{Validation Results}

% ------------------------------------
% Tracking Performance
% ------------------------------------
\subsubsection{Tracking Performance}

The real-time implementation results are shown in Fig.~\ref{fig:ctrl:result} and the quantitative comparison of the tracking performance is given in Table~\ref{tab:sim:L2}.
Across the two episodes, all controllers enhanced their tracking performances about $70 \%$ for (C$_1$), $63.1 \%$ for (C$_2$), $63.7 \%$ for (C$_3$), and $18.5 \%$ for (C$_4$) in $L_2$ norm of $e_1$, and $88.4 \%$ for (C$_1$), $76.9 \%$ for (C$_2$), $72.8 \%$ for (C$_3$), and $67.1 \%$ for (C$_4$) in $L_2$ norm of $e_2$, respectively, as shown in Table~\ref{tab:sim:L2}.
Qualitatively, as described in Fig.~\ref{fig:ctrl:result:q1} and Fig.~\ref{fig:ctrl:result:q2}, the oscillations were suppressed in the second episode.
This results lead to the conclusion that the DNNs in all controllers were able to learn the ideal control input $\rbu^*$ properly, by the filtered error $\fe$, despite the fact that $\rbu^*$ was unknown in advance.
Moreover, since the learning began from the random initialization of the weights and no prior knowledge of the system, the online learning capability of the proposed CONAC was demonstrated.

In addition, it is notable that in both episodes, all controllers failed to track the desired trajectory of $q_1$ in time intervals from $17.5$ s to $19.9$ for Episode $1$, and from $23.5$ s to $25.9$ s and from $29.5$ s to $31.9$ s for Episode $2$.
This is because the control inputs were saturated in the time intervals, as shown in Fig.~\ref{fig:ctrl:result:tau:1} and Fig.~\ref{fig:ctrl:result:tau:norm}.

The tracking performances of CONAC with control input constraints, \ie (C$_1$), and (C$_2$), outperformed CONAC with the auxiliary system (C$_4$) in both episodes, as shown in Table~\ref{tab:sim:L2}.
This is because, (as wil be discussed in Section \ref{sec:sim:constraint}) (C$_1$), and (C$_2$) were able to use the full capability of the actuator, while (C$_4$) was limited its first link actuator due to the feature of the auxiliary system.
The tracking performance of (C$_3$) was also better than (C$_4$), but worse than (C$_1$) and (C$_2$), since the control input constraints were not considered in (C$_3$).
Without consideration of the control input saturation, DDN became confused by the difference between its output (control input) and the saturated control input, which leads to the poor performance of (C$_3$).

% ------------------------------------
% Control Input Saturation
% ------------------------------------
\begin{figure}[t]
    \centering
    \subfloat[Control input $\tau$ in time interval from $29.4$ s to $31.9$ s (bird's eye view).]{
    % \subfloat[Auxiliary states $\mv{\zeta}$ of CONAC-AUX.]{
        \includegraphics[width=\figSizeOneCol\linewidth]
        {
            src/measurement/figures/compare/Fig6.eps
        }%
        \label{fig:ctrl:result:scope:control}}
        \vfill
    \subfloat[Lagrange multipliers $\lambda_j$  in time interval from $29.4$ s to $31.9$ s.]{
        \includegraphics[width=\figSizeOneCol\linewidth]
        {
            src/measurement/figures/compare/Fig8.eps
        }%
        \label{fig:ctrl:result:scope:beta}}
        \vfill
    \subfloat[Auxiliary states in time interval from $29.4$ s to $31.9$ s.]{
        \includegraphics[width=\figSizeOneCol\linewidth]
        {
            src/measurement/figures/compare/Fig9.eps
        }%
        \label{fig:ctrl:result:scope:aux}}
  \caption{
    Constraint handling process of (C$_1$) \protect\colorLine{blue}{solid}, (C$_2$) \protect\colorLine{cyan}{solid}, (C$_3$) \protect\colorLine{gray}{solid}, and (C$_4$) \protect\colorLine{magenta}{solid} in time interval from $29.4$ s to $31.9$ s.
  }
\label{fig:ctrl:result:scope}
\end{figure}

\begin{figure}[!t]
	\centering
	\includegraphics[width=0.6\linewidth]{
		src/figures/lambda_effect.drawio.pdf
		}
	\caption{
		The effect of Lagrange multiplier $\lambda_{j}$ on the adaptation direction of $\widehat{ \mv{\theta}}_1$.
		The notation ($\cdot'$) and ($\cdot''$) represent two different cases of bigger and smaller Lagrange multiplier, respectively (i.e., $\lambda_{j}' > \lambda_{j}''$).
	}
	% \vspace{1mm}
	\label{fig:lambda_effect}
\end{figure}

\hfill
\subsubsection{Constraint Handling} \label{sec:sim:constraint}

In this section, the constraint handling capability of all controllers are investigated.

As shown in Fig.~\ref{fig:ctrl:result:tau:1}, Fig.~\ref{fig:ctrl:result:tau:2}, and Fig.~\ref{fig:ctrl:result:tau:norm}, all controllers except (C$_3$) which did not consider the control input saturation, satisfied the control input saturation illustrated in Fig.~\ref{fig:robot:sat}, with $c_{\rbu},c_{\overline\tau_2}$ and $c_{\underline\tau_2}$ for (C$_1$), and (C$_2$), and $c_{\overline\tau_1},c_{\underline\tau_1},c_{\overline\tau_2}$, and $c_{\underline\tau_2}$ for (C$_4$), as the weights were adapted to reduce the constraint violations, using Lagrange multipliers $\lambda_j,\ \forall j\in\{\rbu,\overline{\tau}_2,\underline{\tau}_2\}$, and auxiliary state $\mv{\zeta}$, respectively.
It is worth to note that the control input constraints were violated repeatedly, since the controllers were implemented in discrete time with a low sampling rate of $250$ Hz.

Even though all controllers except (C$_3$) satisfied the input saturation, (C$_4$) was not able to use the full capability of the actuator, as shown in Fig.~\ref{fig:ctrl:result:scope:control} which shows the bird-eye view of the control input $\rbu$ in time interval from $29.4$ s to $31.9$ s.
This is because, as aforementioned, the auxiliary system only considers saturation of each control input, and the maximum value of the first link's torque $\overline{\tau}_1$ was set to approximately match the maximum allowable norm of the control input $\overline\tau$, instead of $c_{\rbu}$.
In contrast, (C$_1$), and (C$_2$) were able to use the full capability of the actuator with the combination of the input bound constraints $c_{\overline\tau_2}$ and $c_{\underline\tau_2}$, and the nonlinear input norm constraint $c_{\rbu}$ which is completely match the feasible domain of the control input $\Theta_{\mysat}$.

% zeta
The detail of the constraint handling in the time interval from $29.4$ s to $31.9$ s, is now investigated.
First, for (C$_4$), the auxiliary state $\zeta_1$ generated by constraint violation of $c_{\overline\tau_1}$. 
As the weights were adapted to reduce not only $\fe$ but also $\mv{\zeta}$, the constraint violation of $c_{\overline\tau_1}$ was reduced, as shown in Fig.~\ref{fig:ctrl:result:scope:control}.
After the constraint was satisfied, the auxiliary state $\zeta_1$ converged to zero, by the stable state matrix $\mm{A}_\zeta$ in the auxiliary system, \ie $\ddtt\mv{\zeta} = \mm{A}_\zeta \mv{\zeta} + \mm{B}_\zeta \Delta\rbu$ with $\Delta\rbu=\mv{0}_{2\times1}$.

% beta + zeta
For the cases of (C$_1$) and (C$_2$), the Lagrange multipliers $\lambda_{\rbu}$ adjusted the adaptation direction of the weights $\estwth_1$ to reduce the constraint violation $c_{\rbu}$, as $\lambda_{\rbu}$ increased by the constraint violations.
% comparison; big, small beta
For more details, the effect of the update rate $\beta_j,\ \forall j\in\mathcal I$ of Lagrange multipliers is shown in Fig.~\ref{fig:lambda_effect} for convex constraint $c_j,\ \forall j\in\mathcal I$ and two different cases of $\lambda_j$.
With the bigger $\lambda_j$, the adaptation of $\estwth_1$ is more assertive than that with smaller $\lambda_j$.
This leads to a faster satisfaction of the constraint $c_j,\ \forall j\in\mathcal I$, but this also causes oscillation in the adaptation process.
In Fig.~\ref{fig:ctrl:result:scope:control} and Fig.~\ref{fig:ctrl:result:scope:beta}, $\lambda_{\rbu}$ of (C$_1$) repeatedly increased and decreased, which indicates that the constraint $c_{\rbu}$ was satisfied and violated, respectively, in the discrete time implementation.
Besides, with small $\lambda_{\rbu}$, the adaptation direction of $\estwth_1$ is adjusted more conservatively, which leads to a slower satisfaction of the constraint $c_{\rbu}$, as shown in Fig.~\ref{fig:ctrl:result:scope:control} and the positively existence of $\lambda_{\rbu}$ in Fig.~\ref{fig:ctrl:result:scope:beta}.
This allows the controller to show a smoother control input.
However, since the weights were adapted to reduce the filtered error $\fe$ while the control input was saturated, the weights were not able to effectively adapt the weights with respect to the filtered error $\fe$.

\hfill

Finally, the weight norm constraints $c_{\theta_i},\ \forall i\in\{0,\cdots,2\}$ are investigated.
The weight norm constraints were not activated except for the outer layer weights $c_{\theta_2}$ of (C$_3$), as shown in Fig.~\ref{fig:ctrl:result:weight}.
This is because the weights of (C$_1$), (C$_2$), and (C$_4$) were suppressed by the control input constraints, while (C$_3$) was not.
\MSRY{필요하다면 왜 제어입력 제약하는게 가중치 제약도 되는지 추가 가능}
In addition, the inner and hidden layer weights constraints were not violated (\ie $\norm{\estwth_0},\norm{\estwth_1}$ were uniformly lower than $\norm{\estwth_2}$), since the backpropagated error was multiplied by the gradient of the activation function, which is limited from $-1$ to $1$, \ie see, \eqref{eq:gradient:phi}.

The details of the constraint handling process are shown in Fig.~\ref{fig:ctrl:result:scope}, as well.
In Fig.~\ref{fig:ctrl:result:scope:beta}, the Lagrange multipliers $\lambda_{\theta_2}$ were increased by the constraint violations, and the adaptation direction of the weights $\estwth_2$ was adjusted to reduce the constraint violations.

% ------------------------------------
% Karsh-Kuhn-Tucker (KKT) conditions}
% ------------------------------------
\hfill
\subsubsection{Satisfaction of Karsh-Kuhn-Tucker (KKT) conditions}

\MSRY{이 부분도 IECON과 마찬가지로 이산으로 돌아감으로 바로 보여주기는 힘들 수 있을 것 같습니다. 실제로 보시면 많은 진동이 보여서 정상상태의 KKT를 주장할 수 있을지 의문입니다.}

Since the proposed CONAC was implemented in discrete time, demonstration regarding the KKT conditions in steady state is not straightforward.
However, as the weights and Lagrange multipliers were almost constant when the constraints were violated, the satisfaction of the KKT conditions can be implicitly shown, since the Lagrange multipliers of active constraints were positive and inactive constraints were zero.

% ------------------------------------
% Computational Time
% ------------------------------------
\hfill
\subsubsection{Effectiveness of CONAC in Real-time Applications}

The effectiveness of the proposed CONAC in real-time applications was evaluated by measuring the computational time measured using CPU clock ticks of the OpenCR1.0 board.
The computational times are shown in Fig.~\ref{fig:ctrl:real:result:cmp:time} indicates that the proposed CONAC can be implemented in real-time applications with a sampling rate of $250$ Hz.
In addition, it is notable that the computational time decreased as the weights converged.
\MSRY{삭제해야할수있음; C4는 그러하지 않기 때문}

\begin{figure}[t]
    \centering
        \includegraphics[width=\figSizeOneCol\linewidth]
        {
            src/measurement/figures/compare/Fig10.eps
        }%
    \caption{
        Computational time of (C$_1$) \protect\colorLine{blue}{solid}, (C$_2$) \protect\colorLine{cyan}{solid}, (C$_3$) \protect\colorLine{gray}{solid}, and (C$_4$) \protect\colorLine{magenta}{solid}.
    }
    \label{fig:ctrl:real:result:cmp:time}
  \end{figure}

%  SECTION CONCLUSION ======================================
\section{Conclusion}\label{sec:conclusion}

This paper proposed a constrained optimization-based neuro-adaptive controller (CONAC) for uncertain Euler-Lagrange systems, addressing both weight norm and input constraints through a rigorous optimization framework. 
The stability of the proposed controller was analyzed via Lyapunov theory, demonstrating bounded tracking and estimation errors under real-time adaptation.

The controller effectively incorporated both convex input constraints and weight norm constraints, ensuring that actuator limitations and neural network parameters\MSRY{Weight?} remained within predefined bounds. 
By embedding these constraints into the optimization process, CONAC ensured convergence of the weights in accordance with the Karush-Kuhn-Tucker (KKT) conditions, thereby achieving both stability and optimality.

A real-time implementation validated the superior performance of CONAC compared to conventional methods. 
The proposed controller successfully handled complex input constraints while rigorously managing weight norms, resulting in improved tracking accuracy and stable performance without significant oscillations.

Future work may extend this approach to include constraints on both control inputs and system states, further enhancing the flexibility and robustness of neuro-adaptive control systems within constrained optimization frameworks.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
\appendix

\subsection{Input Constraint Candidates}\label{sec:appen:cstr}

This section introduces potential weight and input constraints that can be used in the proposed CONAC.

\subsubsection{Input Bound Constraint}\label{sec:appen:cstr:input:bound}

Most physical systems have control input limits due to electrical and mechanical limitations. These are expressed as $\mv{c}_{\overline \tau}:= [c_{\overline \tau_i}]_{i\in[1,\cdots,n]}$ and $\mv{c}_{\underline\tau}:= [c_{\underline\tau_i}]_{i\in[1,\cdots,n]}$, where
\begin{equation}
    \begin{aligned}
        c_{\overline \tau_i}=\tau_{(i)} - {\tau_{\overline \tau_i}}
        ,
        \quad
        c_{\underline\tau_i}={\tau_{\underline\tau_i}}-\tau_{(i)}
        ,
    \end{aligned}
    \label{eq:cstr:input:bound}
\end{equation}
with $\tau_{\overline \tau_i}\in\R_{>0}$ and $\tau_{\underline\tau_i}\in\R_{>0}$ representing the maximum and minimum control input bounds, respectively.
% The corresponding Lagrange multipliers are $\lambda_{u_j,i},\ \forall j\in[M,m]$. 
The gradients of $\mv{c}_{\overline \tau}$ and $\mv{c}_{\underline\tau}$ with respect to $\estwth$ are given by
\begin{equation}
    \begin{aligned}
        \pptfrac{\mv c_{\overline \tau}}{\estwth}
        =& 
        \left[
            \pptfrac{c_{\overline \tau_i}}{\estwth}^\top
        \right]_{i\in\{1,\cdots,n\}}
        % \begin{bmatrix}
        %     \pptfrac{c_{\overline \tau_1}}{\estwth}^\top \\
        %     \vdots \\
        %     \pptfrac{c_{\overline \tau_n}}{\estwth}^\top
        % \end{bmatrix}
        = 
            +\pptfrac{\estNN}{\estwth}
            \\
        =&
        =
        +
        \begin{bmatrix}
            (\mm{I}_{l_{k+1}}\otimes \estact_{k}^\top)&
            \!\cdots\! &
            (
                \cdot
            )
        \end{bmatrix} 
        \in
        \mathbb R^{n\times \Xi}
        , 
        \\
        \pptfrac{\mv c_{\underline\tau}}{\estwth}         
        =
        & 
        \left[
            \pptfrac{c_{\underline\tau_i}}{\estwth}^\top
        \right]_{i\in\{1,\cdots,n\}}
        % \begin{bmatrix}
        %     \pptfrac{c_{\underline\tau_1}}{\estwth}^\top \\
        %     \vdots \\
        %     \pptfrac{c_{\underline\tau_n}}{\estwth}^\top
        % \end{bmatrix}
        = 
        -\pptfrac{\estNN}{\estwth}
        \\
        =&
        =
        -\begin{bmatrix}
            (\mm{I}_{l_{k+1}}\otimes \estact_{k}^\top)&
            % V_k^\top \phi_{k}' (I_{l_{k}}\otimes  \phi_{k-1}^\top)&
            \!\cdots\! &
            (
                \cdot
            )
        \end{bmatrix} 
        \in
        \mathbb R^{n\times \Xi}
        .
    \end{aligned}
    \label{eq:cstr:input:bound:grad}
\end{equation}

\subsubsection{Input Norm Constraint}\label{sec:appen:cstr:input:norm}

Consider the control input $\rbu$ as the torque of each actuator corresponding to its generalized coordinate. Since torque is typically linearly proportional to current, actuators that share a common power source are often subject to total current limitations. This can be captured by the following inequality constraint: 
\begin{equation}
    c_{\rbu}
    =
    \norm{
        \mv{\tau}
    }^2 
    -
    \overline\tau^2
    ,
    \label{eq:cstr:input:norm}
\end{equation}
with $\overline\tau\in\R_{>0}$ denoting the maximum allowable control input magnitude. This input norm constraint is also commonly applied in current and torque control problems for electric motors \cite{Choi:2024aa}.
% The corresponding Lagrangian multiplier is $\lambda_{u_b}$. 
The gradients of $c_{\rbu}$ with respect to $\estwth$ are given by
\begin{equation}
    \pptfrac{\mv c_{\rbu}}{\estwth}
    \!=\! 
    \textstyle\sum_{i=1}^n 2\tau_{i} 
    \left(
        \myrow_i
        \left(
            \pptfrac{\estNN}{\estwth}
        \right)
    \right)^\top  
    \!=\! 
    {\pptfrac{\estNN}{\estwth}}^\top
    \mv{\tau}
    % \tau^\top (I_{l_{k+1}}\otimes \estNN_k^\top)
    \in \mathbb R^{\Xi}.
    \label{eq:cstr:input:norm:grad}
\end{equation}
It should be noted that constraints \eqref{eq:cstr:input:bound} and \eqref{eq:cstr:input:norm} can be imposed simultaneously, as their gradients \eqref{eq:cstr:input:bound:grad} and \eqref{eq:cstr:input:norm:grad} are linearly independent, satisfying Assumption \ref{assum:LICQ}, \ie the LICQ condition.

\printbibliography

% \begin{IEEEbiographynophoto}{Jane Doe}
%     Biography text here without a photo.
% \end{IEEEbiographynophoto}

\begin{IEEEbiography}[{
    \includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{\template/people/msRyu.jpg}}]{Myeongseok Ryu}
    received the B.S. degree in mechanical engineering from Incheon National University, South Korea, in 2023, and the M.S. degree in mechanical engineering from Gwangju Institute of Science and Technology (GIST), South Korea, in 2025. 
    He is currently serving as a researcher at the Cho Chun Shik Graduate School of Mobility, Korea Advanced Institute of Science and Technology (KAIST), South Korea.
    His research interests include adaptive control, neural-networks, and constrained optimization.
\end{IEEEbiography}

\begin{IEEEbiography}[{
    \includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{\template/people/dhHong.jpg}}]{Donghwa Hong}
    Donghwa Hong received his B.S. degree in Physics and Engineering Physics from Yonsei University, Wonju, South Korea, in 2023. He is currently pursuing the M.S. degree in Mechanical Engineering from GIST, Gwangju, South Korea.  
His research interests include robot control, neural-networks, and model identification. 
\end{IEEEbiography}

\begin{IEEEbiography}[{
    \includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{\template/people/khChoi.jpg}}]{Kyunghwan Choi}
    Kyunghwan Choi received his B.S., M.S., and Ph.D. degrees in Mechanical Engineering from KAIST, Daejeon, South Korea, in 2014, 2016, and 2020, respectively. Following his Ph.D., he worked at the Center for Eco-Friendly \& Smart Vehicles at KAIST as a Postdoctoral Fellow and later as a Research Assistant Professor. In 2022, he joined GIST as an Assistant Professor in the Department of Mechanical and Robotics Engineering. Currently, he serves as an Assistant Professor at the Cho Chun Shik Graduate School of Mobility at KAIST, where he is also the Director of the Mobility Intelligence and Control Laboratory. His research focuses on optimal and learning-based control for connected, automated, and electrified vehicles (CAEVs).
\end{IEEEbiography}

\end{document}