### 2025-06-07 manuscript.tex
improve with teacher

\documentclass[journal]{IEEEtran}

\def\pub{false} % true for publication, false for draft
\newcommand*{\template}{template} 
\input{\template/preamble/preamble_journal.tex}

\newcommand*{\figSizeTwoCol}{.49}
\newcommand*{\figSizeOneCol}{0.98}

% \addbibresource{\template/refs.bib}

\begin{document}

\title{
    Constrained Optimization-Based Neuro-Adaptive Control (CONAC) for Euler-Lagrange Systems Under Weight and Input Constraints
} %% Article title

\author{
  Myeongseok Ryu, Donghwa Hong, and Kyunghwan Choi
\thanks{
    Myeongseok Ryu and Kyunghwan Choi are with the Cho Chun Shik Graduate School of Mobility, Korean Advanced Institute of Science and Technology (KAIST), Daejeon 34051, Korea (e-mail: dding\_98@kaist.ac.kr and kh.choi@kaist.ac.kr).

    Donghwa Hong is with the Department of Mechanical and Robotics Engineering, Gwangju Institute of Science and Technology (GIST), Gwangju 61005, Korea (e-mail: fairytalef00@gm.gist.ac.kr).

    This work was supported in part by a Korea Research Institute for Defense Technology planning and advancement (KRIT) grant funded by Korea government DAPA (Defense Acquisition Program Administration) (No. KRIT-CT-22-087, Synthetic Battlefield Environment based Integrated Combat Training Platform) and in part by the National Research Foundation of Korea (NRF) grant funded by the Korea government (MSIT) (RS-2025- 00554087). (Corresponding author: Kyunghwan Choi)
    }}

\markboth{Journal of \LaTeX\ Class Files,~Vol.~18, No.~9, September~2020}%
{How to Use the IEEEtran \LaTeX \ Templates}

\maketitle

%% Abstract
\begin{abstract}
  This study presents a constrained optimization-based neuro-adaptive controller (CONAC) for uncertain Euler-Lagrange systems subject to weight norm and convex input constraints. 
  A deep neural network (DNN) is employed to approximate the ideal stabilizing control law while simultaneously learning the unknown system dynamics and addressing both types of constraints within a constrained optimization framework.
  The weight adaptation laws are formulated to solve the defined constrained optimization problem, ensuring satisfaction of first-order optimality conditions at steady state.
  The controller's stability is rigorously analyzed using Lyapunov theory, guaranteeing bounded tracking errors and bounded DNN weights. 
  The proposed controller is validated through a real-time implementation on a 2-DOF robotic manipulator, demonstrating its effectiveness in achieving accurate trajectory tracking while satisfying all imposed constraints.
\end{abstract}

\begin{IEEEkeywords}
    Neuro-adaptive control, constrained optimization, deep neural network, input constraint.
\end{IEEEkeywords}

\section*{Notation}
In this study, the following notation is used:

\begin{itemize}
    \item $\otimes$ denotes the Kronecker product \cite[Chap. 7 Def. 7.1.2]{Bernstein:2009aa}.
    \item A vector and a matrix are denoted by $\mv{x}=[x_i]_{i\in\{1,\cdots,n\}}\in\R^n$ and $
        \mm A
        := 
        [a_{ij}]
        _{
            i\in\{1,\cdots,n\},j\in\{1,\cdots ,m\}
        }\allowbreak\in\R^{n\times m}
        $, respectively.
    \item $\myrow_i(\mm A)$ denotes the $i\textsuperscript{th}$ row of the matrix $\mm A\in\R^{n\times m}$. 
    \item For $\mm{A}\in\R^{n\times m}$, $\myvec(\mm A):=(\myrow_1(\mm A^\top),\cdots,\myrow_m(\mm A^\top)  )^\top\in\R^{nm}$ denotes the vectorization of $\mm A$.
    % \item $\mycol_i(A)$ and $\myrow_j(A)$ denote the $i\textsuperscript{th}$ column and $j\textsuperscript{th}$ row of $A\in\R^{n\times m}$, respectively.
    % \item $\myvec(A):= [\mycol_1(A)^\top  ,\cdots,\mycol_m(A)^\top  ]^\top   $ for $A\in\R^{n\times m}$.
    \item $\lambda_{\min}(\mm A)$ denotes the minimum eigenvalue of the matrix $\mm A\in\R^{n\times n}$.
    \item $\mm I_n$ denotes the $n\times n$ identity matrix, and $\mm 0_{n\times m}$ denotes the $n\times m$ zero matrix.
\end{itemize}

%  SECTION INTRODUCTION ===================================
\section{Introduction}

\subsection{Background}

\IEEEPARstart{M}{any} engineering systems, including those in aerospace, robotics, and automotive applications, can be modeled using Euler-Lagrange systems. 
These systems are governed by dynamic equations derived from energy principles and describe the motion of mechanical systems with constraints. 
In practice, however, such systems often exhibit uncertainties due to unmodeled dynamics, parameter variations, or external disturbances. 
These uncertainties can significantly degrade control performance and, in some cases, lead to instability. 
To address these challenges, adaptive control methods have been widely employed to ensure robust performance in the presence of system uncertainties \cite{Ioannou:2006aa, Tao:2003aa}.

More recently, neuro-adaptive control approaches have been introduced to approximate unknown system dynamics or entire control laws using neural networks (NNs) \cite{Farrell:2006aa}. 
NNs are well-known for their universal approximation property, which allows them to approximate any smooth function over a compact set with minimal error. 
Various types of NNs have been utilized in neuro-adaptive control, including simpler architectures like single-hidden layer (SHL) NNs \cite{Ge:2010aa, Yesildirek:1995aa} and radial basis function (RBF) NNs \cite{Liu:2013ab,Ge:2002aa}, as well as more complex models like deep NNs (DNNs) \cite{Patil:2022aa} and their variations. 
SHL and RBF NNs are often employed to approximate uncertain system dynamics or controllers due to their simplicity \cite{Esfandiari:2014aa,Esfandiari:2015aa,Yesildirek:1995aa,Gao:2006aa}.
However, since DNNs offer greater expressive power, making them more effective for complex system approximations \cite{Rolnick:2018aa}, variations of DNNs, such as long short-term memory (LSTM) networks for time-varying dynamics \cite{Griffis:2023aa} and physics-informed NNs (PINNs) for leveraging physical system knowledge \cite{Hart:2024aa}, have further extended the capabilities of neuro-adaptive control systems.

A critical aspect of neuro-adaptive control is the weight adaptation law, which governs how NN weights are updated. 
Most studies derived these laws using Lyapunov-based methods, ensuring the boundedness of the tracking error and weight estimation error, thus maintaining system stability under uncertainty.

However, two significant challenges persist in using NNs for adaptive control. 
First, the boundedness of NN weights is not inherently guaranteed, which can result in unbounded outputs. 
When NN outputs are used directly in the control law, this may lead to excessive control inputs, violating input constraints. 
Such constraints are commonly encountered in industrial systems, where actuators are limited by physical and safety requirements in terms of amplitude, rate, or energy \cite{Esfandiari:2021aa}. 
Failing to address these constraints can degrade control performance or even destabilize the system.

Therefore, addressing these two key issues—ensuring weight boundedness and satisfying input constraints—is essential for the reliable design of neuro-adaptive controllers. 
The following section provides a detailed review of existing solutions to these challenges.

\subsection{Literature Review}

\subsubsection{Ensuring Weight Norm Boundedness}

A common challenge in neuro-adaptive control is maintaining the boundedness of the NN weights to ensure stability. 
In many studies, projection operators were employed to enforce upper bounds on the weight norms, ensuring that the weights do not grow unboundedly. 
For example, in \cite{Zhou:2023aa,Griffis:2023aa,Patil:2022aa}, projection operators were used to constrain the weight norms to remain below predetermined constants. 
However, these constants were often selected as large as possible due to the lack of theoretical guarantees regarding the global optimality of the weight values. 
While this approach ensured that the NN remained stable, it did not necessarily result in optimal performance.

In addition to projection operators, some studies utilized modification techniques like $\sigma$-modification \cite{Ge:2002aa} and $\epsilon$-modification \cite{Esfandiari:2015aa,Gao:2006aa}. 
These methods ensured that the NN weights remained within an invariant set by incorporating stabilizing functions into the adaptation law. 
Although these techniques were effective in ensuring boundedness and avoiding weight divergence, they similarly lacked a formal analysis of the optimality of the adapted weights by biasing the adaptation law toward the origin \cite{Ryu:2024aa}.
This leaves room for improvement in terms of performance optimization.

\hfill

\subsubsection{Satisfying Input Constraints}

The second major issue is satisfying input constraints, particularly in systems where actuators are subject to physical limitations. 
The unpredictable outputs of NNs can sometimes lead to excessively large control inputs, violating these constraints. 
This problem is exacerbated in neuro-adaptive controllers that attempt to cancel out system dynamics using conventional methods like feedback linearization or backstepping. 
In such cases, controllers may produce overly aggressive control inputs, even when the system's natural dynamics are stabilizing, leading to unnecessary saturation of the control inputs \cite{Khalil:2002aa}.

To address input saturation, many studies introduced auxiliary systems. 
These systems mitigated the effects of control input saturation by modifying the control strategy when saturation occurred. 
For instance, in \cite{Esfandiari:2014aa,Karason:1994aa,Esfandiari:2015aa}, auxiliary states were generated whenever input saturation was detected, and these states were incorporated into the adaptation law to adjust the NN weights accordingly. 
This approach helped the controller reduce input saturation by indirectly regulating the auxiliary states.

Alternatively, auxiliary states can also be used as feedback terms in the control law to directly compensate for the effects of input saturation constraints, as demonstrated in \cite{Arefinia:2020aa,He:2016aa,Peng:2020aa}. 
In \cite{Gao:2006aa}, the NN was used to approximate the desired control input, compensating for input saturation. 
However, these approaches typically handle input bound constraints on a per-input basis, \ie applying bounds to each scalar control variable individually, and may not account for more complex, nonlinear constraints, like input norm constraints, which are commonly found in physical systems such as robotic actuators or motor systems.

\hfill

\subsubsection{Limitations of Existing Approaches and Potential of Constrained Optimization}

Although both the projection operator methods for weight norm boundedness and the auxiliary system approach for input constraints have shown effectiveness, they come with significant limitations. 
Projection operators and modification techniques do not guarantee the optimality of the adapted weights. 
Moreover, auxiliary systems typically handle only simple forms of input constraints, such as input bound constraints, limiting their ability to address more complex, nonlinear constraints.

To overcome these limitations, constrained optimization offers a promising approach. 
By formulating the neuro-adaptive control problem as an optimization problem with constraints, it is possible to adapt the NN weights while minimizing an objective function, \eg tracking error, subject to both weight and input constraints. 
Constrained optimization provides a theoretical framework for defining optimality and presents numerical methods for finding solutions that satisfy the constraints \cite{Nocedal:2006aa}.

In existing literature, constrained optimization techniques, such as the Augmented Lagrangian Method (ALM) \cite{Evens:2021aa} and the Alternating Direction Method of Multipliers (ADMM) \cite{Wang:2019aa,Taylor:2016aa}, have been used to train NNs offline. 
These methods impose constraints to address issues like gradient vanishing in backpropagation. 
However, to the best of the authors' knowledge, no prior work has applied constrained optimization to adaptive control systems with real-time weight adaptation. 
This gap suggests that constrained optimization could be key to addressing both weight boundedness and input constraints in a unified, theoretically grounded framework, particularly in real-time neuro-adaptive control.

\subsection{Contributions}

The main contributions of this study are listed as follows:
\begin{itemize}
    \item A constrained optimization-based neuro-adaptive controller (CONAC) is proposed, where the design of a deep neural network (DNN) is formulated as a constrained optimization problem. In this formulation, weight and convex input constraints are incorporated as inequality constraints, while minimizing a given objective function.
    % \item Any convex input constraints can be applied to the proposed CONAC, including input bound constraints, and norm constraints. Moreover, the proposed CONAC can handle any combination of these constraints, which can be useful in practical applications where multiple constraints need to be satisfied simultaneously.
    % \item Since the proposed CONAC approximate the entire ideal control law with no prior knowledge of the system dynamics, the system identification process can be avoided, which is often difficult and time-consuming to implement in practice. 
    \item The adaptation laws for DNN weights in proposed CONAC are systematically derived to solve the defined problem, using constrained optimization theory. These laws guarantee convergence to the first-order optimality conditions, specifically the Karush–Kuhn–Tucker (KKT) conditions, as described in \cite[Chap. 12 Thm. 12.1]{Nocedal:2006aa}.
    \item The proposed CONAC's stability is rigorously analyzed using Lyapunov theory, ensuring that the tracking error and DNN weights remain bounded. This analysis guarantees that the proposed CONAC can achieve stable adaptation online.
    \item The proposed CONAC is implemented in real-time on a 2-DOF robotic manipulator, demonstrating its effectiveness in achieving accurate trajectory tracking while satisfying all imposed constraints. The experimental results validate the theoretical findings and show the practical applicability of the proposed controller.
\end{itemize}

\subsection{Organization}

The remainder of this paper is organized as follows. 
Section \ref{sec:Problem Formulation} presents the target system and control objective.
Section \ref{sec:ctrl design} introduces the proposed controller and the architecture of DNN in the controller. 
In Section \ref{sec:adap_laws}, the weight adaptation laws are developed.
Section \ref{sec:stability} analyzes the stability of the proposed controller.
A real-time implementation of the proposed controller is presented in Section \ref{sec:sim}, where the proposed controller is applied to a 2-DOF robotic manipulator.
Finally, Section \ref{sec:conclusion} concludes the paper and discusses potential future work.
The candidates of input constraints are presented in Appendix \ref{sec:appen:cstr}.

% ====================================
%  SECTION PROBLEM FORMULATION 
% ====================================
\section{Problem Formulation}\label{sec:Problem Formulation}

\subsection{Model Dynamics and Control Objective}

We consider an uncertain Euler-Lagrange system modeled as
\begin{equation}
    \rbM\ddq + \rbVm + \rbF + \rbG + \rbd
    =
    \mysat(\rbu)
    \label{eq:sys1}
\end{equation}
where $\q\in \R^n$ denotes the generalized coordinate, $\rbd$ represents the disturbance and $\rbu\in\R^n$ denotes the generalized control input. 
The terms $\rbM:=\rbM(\q)\in\R^{n\times n}$, $\rbVm:=\rbVm(\q,\dq)\in\R^{n\times n}$, $\rbF:=\rbF(\dq)\in\R^{n}$, and $\rbG:=\rbG(\q)$ represent the unknown inertia matrix, Coriolis/centripetal matrix, friction vector, and gravity vector, respectively.
The function $\mysat(\cdot):\R^n\to\R^n$ represents the inherent physical limitations of the actuators such that $\norm{\mysat(\rbu)}\le \overline\tau$, where $\overline\tau$ denotes the maximum norm of control input.
In this study, the saturation function $\mysat(\cdot)$ is assumed to be convex, which is practically common in many engineering systems, since the actuator limits are naturally convex, such as min/max torque limits, and bounds look like a box or ball in the input space.
To account for these limitations, it is essential to incorporate physically motivated constraints into the controller design.
Appendix \ref{sec:appen:cstr} introduces candidate constraints that can be applied to ensure compliance with these physical limitations.

The Euler-Lagrange system \eqref{eq:sys1} satisfies several important physical properties, as presented in \cite[Chap. 3 Tab. 3.2.1]{Lewis:1998aa}. 
The key properties are introduced below:
\begin{prop} 
    The inertia matrix $\rbM$ is symmetric, positive definite, and bounded.
    \label{prop:M}
\end{prop}

\begin{prop} 
    The Coriolis/centripetal matrix $\rbVm$ can always be selected, so that the matrix $\ddtt{\rbM}-2\rbVm$ is skew-symmetric, \ie $\mv{x}^\top(\ddtt{\rbM}-2\rbVm)\mv{x}=0,\ \forall \mv{x}\in\R^n$.
    \label{prop:skew}
\end{prop}

\begin{prop}
    The disturbance $\rbd$ is bounded, \ie $\norm{\rbd}\le \overline\tau_d$.
    \label{prop:dis_bound}
\end{prop}

The control objective is to develop a neuro-adaptive controller that enables $\q$ to track a continuously differentiable desired trajectory $\qd:=\qd(t): \R_{\ge0} \to \R^n$, compensating for the unknown system dynamics while addressing the imposed constraints.
The desired trajectory $\qd(t)$ is supposed to satisfy the following assumption:
\begin{assum}
    The desired trajectory $\qd(t)$ is assumed to be bounded, i.e., $\norm{\qd(t)}\le \overline{q}_d$ , and available for designing a bounded control input.
    \label{assum:feasible}
\end{assum}

%  SECTION CONTROLLER DESIGN ==============================
\section{Control Law Development}\label{sec:ctrl design}

The architecture of the proposed CONAC is illustrated in Fig.~\ref{fig:ctrl:diagram}, and consists of a DNN that functions as a neuro-adaptive controller, and a weight optimizer for the DNN.
Before proceeding to the controller design, some mathematical preliminaries are revisited in Section \ref{sec:sub:math preliminaries}. 
Section \ref{sec:sub:NAC} introduces the neuro-adaptive controller, followed by the DNN model in Section \ref{sec:sub:NN definition}. 
The constrained optimization-based weight adaptation laws are presented in Section \ref{sec:adap_laws}.

\begin{figure}[!t]
    \centering
    \includegraphics[width=.99\linewidth]{src/figures/Controller.drawio.pdf}
    % \includesvg[width=0.75\linewidth]{Controller.drawio.svg}
    \caption{Architecture of the constrained optimization-based neuro-adaptive controller (CONAC).}
    \label{fig:ctrl:diagram}
\end{figure}

\subsection{Mathematical Preliminaries}\label{sec:sub:math preliminaries}

The following proposition and lemma will be used in subsequent sections:

\begin{propsit}[see, {\cite[Chap. 7 Prop. 7.1.9]{Bernstein:2009aa}}] \label{propsit:kron}
	For matrix $\mm A\in\R^{n\times m}$ and vector $\mv b\in\mathbb R^n$, the following property holds:
	\begin{equation}
		\mm A^\top \mv b 
		= 
		\myvec(\mm A^\top\mv b)
		=
		\myvec(\mv b^\top\mm A)
		= 
		(\mm{I}_{m}\otimes \mv b^\top) \myvec(\mm A)
		.
	\end{equation}
\end{propsit}

\begin{lem} \label{lem:stable:set}
    For Lyapunov function $V:=V(\mv{x})\ge0$ with $\mv{x}\in\R^n$, if the time derivative of $V$ is given by $\ddtt V\le -a_1 \mv{x}^\top \mv{x} + a_2 \mv{x}$ for some positive constants $a_1\in\R$ and $a_2\in\R$, then $\mv{x}$ is bounded as $\mv{x}\in\{\mv{x}\mid\norm{\mv{x}}\le\tfrac{a_2}{a_1}\}$.
\end{lem}

\begin{proof}
    The time derivative of $V$ can be rewritten as
    \begin{equation}
        \ddtt V
        \le
        -a_1 \norm{\mv{x}}^2 + a_2 \norm{\mv{x}}
        =
        -a_1
        \left(
            \norm{\mv{x}}-\tfrac{a_2}{2a_1}
        \right)^2
        -
        \tfrac{a_2^2}{4a_1}
        .
        \label{eq:lem:lya:dot}
    \end{equation}
    According to \eqref{eq:lem:lya:dot}, one can conclude that $\ddtt V$ is negative definite if $\norm{\mv{x}}\le\tfrac{a_2}{a_1}$ holds.
    In other words, $\mv{x}$ remains within the bounded set $\{\mv{x}\mid\norm{\mv{x}}\le\tfrac{a_2}{a_1}\}$.
\end{proof}

\subsection{Neuro-Adaptive Controller Design}\label{sec:sub:NAC}

First, the filtered tracking error $\fe\in\R^n$ is defined as 
\begin{equation}
    \fe := \ddtt{\mv e}-\mm\Lambda \mv e
    ,
    \label{eq:err:filtered}
\end{equation}
where $\mv e := \q - \qd$ denotes the tracking error, $\ddtt{\mv e} := \dq - \dqd$ represents the time derivative of the tracking error, and $\mm\Lambda\in\R^{n\times n}_{>0}$ is a user-designed filtering matrix.
Since \eqref{eq:err:filtered} is stable system, it implies that $\mv e$ is bounded if $\fe$ is bounded.

Using $\fe$, the system dynamics \eqref{eq:sys1} can be rewritten as
\begin{equation}
    \rbM \ddtt\fe
    =
    -\rbVm \fe
    -\mm K \fe
    + \mv f
    -\rbd + \mysat(\rbu)
    ,
    \label{eq:sys:filtered}
\end{equation}
where $\mm{K}=\mm{K}^\top\in\R^{n\times n}_{>0}$ denotes an arbitrary unknown matrix and $
    \mv f:= \mv f(\q,\dq,\qd,\dqd,\ddqd)
    =
    \mm K \fe
    +\rbM(-\ddqd+\mm\Lambda\ddtt{\mv e})
    +
    \rbVm(-\dqd+\mm\Lambda\mv e)
    -
    \rbF
    -
    \rbG
    \in\R^n
$ denotes the lumped system uncertainty.
% \color{red}
% The unknown matrix $\mm{K}$ is not available to design in practice, 
% \color{black}

Consider the Lyapunov function $V_1:= \tfrac{1}{2}\fe^\top \rbM\fe$. 
Invoking Property \ref{prop:skew}, the time derivative of $V_1$ is
\begin{equation}
    \begin{aligned}
        \ddtt{V}_1
        =&
        \fe^\top \rbM \ddtt \fe
        +
        \tfrac{1}{2}\fe^\top \ddtt{\rbM}\fe
        \\
        =&
        \fe^\top (
            -\rbVm \fe -\mm K \fe + \mv f
            -\rbd + \mysat(\rbu)
        )
        \\
        &
        +
        \tfrac{1}{2}
        \fe^\top \ddtt{\rbM}\fe
        \\
        =&
        -
        \fe^\top \mm K \fe 
        +
        \fe^\top (
            \mv f+\mysat(\rbu)
            -\rbd
        )
        \\
        &
        +
        \tfrac{1}{2}
        \fe^\top(\ddtt{\rbM}- 2\rbVm)\fe
        \\
        \le&
        -\lambda_{\min}(\mm K)\norm{\fe}^2
        +
        \overline\tau_d\norm{\fe}
        +
        \fe^\top (\mv f+\mysat(\rbu))
        % \\
        % \le&
        % -\lambda_{\min}(\mm K)\norm{\fe}^2
        % +
        % \overline\tau_d\norm{\fe}
        % +
        % \fe^\top (\mysat(\rbu)-\rbu^*)
        .
    \end{aligned}
    \label{eq:lya1:dot1}
\end{equation} 
Therefore, invoking Lemma \ref{lem:stable:set}, it can be concluded that the filtered error $\fe$ is exponentially stable, with $\lim_{t\to\infty}\norm{\fe}=\tfrac{\overline\tau_d}{\lambda_{\min}(\mm K)}$, provided that the lumped system uncertainty $\mv{f}$ can be perfectly cancelled by the control input $\rbu$, neglecting the effect of input saturation.
Hence, the ideal control input $\rbu^*$ is defined as $-\mv f$.
However, $\rbu^*$ is not available in practice, since system dynamics encapsulated in $\mv f$ are unknown.

To overcome this issue, a DNN is employed to approximate $\rbu^*$.
Let $\NN:=\NN({\q}_n;\wth): \R^{l_0+1}\times\R^{\Xi}\to\R^{n}$ represent the DNN, where ${\q}_n\in\R^{l_0+1}$ is the DNN input vector, and $\wth\in\R^{\Xi}$ is the vector of trainable weights.
The architecture of $\NN({\q}_n;\wth)$ will be presented in Section \ref{sec:sub:NN definition}.
According to the universal approximation theorem for DNNs \cite{Kidger:2020aa}, $\NN({\q}_n;\wth)$ can approximate a nonlinear function, denoted as $\mv g(\cdot)$, with an ideal weight vector $\idealwth$ over a compact subset $\Omega_{NN}\in\R^{l}$ to within $\epsilon$-accuracy, such that $\sup_{{\q}_n\in\Omega_{NN}}\norm{ \NN({\q}_n;\idealwth) - \mv g(\cdot)} = \epsilon < \infty$.
% Furthermore, the theorem states that the norm of $\idealwth$ is bounded, \ie $\norm{\idealwth}\le \overline\wth<\infty$.
In this study, $\idealwth$ is considered a locally optimal solution, rather than a global one.

Accordingly, the ideal control law $\rbu^*$ can be approximated by the DNN with the ideal weight vector $\idealNN:=\NN({\q}_n;\idealwth)$ as follows:
\begin{align}
    \rbu^*=& \idealNN+\mv\epsilon
    ,
    \label{eq:control:ideal}
\end{align}
where $\mv\epsilon\in\R^n$ is the approximation error vector, bounded by $\norm{\mv\epsilon}\le \bar{\epsilon}$.
The control input $\rbu^*$ is estimated online by
\begin{align}
    \rbu =& \estNN,
    \label{eq:control:est}
\end{align}
where $\hat\NN:=\NN({\q}_n;\estwth)$, and  $\estwth$ is the estimated weight vector for $\idealwth$.

Substituting \eqref{eq:control:ideal} and \eqref{eq:control:est} into \eqref{eq:lya1:dot1}, the time derivative of $V_1$ can be rewritten as
\begin{equation}
    \ddtt{V}_1
    \le 
    -\lambda_{\min}(\mm K)\norm{\fe}^2
    +
    \overline\tau_d\norm{\fe}
    +
    \fe^\top (\mysat(\hat\NN)-\idealNN-\mv\epsilon)
    .
    \label{eq:lya1:dot2}
\end{equation}
Therefore, one can conclude that the result of \eqref{eq:lya1:dot1} can be obtained by adapting $\estwth$ to $\idealwth$ (\ie $\hat\NN\to\idealNN$), neglecting the effect of input saturation.
Note that the control input saturation has not yet been considered in the derivation of the ideal control input; this will be addressed in the subsequent section.

\subsection{Deep Neural Network (DNN) Model}\label{sec:sub:NN definition}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.99\linewidth]{src/figures/DNN.drawio.pdf}
    \caption{
        Architecture of the DNN $\NN({\q}_n;\wth)$ with $k=2,l_0=2,l_1=l_2=3$, and $l_3=2$.
    }
    \label{fig:DNN}
\end{figure}

The DNN architecture $\NN({\q}_n;\wth) := \NN_k$ can be recursively represented as follows:
\begin{equation}
    \NN_i :=
    \begin{cases}
        \wV_i^\top \act_i(\NN_{i-1}), 
        &
        i\in\{1,\dots ,k\},
        \\
        \wV_0^\top {\q}_n,
        &
        i=0
        ,
    \end{cases}
    \label{eq:DNN:architecture}
\end{equation}
where $\wV_i\in\R^{(l_i+1)\times l_{i+1}}$ is the weight matrix of the $i\textsuperscript{th}$ layer, and $\act_i: \R^{l_i}\to\R^{l_i+1}$ represents the activation function of the $i\textsuperscript{th}$ layer. 
The activation function is defined as $\act_i(\mv x)=\left(\sigma(x_1),\sigma(x_2),\cdots, \sigma(x_{l_{i}}), 1\right)^\top,\ \forall\mv x\in\R^{l_i}$, where $\sigma: \R\to\R$ is a nonlinear function, and the augmentation of $1$ is used to account for bias terms in the weight matrices. 
Notice that the output size of $\NN(\cdot)$ is the same as that of the control input $\rbu$ (\ie $l_{k+1}=n$).

One of the widely used activation functions for large DNNs is from the ReLU family \cite{Maas:2013aa}, which effectively avoids the gradient vanishing problem during error backpropagation. 
However, for control applications, where relatively shallow DNNs are typically sufficient, and the gradient vanishing issue is less severe, the sigmoid function or the hyperbolic tangent function is commonly used as the activation function. 
These functions simplify stability analysis, since they are continuously differentiable and their outputs and gradients are bounded. 
In this study, the hyperbolic tangent function $\tanh(\cdot)$ was selected as the activation function (\ie $\sigma(x) = \tanh(x),\ \forall x\in\R$), which provides desirable boundedness with $\norm{\sigma(x)}<1$ and $\norm{\ddtfrac{\sigma(x)}{x}}\le 1$.

For simplicity, each layer's weights are vectorized as $\wth_i:=\myvec(\wV_i)\in\R^{\Xi_i}$, where $\Xi_i:= (l_i+1)l_{i+1}$ is the number of weights in the $i\textsuperscript{th}$ layer. 
The total weight vector $\wth\in\R^{\Xi}$ is defined by augmenting $\wth_i$ for all $i\in \{0,\cdots,k\}$ as 
\begin{equation}
    \wth := 
    \begin{pmatrix}
        \wth_k\\
        \wth_{k-1}\\
        \vdots\\
        \wth_0
    \end{pmatrix}
    =
    \begin{pmatrix}
        \myvec(\wV_k)\\
        \myvec(\wV_{k-1})\\
        \vdots\\
        \myvec(\wV_0)
    \end{pmatrix},
\end{equation}
where $\Xi={\sum_{i=0}^{k} \Xi_i}$ represents the total number of weights. 
The gradient of $ \NN({\q}_n;\wth)$ with respect to $\wth$ can be obtained using Proposition \ref{propsit:kron} and the chain rule as follows:
\begin{equation}
    \pptfrac{\NN}{\wth}=
    \begin{bmatrix}
        \pptfrac{\NN}{\wth_k}&
        \pptfrac{\NN}{\wth_{k-1}}&
    \cdots &
        \pptfrac{\NN}{\wth_0}
    \end{bmatrix}
    \in\R^{n \times \Xi}
    \label{eq:gradient:phi}
\end{equation}
where
\begin{equation}
    \pptfrac{\NN}{\wth_i} = 
    \begin{cases}
        (
            \mm I_{l_{k+1}}
            \otimes 
            \act_{k}^\top  
        )
        , 
        &
        i=k
        ,
        \\
        \wV_k^\top\act_{k}' 
        (
            \mm I_{l_{k}}
            \otimes  
            \act_{k-1}^\top  
        )
        , 
        & 
        i=k-1
        ,
        \\
        &
        \vdots 
        \\
        \wV_k^\top\act'_{k} 
        \cdots 
        \wV_1^\top\act_1' 
        (
            \mm I_{l_1}
            \otimes 
            {\q}_n^\top  
        )
        , 
        &
        i = 0
        ,
    \end{cases}
    \label{eq:NN:grad}
\end{equation}
and $\act_i:= \act_i(\NN_{i-1})$ and $\act_i':= \ddtfrac{\act_i}{\NN_{i-1}}$.

In the following sections, let $\idealNN_i$ represent the output of the $i\textsuperscript{th}$ layer with the ideal weight vector $\idealwth$. 
Additionally, define $\idealact_i:=\act_i(\idealNN_{i-1})$ and ${\idealact}'_i:= \ddtfrac{\idealact_i}{\idealNN_{i-1}}$. 
Similarly, $\estNN_i$ denotes the output of the $i\textsuperscript{th}$ layer with the estimated weight vector $\estwth$, and define $\estact_i:=\act_i(\estNN_{i-1})$ and $\estact_i':= \ddtfrac{\estact_i}{\estNN_{i-1}}$.

%  SECTION ADAPTATION LAW DERIVATION =======================
\section{Weight Adaptation Laws}\label{sec:adap_laws}

% \subsection{Adaptation Law using Lagrangian Function}
\subsection{Weight Optimizer Design}\label{sec:sub:weight optimizer}

The control objective can be represented as follows:
\begin{equation}
    \begin{matrix}
        \min_{\estwth} \ J(\fe;\estwth):= 
        \tfrac{1}{2} \fe^\top \fe
        \\ \\
        \begin{aligned}
        \text{subject to }&c_{j}(\estwth) 
        \le0, \quad \forall j\in\mathcal{I},
        \end{aligned}
    \end{matrix}
    \label{eq:opt:problem1}
\end{equation}
where $J(\fe;\estwth):= \tfrac{1}{2} \fe^\top \fe$ is the objective function.
Inequality constraints $c_j:=c_j(\estwth),\ \forall j\in\mathcal{I}$, are imposed during the weight adaptation process to minimize the objective function, where $\mathcal I$ denotes the set of the imposed inequality constraints. 
Here, $\fe$ is considered a pre-defined data or static parameter for this optimization problem, so that the adaptation process does not consider the dynamics of $\fe$.
The Lagrangian function is defined as
\begin{equation}
    L(
        \fe,\estwth,[\lambda_j]_{j\in\mathcal I}
    ) 
    := 
    J(\fe;\estwth) 
    + 
    \sum_{j\in\mathcal I}
    \lambda_{j}
    c_{j}(\estwth)
    ,
    \label{eq:lagrangian}
\end{equation}
where $\lambda_j$ denotes the Lagrange multiplier for each constraint.

The adaptation laws for $\estwth$ and $[\lambda_j]_{j\in\mathcal I}$ are derived to solve the dual problem of \eqref{eq:opt:problem1} (\ie  $\min_{\estwth} \max_{[\lambda_j]_{j\in\mathcal I}}L(\fe,\estwth,[\lambda_j]_{j\in\mathcal I})$), as follows:
\begin{subequations}
    \begin{align}
            \ddtt{\estwth}
            &
            =
            -\alpha \pptfrac{L}{\estwth}
            =
            -\alpha 
            \left(
                \pptfrac{J}{\estwth}
                +
                \textstyle\sum_{j\in\mathcal{I}}
                \lambda_j 
                \pptfrac{c_j}{\estwth}
            \right),
        \label{eq:adap:th}
            \\
            \ddtt\lambda_j
            & 
            = 
            \beta_j\pptfrac{L}{\lambda_j} 
            = 
            \beta_j c_j ,
            \quad \forall j\in\mathcal I,
        \label{eq:adap:lbd}
            \\
            \lambda_j & \leftarrow \max(\lambda_j,0) ,
        \label{eq:adap:lbd:max}
    \end{align}
    \label{eq:adaptation law}%             %%
\end{subequations}
where $\alpha\in\R_{>0}$ denotes the adaptation gain (learning rate) and $\beta_j\in\R_{>0}$ denotes the update rate of the Lagrange multipliers in $\mathcal I$, and the arguments of $L(\cdot)$ and $J(\cdot)$ are suppressed for brevity. 
The Lagrange multipliers associated with inequality constraints are non-negative, \ie $\lambda_j\ge 0$, due to \eqref{eq:adap:lbd:max}. 
When a constraint $c_j$ becomes active, \ie $c_j > 0$, the corresponding Lagrange multiplier $\lambda_j$ increases to address the violation, \ie see, \eqref{eq:adap:lbd}.
Once the violation is resolved and the constraint is no longer active, \ie $c_j \le 0$, the multiplier decreases gradually until it returns to zero.
Note that this adaption law is similar to the ALM in \cite{Nocedal:2006aa}, where the adaptation law for Lagrange multipliers is given by $\lambda_j\leftarrow \max\left(\lambda_j+\tfrac{c_j}{\mu},0\right)$, with $\mu\in\R_{>0}$ being the penalty parameter.

At steady state, where $\ddtt{\estwth}=0$ and $\ddtt\lambda_j=0$, the KKT conditions are satisfied, \ie $\pptfrac{L}{\estwth}=0$, $c_j \le 0$, $\lambda_j \ge 0$, and $\lambda_j c_j=0$.
In other words, the proposed optimizer updates $\estwth$ and $\lambda_j$ in a way that satisfies the KKT conditions. 
These conditions represent the first-order necessary conditions for optimality, guiding the updates toward candidates for a locally optimal point.
% The KKT condition is satisfied, when $\estwth$ converges to $\idealwth$, since $\partial L/\partial \estwth=0$ (i.e. the first-order necessary condition of the optimality is satisfied.).

\subsection{Approximation of the Gradient of Objective Function}

The adaptation law for $\estwth$ defined in \eqref{eq:adap:th} involves the partial derivative of the filtered error with respect to control input $\pptfrac{\fe}{\rbu}$, \ie 
$
    \pptfrac{J}{\rbu}
    =
    \left(
        (\pptfrac{\fe}{\rbu})
        (\pptfrac{\rbu}{\estwth})
    \right)
    ^\top \fe
    =
    \left(
        (\pptfrac{\fe}{\rbu})
        (\pptfrac{\estNN}{\estwth})
    \right)
    ^\top \fe
$. 
Since the objective function depends on $\fe$ of a dynamic system, obtaining $\pptfrac{\fe}{\rbu}$ is not straightforward. 
The recommended method to calculate the exact value of $\pptfrac{\fe}{\rbu}$ is to use the forward sensitivity method \cite{Sengupta:2014aa} by simulating the sensitivity equation as follows: 
\begin{equation}
    \ddtt 
    \left(
        \pptfrac{\fe}{\rbu}
    \right)
    =
    \pptfrac{}{
        \rbu
    }
    \left(
        \rbM^{-1}(
            -\rbVm \fe
            -\mm K \fe
            + \mv f
            -\rbd + \mysat(\rbu)
        )
    \right)
    .
\end{equation}
However, this method cannot be realized, since we do not have the exact system dynamics, \ie $\rbM,\rbVm, \rbF$ and $ \rbG$.
Additionally, the computational cost of the forward sensitivity method is high, as the number of DNN's weights are generally large.

In \cite{Douratsos:2007aa,Saerens:1991aa}, the authors approximate $\pptfrac{\fe}{\rbu}$ by taking the sign of each entry, \ie
$
    \pptfrac{\fe}{\rbu}
    \approx
    \left[
        \mysign(\pptfrac{r_i}{\tau_j})
    \right]_{i,j\in\{1,\cdots,m\}}
$ 
assuming known control directions.
However, this approach is not applicable to \eqref{eq:sys:filtered}, where control directions are unknown.
Instead, since the sign of the control input matrix is known—owing to the positive definiteness of $\rbM^{-1}$, see, Property \ref{prop:M}—we approximate $\pptfrac{\fe}{\rbu}\approx \mm I_n$.
Consequently, the adaptation law in \eqref{eq:adap:th} becomes:
\begin{equation}
    \ddtt{\estwth}
    \approx
    -\alpha 
    \left(
        \pptfrac{\estNN}
        {\estwth}^\top
        \fe
        +
        \textstyle\sum_{j\in\mathcal{I}}
        \lambda_j 
        \pptfrac{c_j}{\estwth}
    \right)
    .
    \label{eq:adap:th:approx}
\end{equation}

%% ALGORITHM PART; Now omitted!
% The proposed controller is implemented using Algorithm \ref{alg: alg1}. For implementation in the discrete-time domain, it is recommended to use a sufficiently small sampling time $T_s$. If a large $T_s$ is used, $\alpha$ and $\beta_j$ should satisfy the Armijo condition \cite[Chap.~3 eq.~(3.4)]{Nocedal:2006aa} to ensure that the objective function decreases.

% \begin{algorithm}[!t]
%     \caption{Weight Optimizer Implementation.}\label{alg: alg1}
%     \begin{algorithmic}[1]
%         \renewcommand{\algorithmicrequire}{\textbf{Input:}}
%         \renewcommand{\algorithmicensure}{\textbf{Output:}}
%         \REQUIRE $\xi$, $\estwth$, $\lambda_j$, $\eta$
%         \ENSURE  $\estwth$, $\lambda_j$, $\eta$
%         \STATE Set $\mathcal A \leftarrow \mathcal A\cup \{j\}$ for all $c_j\ge0$;
%         \STATE Determine update matrix $\ddtt\eta$ using \eqref{eq. eta dynamics 1};
%         \STATE Update $\eta\leftarrow \eta +\ddtt\eta\cdot T_s$; 
%         \STATE Determine update directions $\ddtt{\estwth}$, $[\ddtt\lambda_j]_{j\in\mathcal A}$ using \eqref{eq. adaptation law th}, \eqref{eq. adaptation law L};
%         \STATE Update weight vector $\estwth\leftarrow \estwth+\ddtt{\estwth}\cdot T_s$;
%         \STATE Update multipliers $[\lambda_j]_{j\in\mathcal A}\leftarrow [\lambda_j]_{j\in\mathcal A}+[\ddtt\lambda_j]_{j\in\mathcal A}\cdot T_s$;
%         \STATE $[\lambda_j]_{j\in\mathcal A}\leftarrow \max([\lambda_j]_{j\in\mathcal A}, 0)$;
%         \STATE Set $\mathcal A \leftarrow \mathcal A - \{j\}$ for all $\lambda_j=0$;
%         \STATE \textbf{RETURN};
%     \end{algorithmic}
%     \label{alg1}
% \end{algorithm}

\subsection{Essential and Potential Constraint Candidates}\label{sec:sub:cstr}

This section introduces essential constraints required to ensure the stability of the proposed CONAC, as well as candidate conditions for potential control input constraints.

First, weight norm constraints are essential to prevent the weights from diverging during the adaptation process.
The weight norm constraints $\mv{c}_{\theta}(\estwth):= [c_{\theta_i}(\estwth)]_{i\in\{0,\cdots ,k\}}\in\mathbb R^{k+1}$ are defined for each layer's weight as follows:
\begin{equation}
    c_{\theta_i}
    =
    \tfrac{1}{2}
    \left(
        \Vert \estwth_i\Vert^2 
        -
        \overline\theta_i^2 
    \right)    
    % \le 0
    \label{eq:cstr:weight:ball}
\end{equation}
with $\overline\theta_i<\infty$ denoting the maximum allowable norm for $\estwth_i$, and the arguments of $c_{\theta_i}(\cdot)$ are omitted for brevity.
The gradient of $\mv{c}_\theta:=\mv c_{\theta}(\estwth)$ with respect to $\estwth$ is given by
\begin{equation}
    \pptfrac{c_{\theta_i}}{\estwth_j} 
    =
    \begin{cases}
        \estwth_i,
        &
        \text{if } i=j,
        \\
        \mv{0},
        &
        \text{if } i\ne j
        .
    \end{cases} 
    \label{eq:cstr:weight:ball:grad}
\end{equation}

Next, we present the following assumptions, which specify the conditions for the potential control input constraints introduced in Appendix \ref{sec:appen:cstr}:

\begin{assum}
    The constraint functions $c_j(\estwth),\ \forall j\in\mathcal I$, are convex in the $\tau$-space and satisfy $c_j(0) \le 0$ and $c_j(\idealwth)\le 0$.
    \label{assum:convex}
\end{assum}

\begin{assum}
    The selected constraints satisfy the Linear Independence Constraint Qualification (LICQ) \cite[Chap.~12 Def.~12.1]{Nocedal:2006aa}.
    \label{assum:LICQ}
\end{assum}

\begin{remark}
    Assumption \ref{assum:convex} is not restrictive, since, as aforementioned, the control input saturation is practically convex.
    Moreover, Assumption \ref{assum:LICQ} is a standard assumption in optimization problems, which ensures that the gradients of the active constraints are linearly independent.
    This assumption will be used in the stability analysis (see, Lemma \ref{lem:convex:angle}).
\end{remark}

%  SECTION STABILITY ANALYSIS ==============================
\section{Stability Analysis}\label{sec:stability}

Before conducting the stability analysis, let us define the weight estimation error as $\tilde\wth:= [\tilde\wth_i]_{i\in\{0,\cdots,k\}}$, where $\tilde\wth_i:=\estwth_i-\idealwth_i$.
The following lemmas are introduced for the stability analysis.

\begin{lem}
    If Assumptions \ref{assum:convex} and \ref{assum:LICQ} hold, then for each active constraint $c_j$, the angle between $\pptfrac{c_j}{\estwth_k}$ and $\estwth_k$ is positive; that is $\pptfrac{c_j}{\estwth_k}^\top\estwth_k>0$.
    \label{lem:convex:angle}
\end{lem}

\begin{proof}

Since $\rbu = \estNN$, using Proposition \ref{propsit:kron}, a linear mapping $\mm T(\cdot):\estwth_k\to\rbu$ that is independent of $\estwth_k$, can be derived using Proposition \ref{propsit:kron}:
\begin{equation}\label{eq:linear:map}
    \begin{aligned}
    \rbu 
    = 
    &
    \estNN 
    % = 
    % \myvec(\estNN)
    =
    \myvec(
        \estwV_k^\top \estact_k
    ) 
    = 
    (
        \mm I_{l_{k+1}}
        \otimes 
        \estact_k^\top
    )^\top
    \myvec(\estwV_k)
    \\
    = &
    \underbrace{
        (
        \mm I_{l_{k+1}}
        \otimes 
        \estact_k^\top
    )^\top}_{=:\mm T(\estact_k)}
    \estwth_k 
    =
    \mm T(\estact_k) \estwth_k.
    \end{aligned}
\end{equation}
Therefore, the convexity of the input constraints in the $\rbu$-space (see, Assumption \ref{assum:convex}) is preserved in $\estwth_k$-space, implying
that $\pptfrac{c_j}{\estwth_k}^\top\estwth_k>0$.

\end{proof}

\begin{lem} 
    If control input constraint $c_j(\estwth),\ \forall j\in\mathcal I \setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}$ satisfies Assumption \ref{assum:convex}, then $\norm{\pptfrac{c_j}{\estwth_i}},\ \forall i\in\{k-1,\cdots, 0\}$, is bounded, provided the norms of $\estwth_i,\ \forall i\in\{k,\cdots, i+1\}$, remain bounded.
    \label{lem:cstr:grad:bound}
\end{lem}

\begin{proof}

The derivative of $c_j,\ \forall j\in\mathcal I \setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}$, with respect to $\estwth_i$ is represented as
\begin{equation}
    \pptfrac{c_j}{\estwth_i} 
    = 
    \pptfrac{c_j}{\rbu} 
    \pptfrac{\rbu}{\estNN} 
    \pptfrac{\estNN}{\estwth_i}
\end{equation}
where $\pptfrac{\tau}{\estNN}=\mm I_n$, which is bounded. 
From the linear mapping in \eqref{eq:linear:map}, $\rbu$ is bounded as long as $\estwth_k$ is bounded (by the condition of the lemma), and $\norm{\act_k(\cdot)}$ is bounded due to the properties of the activation functions. 
By Assumption \ref{assum:convex}, the function $c_j$ is convex in $\rbu$-space.
The convex function has a bounded derivative with respect to $\rbu$, since $\rbu$ is a bounded variable (\ie $\pptfrac{c_j}{\rbu}$ is bounded). 
Furthermore, $\pptfrac{\estNN}{\estwth_i}$ is bounded, provided that the norms of $\estwth_i,\ \forall i \in\{k,\cdots,i+1\}$, are bounded. 
This can be verified by using the definition of $\pptfrac{\estNN}{\estwth_i}$ given in \eqref{eq:NN:grad}.
Consequently, $\norm{\pptfrac{c_j}{\estwth_i}},\ \forall j\in\mathcal I \setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}$, is bounded, when $\estwth_i,\ \forall i\in \{k,\cdots,i+1\}$ are bounded.

\end{proof}

The following theorem states that $\estwth$ and $\fe$ are bounded.

\begin{theorem}
    For the dynamical system described in \eqref{eq:sys1}, the neuro-adaptive controller in \eqref{eq:control:est} with the weight adaptation laws in \eqref{eq:adaptation law} ensure the boundedness of the filtered error $\fe$ and the weight estimate $\estwth$, under the control input constraints satisfying Assumption \ref{assum:convex} and \ref{assum:LICQ}.
    This holds under the weight norm constraint \eqref{eq:cstr:weight:ball}.
\end{theorem}

\begin{proof}

% The boundednesses of $\estwth$ and $\fe$ are analyzed recursively from the last $k\textsuperscript{th}$ layer to the first layer of $\estNN$. 
Invoking the fact that the amplitude of the control input depends only on $\estwth_k$, \ie see, \eqref{eq:control:est}, and \eqref{eq:DNN:architecture} and $\norm{\act_k}<\infty$, the boundedness of inner layers' weights follows once the boundedness of $\estwth_k$ and $\fe$ is established.
Without loss of generality, weight norm constraint $c_{\theta_i},\ \forall i\in\{0,\cdots,k\}$ is supposed to be active.
This assumption is justified because, even if $c_{\theta_i}$ is inactive, $\estwth$ is adapted to minimize the objective function $J(·)$ as long as the constraint is not violated.
We consider two cases: (1) the control input is saturated, and (2) the control input is not saturated.

\subsection*{Case 1: Control input is saturated.}

As a result of the time derivative of $V_1$ in \eqref{eq:lya1:dot2}, the invariant set $\Theta_{\fe}^{1}$ of $\fe$ (\ie if $\fe$ leaves $\Theta_{\fe}^{1}$, $\ddtt V_1$ is negative) can be obtained using Lemma \ref{lem:stable:set} as follows:
\begin{equation}
    \Theta_{\fe}^{1} 
    = 
    \left\{ 
        \fe \in \R^n 
        \mid 
        \norm{\fe} 
        \le 
        \tfrac{
            \overline\tau_d
            +
            \overline\tau
            +
            \bar{\theta}_k\sqrt{l_{k}+1}
            +
            \bar{\epsilon}
        }{
            \lambda_{\min}(\mm K)
        }
    \right\}
    .
    \label{eq:invariant:c1:fe}
\end{equation}
It is notable that the ideal DNN $\idealNN_k={\idealwV_k}^\top\idealact$ is bounded by $\bar{\theta}_k\sqrt{l_{k}+1}$, since $\norm{\idealwV_k}_F = \norm{\idealwth}_k \le \bar{\theta}_k$, and $\norm{\idealact_k}\le \sqrt{l_{k}+1}$.

To investigate the boundedness of $\estwth_k$, consider the Lyapunov function candidate $V_2:=\tfrac{1}{2\alpha}\estwth_k^\top \estwth_k$.
Taking the time derivative of $V_2$ yields:
\begin{equation}
    \begin{aligned}
        \ddtt  V_2 
        =& 
        -\estwth_k^\top 
        \left(
            (\mm I_{l_{k+1}}\otimes \estact_k^\top)^\top
            \fe
            +
            \textstyle\sum_{j\in\mathcal{I}}
            \lambda_j 
            \pptfrac{c_j}{\estwth_k}
        \right)
        \\
        =&
        -\estwth_k^\top 
        \Big(
            (\mm I_{l_{k+1}}\otimes \estact_k^\top)^\top
            \fe
            +
            \lambda_{\theta_k}\estwth_k
        \\
        &
            +
            \textstyle\sum_{
                j\in\mathcal{I}\setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}
            }
            \lambda_j 
            \pptfrac{c_j}{\estwth_k}
        \Big)
        \\
        \le&
        -
        \lambda_{\theta_k}\norm{\estwth_k}^2
        +
        \underbrace{
            \norm{(\mm I_{l_{k+1}}\otimes {\estact}_k^\top)}
        }_{=:{\gamma}_1\in\R_{>0}}
        \norm{\fe}
        \norm{\estwth_k}
        \\
        &
        \underbrace{
            -
            \textstyle\sum_{
                j\in\mathcal{I}\setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}
            }
            \lambda_j 
            \estwth_k^\top \pptfrac{c_j}{\estwth_k}
        }_{=:{\gamma}_2<0,\ \text{by Lemma \ref{lem:convex:angle} and Assumption \ref{assum:LICQ}}}
        \\
        \le&
        -
        \lambda_{\theta_k}\norm{\estwth_k}^2
        +
        {\gamma}_1     
        \norm{\fe}
        \norm{\estwth_k}
        .
    \end{aligned}
    \label{eq:lya2:dot}
\end{equation}
According to \eqref{eq:lya2:dot} and Lemma \ref{lem:stable:set}, the invariant set of $\estwth_k$ is defined as
\begin{equation}
    \Theta_{\theta_k}^{1} 
    = 
    \left\{ 
        \estwth_k \in \R^{\Xi_{k}} 
        \mid 
        \norm{\estwth_k} 
        \le 
        \tfrac{
            {\gamma}_1
            (
                \overline\tau_d
                +
                \overline\tau
                +
                \bar{\theta}_k\sqrt{l_{k}+1}
                +
                \bar{\epsilon}
            )
        }{
            \lambda_{\theta_k}\lambda_{\min}(\mm K)
        }
    \right\}
    .
    \label{eq:invariant:c1:wth}
\end{equation}

The satisfaction of the constraints can be verified through the boundedness of $\estwth_k$.
For the weight norm constraint $c_{\theta_k}$, if the corresponding Lagrange multiplier $\lambda_{\theta_k}$ increases sufficiently large, due to the constraint violation, $\estwth_k$ approaches to the origin, as the invariant set $\Theta_{\estwth_k}^{1}$ shrinks to a point.
The remaining control input constraints can be validated implicitly through the term ${\gamma}_2$ in \eqref{eq:lya2:dot}.
Similarly to $c_{\theta_k}$, when a control input constraint $c_j,\ \forall j\in\mathcal{I}\setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}$ is violated, the corresponding Lagrange multiplier $\lambda_j$ increases sufficiently large, leading to increase in ${\gamma}_2$.
It makes ${\gamma}_2$ dominates the right-hand side of \eqref{eq:lya2:dot}, which leads $\ddtt V_2$ negative definite.
This drives $\estwth_k$ to the origin until the constraint is satisfied.
By Assumption \ref{assum:convex} and Lemma \ref{lem:convex:angle}, the control input constraints can be considered as convex constraint in the $\estwth_k$-space.

Therefore, the boundedness of $\estwth_k$ and $\fe$ is guaranteed by the invariant sets $\Theta_{\fe}^{1}$ and $\Theta_{\theta_k}^{1}$, and the satisfactions of $c_j,\ \forall j\in\mathcal{I}$ is ensured.

\subsection*{Case 2: Control input is not saturated.}

Since, the control input constraints are inactive, the saturation function $\mysat(\cdot)$ in \eqref{eq:lya1:dot2} and the constraint functions $c_j,\ \forall j\in\mathcal I \setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}$ in \eqref{eq:lya2:dot} can be removed.
Consider the Lyapunov function candidate $V_3:=V_1+V_2$, whose time derivative is given by:
\begin{equation}
    \begin{aligned}
        \ddtt V_3
        =&
        -\lambda_{\min}(\mm K)\norm{\fe}^2
        +\overline\tau_d\norm{\fe}
        +\fe^\top (\estNN - \idealNN - \mv\epsilon)
        \\
        &
        -\estwth_k^\top (
            (\mm I_{l_{k+1}}\otimes \estact_k^\top)^\top
            \fe
            +
            \lambda_{\theta_k}\estwth_k
        )
        \\
        \le&
        -\lambda_{\min}(\mm K)\norm{\fe}^2
        +\fe^\top \estNN
        +(
            \overline\tau_d+\norm{
                \idealNN+\mv\epsilon
            }
        )
        \norm{\fe}
        \\
        &
        -\estNN^\top\fe 
        -\lambda_{\theta_k}\norm{\estwth_k}^2
        \\
        \le&
        -
        \lambda_{\min}(\mm K)
        \norm{\fe}^2
        +
        \left(
            \overline\tau_d
            +
            \bar{\theta}_k\sqrt{l_{k}+1}
            +
            \bar{\epsilon}
        \right)
        \norm{\fe}
        \\
        &
        -
        \lambda_{\theta_k}\norm{\estwth_k}^2
        .
    \end{aligned}
    \label{eq:lya3:dot}
\end{equation}
Following the same reasoning as in Case 1, the boundedness of $\fe$ is ensured by the invariant set:
\begin{equation}
    \Theta_{\fe}^{2}
    =
    \left\{ 
        \fe \in \R^n 
        \mid 
        \norm{\fe} 
        \le 
        \tfrac{
            \overline\tau_d
            +
            \bar{\theta}_k\sqrt{l_{k}+1}
            +
            \bar{\epsilon}
        }{
            \lambda_{\min}(\mm K)
        }
    \right\}
    .
\end{equation}
The boundedness of $\estwth_k$ also can be ensured by the invariant set $\Theta_{\theta_k}^{2}=\{ \estwth_k \in \R^{\Xi_{k}} \mid \norm{\estwth_k} \le \overline\theta_k \}$, as $\lambda_{\theta_k}$ remains positive unless the weight norm constraint $c_{\theta_k}$ is satisfied.

\hfill

The boundedness of the inner-layer weights and the satisfaction of their corresponding constraints $c_j,\ \forall j\in\{\theta_i\}_{i\in\{k-1,\cdots,0\}}$, can be established recursively using \cite[Chap.~4 Thm.~1.9]{Desoer:2009aa}.
The dynamics of $\estwth_i,\ \forall i\in\{k-1,\cdots,0\}$, are represented as
\begin{equation}
    \ddtt \estwth_i 
    =
    -\alpha
    \left(
        \pptfrac{\estNN}{\estwth_i}^\top
        \fe
        +
        \lambda_{\theta_i}\estwth_i
        +
        \textstyle\sum_{
            j\in\mathcal{I}\setminus \{\wth_i\}_{i\in\{0,\cdots,k\}}
        }
        \lambda_j
        \pptfrac{c_j}{\estwth_i}
    \right)
    .
\end{equation}
By Lemma \ref{lem:cstr:grad:bound} $\estwth_i$ is bounded, provided that $\estwth_i,\ \forall i\in\{k,\cdots,i+1\}$, are bounded, since the system matrix $-\lambda_{\theta_i} \mm I_{\Xi_i}$ is stable and the residual terms are bounded.
Therefore, starting from the $(k-1)\textsuperscript{th}$ layer, the boundedness of $\estwth_i$ can be established recursively down to the input layer ($i=0$).

\hfill

In conclusion, the filtered tracking error $\fe$ and the entire set of estimated weights $\estwth$ are bounded, and all imposed constraints $c_j,\ \forall j\in\mathcal I$ are satisfied.

\end{proof}

%  SECTION SIMULATION ======================================
\section{Real-Time Implementation and Validation}\label{sec:sim}

\subsection{Validation Setup}

% ------------------------------------
% Real-time Implementation Set-up
% ------------------------------------
\begin{figure}[t]
    \centering
        \includegraphics[width=\figSizeOneCol\linewidth]
        {
            src/figures/exp_set.drawio.png
        }%
    \caption{
        Experimental setup of the two-link robotic manipulator used for real-time validation.
    }
    \label{fig:ctrl:exp:set}
\end{figure}

To validate the effectiveness of the proposed CONAC, a real-time experiment was conducted on a two-link robotic manipulator, as shown in Fig.~\ref{fig:ctrl:exp:set}. 
The OpenCR1.0 board \cite{opencr} with a 32-bit ARM Cortex and a 216 MHz clock frequency was used for the real-time implementation.
The two-link robotic manipulator actuated by two Cubemars AK10-90 servo motors. 
The servos were controlled using the OpenCR1.0 board which transmits the reference angle to the corresponding servo via controller area network (CAN) communication.
The system dynamics of the two-link robotic manipulator can follow the Euler-Lagrange equation \eqref{eq:sys1} and are illustrated in Fig.~\ref{fig:robot:model}. 
The measured and/or estimated system parameters are provided in Table~\ref{table:system:params}.

% ------------------------------------
% Robot Model Figure and Table
% ------------------------------------
\begin{figure}[t]
    \centering
    \subfloat[Two-link robotic manipulator.]{
        \includegraphics[width=0.65\linewidth]{src/figures/RobotModel.drawio.pdf}
        \label{fig:robot:model}
    }
    \hfill
    \subfloat[Control input saturation function.]{
        \includegraphics[width=0.55\linewidth]{src/figures/saturation.drawio.pdf}
        \label{fig:robot:sat}
    }
    \caption{Two-link robotic manipulator and control saturation function.}
    \label{fig:robot}
\end{figure}

% ------------------------------------
% Robot Model Parameters
% -----------------------------------
\begin{table}[t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Two-link robotic manipulator's parameters.}
    \centering
    \begin{tabular}{c m{9.5em} c c c }
    % \begin{tabular}{c c c c c}
    \hline
    \textbf{Symbol} & \textbf{Description} & \textbf{Value} \\
    \hline
    \hline 
    $m_1,m_2$ & Mass & 2.465 kg \\
    \hline
    $l_1,l_2$  & Length & 0.2 m \\  
    \hline
    ${l_c}_1,{l_c}_2$ & Center of mass & 0.139 m \\
    \hline
    $I_1,I_2$  & Inertia & 0.069 kgm\textsuperscript{2} \\
    \hline
    \end{tabular}
    \label{table:system:params}
\end{table}

% \begin{table}[t]
%     \renewcommand{\arraystretch}{1.3}
%     \caption{Two-link manipulator parameters of $p\textsuperscript{th}$ link.}
%     \centering
%     \begin{tabular}{c m{9.5em} c c c }
%     % \begin{tabular}{c c c c c}
%     \hline
%     \textbf{Symbol} & \textbf{Description} & \textbf{Link 1} & \textbf{Link 2} \\
%     \hline
%     \hline 
%     $m_p$ & Mass & 2.465 (kg) & 2.465 (kg) \\
%     \hline
%     $l_p$  & Length & 0.2 (m) & 0.18 (m) \\
%     \hline
%     ${l_c}_p$ & Center of mass & 0.139 (m) & 0.087 (m) \\
%     \hline
%     $b_p$   & Viscous friction coef. &  0.1 (Nms) & 0.1 (Nms) \\
%     \hline
%     ${f_c}_p$  & Coulomb friction coef. &  0.24 (Nm) & 0.24 (Nm) \\
%     \hline
%     ${f_s}_p$  & Static friction coef. &  0.3 (Nm) & 0.2 (Nm) \\
%     \hline
%     $I_p$  & Inertia & 0.069 (kgm\textsuperscript{2}) & 0.015 (kgm\textsuperscript{2}) \\
%     \hline
%     \end{tabular}
%     \label{table:system:params}
% \end{table}

% ------------------------------------
% Reference Trajectory
% ------------------------------------
The desired trajectory $\qd=\left({q_d}_1,{q_d}_2\right)^\top $ was generated using quintic (5$\textsuperscript{th}$-order) polynomial functions $\mathcal{P}^5(t;\mv{x}_1,\mv{x}_2,t_f)$, which satisfy the initial and final joint angles $\mv{x}_1$ and $\mv{x}_2$, as well as zero initial and final joint velocities. 
Here, $t_f$ denotes the duration of each segment.
The desired (reference) trajectory is defined as follows:
\begin{equation}
    \qd(t) 
    =
    \begin{pmatrix}
        {q_d}_1
        \\
        {q_d}_2
    \end{pmatrix}
    =
    \begin{cases}
    \mathcal{P}^5(t;\mv{q}_d^{(0)},\mv{q}_d^{(1)},t_f),
        &
        \text{if } 0 \le t < t_f 
        ,
        \\
    \mathcal{P}^5(t;\mv{q}_d^{(1)},\mv{q}_d^{(2)},t_f),
        &
        \text{if } t_f \le t < 2t_f 
        ,
        \\
    \mathcal{P}^5(t;\mv{q}_d^{(2)},\mv{q}_d^{(1)},t_f),
        & 
        \text{if } 2t_f \le t < 3t_f 
        ,
        \\
    \mathcal{P}^5(t;\mv{q}_d^{(1)},\mv{q}_d^{(0)},t_f),
        &    
        \text{if } 3t_f \le t < 4t_f 
        ,
        \\
    \end{cases}
    \label{eq:desired trajectory}
\end{equation}
where 
$
    \mv{q}_d^{(0)}:=\left(-\tfrac{1}{3}\pi,\tfrac{1}{3}\pi\right)^\top
$, 
$
    \mv{q}_d^{(1)}:=\left(\tfrac{1}{4}\pi,-\tfrac{1}{2}\pi\right)^\top
$, 
$
    \mv{q}_d^{(2)}:=\left(-\tfrac{1}{4}\pi,\tfrac{1}{4}\pi\right)^\top
$, 
and $t_f=3$ s, and the reference trajectory is illustrated in Fig.~\ref{fig:robot:model}.
To demonstrate the learning capability of the DNN, the desired trajectory was applied twice. 
The first and second repetitions of the reference trajectory are referred to as Episode $1$ and Episode $2$, respectively.

However, due to the random initialization of the DNN weights, large reference trajectories should be avoided to reduce the impact of the controller's initial poor performance.
Therefore, prior to applying the reference trajectory, the manipulator was first positioned at a stable equilibrium point, denoted as ${\qd}^{\rm{origin}} = \left(-\tfrac{1}{2}\pi, 0\right)^\top$.
Subsequently, a trajectory $\mathcal{P}^5(t; {\qd}^{\rm{origin}}, {\qd}_0, 3)$ was used to move the manipulator from ${\qd}^{\rm{origin}}$ to ${\qd}_0$ over a duration of 3 seconds, defining the start and end positions of each episode.
Once the manipulator reached ${\qd}_0$, a constant reference trajectory $\qd = {\qd}_0$ was maintained for 8 seconds to allow the system to settle and eliminate the influence of transient dynamics.

% ------------------------------------
% Control Input Saturation
% ------------------------------------
The effectiveness of the proposed CONAC in handling combined convex input constraints was demonstrated by introducing a saturation function that projects the control input onto the convex set $\Theta_{\mysat} :=\left\{ \Theta_{\mysat, {\tau_2}} \cap \Theta_{\mysat, {\rbu}} \right\}$, as illustrated in Fig.~\ref{fig:robot:sat}, where
\begin{subequations}
    \begin{align}
        \Theta_{\mysat, {\tau_2}}
        :=&
        \left\{
            \tau_2
            \mid
            \vert{\tau_2}\vert \le 
            3.5
            % \overline\tau_2
        \right\}
        ,
        \label{eq:sat:tau2}
        \\    
        \Theta_{\mysat, {\rbu}}
        :=&
        \left\{
            \rbu
            \mid
            \norm{\rbu} \le 
            11
            % \overline\tau
        \right\}
        .
        \label{eq:sat:norm}
    \end{align}
    \label{eq:sat}% %%
\end{subequations}

The sets defined in \eqref{eq:sat:tau2} and \eqref{eq:sat:norm} can be physically interpreted as the current limit imposed by the torque limit of the second joint actuator and the power source, respectively.
The corresponding control input constraints are mathematically described in Appendix~\ref{sec:appen:cstr} as the nonlinear input norm constraint \eqref{eq:cstr:input:bound} and the input bound constraint \eqref{eq:cstr:input:norm}, respectively.

% ------------------------------------
% Comparative Study
% ------------------------------------

\hfill

Four controllers were implemented for comparative analysis, with their properties summarized in Table~\ref{table:controllers}.

% ------------------------------------
% Controllers: CONAC 
% ------------------------------------
\subsubsection*{Controller 1: \normalfont{(C$_1$)} \textit{CONAC with large update rate, $\beta_j,\forall j\in\mathcal{I}$, of Lagrange Multipliers}}

To handle input saturation, the proposed CONAC was implemented with the input bound constraints $c_{\overline\tau_2}$ and $c_{\underline\tau_2}$ \eqref{eq:cstr:input:bound} for the second link, and the nonlinear input norm constraint $c_{\rbu}$ \eqref{eq:cstr:input:norm}, as follows:
\begin{equation}
    c_{\overline\tau_2}     
    =
    \tau_2-\overline\tau_2
    % \le 0
    ,
    \;
    c_{\underline\tau_2} 
    =
    \underline\tau_2-\tau_2
    % \le 0
    ,
    \;
    c_{\rbu}
    =
    \tfrac{1}{2}
    \left(
        \norm{\rbu}-\overline\tau
    \right)^2 
    % <0
    ,
    \label{eq:sim:cstr:CONAC}
\end{equation}
where $\overline\tau_2=-\underline{\tau}_2=3.5$, and $\overline\tau=11$.
Therefore, controller (C$1$) utilizes the full capability of the actuator.
The weight norm constraints $c_{\theta_i}(\cdot),\ \forall i\in\{0,\cdots,k\}$ defined in \eqref{eq:cstr:weight:ball}, were also incorporated into the CONAC framework.
As described in \eqref{eq:control:est}, the control input (torque) corresponds to the output of the DNN, and the adaptation law is defined in \eqref{eq:adaptation law}.

The update rates of the Lagrange multipliers for the weight norm constraints were set to $\beta_{\theta_i}=1,\ \forall i\in\{0,\cdots,k\}$.
For the control input constraints, the update rates were assigned large values: $\beta_{\rbu} = 10$ and $\beta_{\overline{\tau}2} = \beta{\underline{\tau}_2} = 10^2$, in order to investigate the effect of large update rates.

% ------------------------------------
% Controllers: CONAC with small beta
% ------------------------------------
\subsubsection*{Controller 2: \normalfont{(C$_2$)} \textit{CONAC with small update rate, $\beta_j,\forall j\in\mathcal{I}$, of Lagrange Multipliers}}

As the second controller, the proposed CONAC was implemented with the same input bound constraints as (C$_1$), but with $100$ times smaller update rate values: $\beta_{\rbu}=0.1$ and $\beta_{\overline{\tau}_2}=\beta_{\underline{\tau}_2}=1$.
All other properties and parameters were kept identical to those of (C$_1$).

% ------------------------------------
% Controllers: CONAC without ctrl cstr
% ------------------------------------
\subsubsection*{Controller 3: \normalfont{(C$_3$)} \textit{CONAC without control input constraints}}

The third controller was identical to (C$1$) and (C$2$), except that the control input constraints were removed to highlight the constraint handling capability of the proposed CONAC.
In this case, the update rates of the Lagrange multipliers associated with the control input constraints, \ie $\beta_{\rbu},\beta_{\overline{\tau}_2},\beta_{\underline{\tau}_2}$, were set to zero, such that the corresponding multipliers $[\lambda_j]_{j\in\{\rbu,\overline{\tau}_2,\underline{\tau}_2\}}$ were not updated (see \eqref{eq:adap:lbd}).
All other properties and parameters were kept the same as in (C$_1$) and (C$_2$).

% ------------------------------------
% Controllers: Aux. system in CONAC-AUX
% ------------------------------------
\subsubsection*{Controller 4: \normalfont{(C$_4$)} \textit{CONAC with auxiliary system instead of control input constraints}}

As a baseline for comparison, the existing auxiliary system approach \cite{Esfandiari:2014aa, Karason:1994aa, Esfandiari:2015aa} was implemented in place of the proposed CONAC for handling control input constraints.
The auxiliary system is defined by $\ddtt{\mv\zeta} = \mm{A}_\zeta \mv{\zeta} + \mm{B}_\zeta \Delta\rbu$, with initial condition $\mv{\zeta}\vert_{t=0} = \mv{0}_{2\times1}$, where $\mv{\zeta}\in\R^2$ represents the auxiliary state variables, and $\mm{A}_\zeta\in\R_{<0}^{2\times 2}$ and $\mm{B}_\zeta\in\R^{2\times 2}$ are user-designed matrices.
The auxiliary state $\mv{\zeta}$ is driven by the overflow amount of the control input $\Delta\rbu$, where each element is defined as $\Delta\tau_{i} = \tau_{i}-\max(\underline\tau_i,\min(\overline\tau_i,\tau_i)),\ \forall i\in\{1,2\}$..

This auxiliary state was incorporated into the adaptation law \eqref{eq:adap:th:approx} by replacing the filtered tracking error $\fe$ with $\fe+\mv{\zeta}$.
As a result, the weights were adapted to minimize both the filtered tracking error and the auxiliary state, thereby resolving the control input constraint saturation.

The matrices for the auxiliary system were chosen as $\mm{A}_\zeta=\mydiag\left((-10,-10)^\top\right)$ and $\mm{B}_\zeta=\mydiag\left((10^3,10^3)^\top\right)$, reflecting a trade-off between sensitivity to input saturation and convergence speed of the auxiliary dynamics.
Since the auxiliary system only accounts for individual input saturation—corresponding to the input bound constraint \eqref{eq:cstr:input:bound}—the upper limit for the first actuator torque, $\overline{\tau}_1$, was set to $10.428$ to approximately match the allowable control input norm $\overline{\tau}$, given $\overline{\tau}_2 = 3.5$ and $\overline{\tau} = 11$, satisfying $\overline{\tau}_1^2 + \overline{\tau}_2^2 = \overline{\tau}^2$.

\hfill

% ------------------------------------
% Controller Parameters
% ------------------------------------
For all controllers, \ie (C$_1$), (C$_2$), (C$_3$), and (C$_4$), the input vector to the DNN was defined as ${\q}_n=\left(\q^\top,\dq^\top,\fe^\top,1\right)^\top\in\R^{7}$, where the scalar $1$ was included to account for the bias term in the weight matrix.
All DNNs shared the same architecture, consisting of two hidden layers with four nodes each, \ie $k=2$ and $l_0=6, l_1=4, l_2=4$, and $l_3=2$ in \eqref{eq:DNN:architecture}.
The adaptation gain was set to $\alpha = 0.2$, and the filtering matrix in \eqref{eq:err:filtered} was chosen as $\mm{\Lambda}=\mydiag([5,15])$.
The DNN weights were initialized using a uniform distribution in the range $[-0.1, 0.1]$, with the same random seed used across all experiments to ensure consistency.
The controller sampling rate was set to $250$ Hz to support real-time implementation.
Weight norm constraint parameters were selected as $\overline\theta_0=\overline\theta_1=\overline\theta_2=6$.

Control results were collected via CAN communication using a PCAN-USB interface and recorded on a computer at a sampling rate of $250$ Hz.

The tracking performance was quantitatively evaluated using the $L_2$-norm of the tracking error, defined as $\sqrt{\int_0^T\norm{h}\,\der t}$, where $h\in\{e_1,e_2\}$ and $T\in\R_{>0}$ denotes the duration of each episode.

% ------------------------------------
% Controller Table
% ------------------------------------
\begin{table}[t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Properties of the controllers}
    \label{table:controller}
    \centering
    \begin{tabular}{c l l}
    \hline
        &\bf{Description}&\bf{Input Constraint Handling}\\
    \hline
    \hline
        \multirow{2}{*}{(C$_1$)}&large&\eqref{eq:cstr:input:norm} $c_{\rbu}$ with $\overline{\tau}=11$ \\
        &  update rate $\beta_j$ &\eqref{eq:cstr:input:bound} $c_{\overline\tau_2}, c_{\underline\tau_2}$ with $\overline{\tau}_2=-\underline{\tau}_2=3.5$ \\
    \hline
        \multirow{2}{*}{(C$_2$)}&small&\eqref{eq:cstr:input:norm} $c_{\rbu}$ with $\overline{\tau}=11$ \\
        &  update rate $\beta_j$ &\eqref{eq:cstr:input:bound} $c_{\overline\tau_2}, c_{\underline\tau_2}$ with $\overline{\tau}_2=-\underline{\tau}_2=3.5$ \\
    \hline
        \multirow{2}{*}{(C$_3$)}&no input& \multirow{2}{*}{$[\beta_j]_{j\in\{\rbu,\overline{\tau}_2,\underline{\tau}_2\}}=\mv{0}$} \\
        & constraint &  \\
    \hline
        \multirow{2}{*}{(C$_4$)}&\multirow{2}{*}{auxiliary system} & $\overline{\tau}_1=-\underline{\tau}_1=10.428$ \\
        &  & $\overline{\tau_2}=-\underline{\tau_2}=3.5$ \\    
    \hline
    \end{tabular}
    \label{table:controllers}
\end{table}

\begin{figure*}[t]
  \centering
    \subfloat[Tracking result of $q_1$.]{
    %   \includegraphics
      \includegraphics[width=\figSizeTwoCol\linewidth]
      {
        % fig/Fig1.pdf
        src/measurement/figures/compare/Fig1.eps
      }%
      \label{fig:ctrl:result:q1}}
    \hfill
    \subfloat[Tracking result of $q_2$.]{
      \includegraphics[width=\figSizeTwoCol\linewidth]
      {
        % src/measurement/figures/compare/Fig11.eps
        src/measurement/figures/compare/Fig2.eps
      }%
      \label{fig:ctrl:result:q2}}
    \vfill
      \subfloat[Control input $\tau_1$.]{
        \includegraphics[width=\figSizeTwoCol\linewidth]
        {
            src/measurement/figures/compare/Fig3.eps
        }%
        \label{fig:ctrl:result:tau:1}}
    \hfill
      \subfloat[Control input $\tau_2$.]{
        \includegraphics[width=\figSizeTwoCol\linewidth]
        {
            src/measurement/figures/compare/Fig4.eps
        }%
        \label{fig:ctrl:result:tau:2}}
        \vfill
    \subfloat[Norm of control input $\rbu$.]{
        \includegraphics[width=\figSizeTwoCol\linewidth]
        {
            src/measurement/figures/compare/Fig5.eps
        }%
        \label{fig:ctrl:result:tau:norm}}
    \hfill
    \subfloat[Norm of weights ${\estwth}_i$.]{
        % \includegraphics
        \includegraphics[width=\figSizeTwoCol\linewidth]
        {
            % fig/Fig7.pdf
            src/measurement/figures/compare/Fig7.eps
        }%
        \label{fig:ctrl:result:weight}}        
    \caption{
    Real-time implementation results of (C$_1$) \protect\colorLine{blue}{solid}, (C$_2$) \protect\colorLine{cyan}{solid}, (C$_3$) \protect\colorLine{gray}{solid}, (C$_4$) \protect\colorLine{magenta}{solid}, and reference signal $\qd$ \protect\colorLine{red}{dashed}.
  }
\label{fig:ctrl:result}
\end{figure*}

% ------------------------------------
% Norm of error in Episode 1: 
% C1 e1 ep1: 25.130
% C1 e2 ep1: 25.016
% C2 e1 ep1: 21.719
% C2 e2 ep1: 10.530
% C3 e1 ep1: 23.168
% C3 e2 ep1: 13.061
% C4 e1 ep1: 24.195
% C4 e2 ep1: 12.927
% Norm of error in Episode 2: 
% C1 e1 ep2: 7.066
% C1 e2 ep2: 2.198
% C2 e1 ep2: 8.021
% C2 e2 ep2: 2.435
% C3 e1 ep2: 8.408
% C3 e2 ep2: 3.554
% C4 e1 ep2: 19.721
% C4 e2 ep2: 4.256
% Improvement in Episode 2: 
% C1 e1: 0.719
% C1 e2: 0.912
% C2 e1: 0.631
% C2 e2: 0.769
% C3 e1: 0.637
% C3 e2: 0.728
% C4 e1: 0.185
% C4 e2: 0.671
% ------------------------------------

\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Quantitative\,comparison\,of\,performances'\,$L_2$\,norm.}
    \centering
    \begin{tabular}{c c c c c}
    \hline
		& \multicolumn{4}{c}{\textbf{Episode 1}} \\
    \hline
	\hline 
		& (C$_1$) & (C$_2$) & (C$_3$) & (C$_4$) \\
	\hline
		$e_1\times10^{3}$ / rad & $25.130$ & $21.719$ & $23.168$ & $24.195$ \\ 
	\hline
        $e_2\times10^{3}$ / rad & $25.016$ & $10.530$ & $13.061$ & $12.927$ \\
	\hline
        & \multicolumn{4}{c}{\textbf{Episode 2}} \\
    \hline
    \hline
        & (C$_1$) & (C$_2$) & (C$_3$) & (C$_4$) \\
	\hline
    \multirow{2}{*}{$e_1\times10^{3}$ / rad} 
        & $7.066$ & $8.021$ & $8.408$ & $19.721$ \\
        & ($-71.6\%$) & ($-63.1\%$) & ($-63.7\%$) & ($-18.5\%$) \\
    \hline
    \multirow{2}{*}{$e_2\times10^{3}$ / rad} 
        & $2.198$ & $2.435$ & $3.554$ & $4.256$ \\
        & ($-88.4\%$) & ($-76.9\%$) & ($-72.8\%$) & ($-67.1\%$) \\
    \hline
    \end{tabular}
    \label{tab:sim:L2}
\end{table}

\subsection{Validation Results}

% ------------------------------------
% Tracking Performance
% ------------------------------------
\subsubsection{Tracking Performance}

The real-time implementation results are presented in Fig.\ref{fig:ctrl:result}, and the quantitative comparison of tracking performance is summarized in Table\ref{tab:sim:L2}.
Across the two episodes, all controllers improved their tracking performance. 
Specifically, the $L_2$-norm of $e_1$ was reduced by $71.6 \%$ for (C$_1$), $63.1 \%$ for (C$_2$), $63.7 \%$ for (C$_3$), and $18.5 \%$ for (C$_4$); while the $L_2$-norm of $e_2$ was reduced by $88.4 \%$, $76.9 \%$, $72.8 \%$, and $67.1 \%$, respectively.
Qualitatively, as illustrated in Fig.\ref{fig:ctrl:result:q1} and Fig.\ref{fig:ctrl:result:q2}, oscillations were suppressed during the second episode.
These results indicate that the DNNs in all controllers successfully learned the ideal control input $\rbu^*$ using the filtered tracking error $\fe$, despite $\rbu^*$ being completely unknown a priori.
Furthermore, the stable learning achieved under random weight initialization and without prior knowledge of the system demonstrates the effectiveness of the proposed CONAC in enabling online learning.

Notably, in both episodes, all controllers failed to track the desired trajectory of $q_1$ during certain time intervals: from $17.5$ s to $19.9$ s in Episode $1$, and from $23.5$ s to $25.9$ s and from $29.5$ s to $31.9$ s in Episode $2$.
This degradation in tracking performance is attributed to actuator saturation during those intervals, as observed in Fig.\ref{fig:ctrl:result:tau:1} and Fig.\ref{fig:ctrl:result:tau:norm}.

The tracking performance of CONAC with control input constraints—\ie (C$_1$) and (C$_2$)—outperformed the auxiliary system-based controller (C$_4$) in the second episode, as shown in Table~\ref{tab:sim:L2}.
This improvement is due to the fact that (C$_1$) and (C$_2$) were able to fully utilize the actuator's capabilities, whereas (C$_4$) was constrained by the auxiliary system's design, which limited the first joint actuator.
While (C$_3$) also showed better performance than (C$_4$), this was primarily because it did not consider input saturation and thus used more of the actuator's capacity.
\KHCH{
        여기부터 문단 끝날때까지의 표현이 명료하지 않습니다. C3의 tracking performance가 좋은 것은 지표상으로 명확히 맞는 말이기 때문에 마치 tracking performance가 실제로 좋은 것은 아니다라는 뉘앙스의 논리 전개가 적절치 않습니다.
    Tracking performanc는 더 좋지만 과도한 control input을 사용하고, saturation이 지난 이후에는 진동 문제가 있다는 것만 명료하게 표현해주세요.
}
One may conclude that (C$_3$) outperformed (C$_4$) since the tracking performance of (C$_3$) was better than (C$_4$), as shown in Table~\ref{tab:sim:L2}.
\KHCH{어색한 표현; 수정 요}\MSRY{필요없는 내용으로 판단하여 삭제했습니다.}
However, during the episodes, (C$_3$) increases the control input excessively large to reduce the tracking error, even though it can not be regulated due to the actuator limits as shown in Fig.~\ref{fig:ctrl:result:tau:2} and Fig.~\ref{fig:ctrl:result:tau:norm}, \ie the violations of the control input saturations of (C$_3$) grow larger in every saturated time interval.
Moreover, this large violation occurs oscillations when the control saturation is released as the reference trajectory is changed, since the weights are adapted to produce large control inputs.

% ------------------------------------
% Control Input Saturation
% ------------------------------------
\begin{figure}[t]
    \centering
    \subfloat[Control input $\tau$ locus in time interval from $29.4$ s to $31.9$ s.]{
    % \subfloat[Auxiliary states $\mv{\zeta}$ of CONAC-AUX.]{
        \includegraphics[width=\figSizeOneCol\linewidth]
        {
            src/measurement/figures/compare/Fig6.eps
        }%
        \label{fig:ctrl:result:scope:control}}
        \vfill
    \subfloat[Lagrange multipliers $\lambda_j$  in time interval from $29.4$ s to $31.9$ s.]{
        \includegraphics[width=\figSizeOneCol\linewidth]
        {
            src/measurement/figures/compare/Fig8.eps
        }%
        \label{fig:ctrl:result:scope:beta}}
        \vfill
    \subfloat[Auxiliary states in time interval from $29.4$ s to $31.9$ s.]{
        \includegraphics[width=\figSizeOneCol\linewidth]
        {
            src/measurement/figures/compare/Fig9.eps
        }%
        \label{fig:ctrl:result:scope:aux}}
  \caption{
    Constraint handling process of (C$_1$) \protect\colorLine{blue}{solid}, (C$_2$) \protect\colorLine{cyan}{solid}, (C$_3$) \protect\colorLine{gray}{solid}, and (C$_4$) \protect\colorLine{magenta}{solid} in time interval from $29.4$ s to $31.9$ s.
  }
\label{fig:ctrl:result:scope}
\end{figure}

\begin{figure}[!t]
	\centering
	\includegraphics[width=0.6\linewidth]{
		src/figures/lambda_effect.drawio.pdf
		}
	\caption{
        The effect of the Lagrange multiplier $\lambda_{j}$ on the adaptation direction of $\widehat{\mv{\theta}}_2$ is illustrated. 
        The notations $(\cdot)'$ and $(\cdot)''$ denote two distinct cases corresponding to large and small values of the Lagrange multiplier, respectively, \ie $\lambda_{j}' > \lambda_{j}''$.
	}
	% \vspace{1mm}
	\label{fig:lambda_effect}
\end{figure}

\hfill

\subsubsection{Constraint Handling} \label{sec:sim:constraint}

In this section, the constraint handling capability of all controllers are investigated.
As shown in Fig.~\ref{fig:ctrl:result:tau:1}, Fig.~\ref{fig:ctrl:result:tau:2}, and Fig.~\ref{fig:ctrl:result:tau:norm}, all controllers except (C$_3$) which did not consider the control input saturation, satisfied the control input saturation illustrated in Fig.~\ref{fig:robot:sat}.
For (C$_1$) and (C$_2$), the control input saturation was handled using the Lagrange multipliers and, for (C$_4$), the auxiliary system was used to handle the control input saturation.

\KHCH{
    여기에 각 제어기의 constraint를 표현해주는게 좋겠습니다. FIg6a만 보고 아래의 설명을 이해하기에는 리뷰어들에게 어려움이 있을 겁니다.
}
\MSRY{
    아래의 설명을 보완하여 각 제어기의 constraint를 표현해주었습니다.
}
Even though all controllers except (C$_3$) satisfied the input saturation, (C$_4$) was unable to fully utilize the actuator's capability.
As shown in Fig.~\ref{fig:ctrl:result:scope:control}, which illustrates the control input locus during the time interval from $29.4$ s to $31.9$ s, the control inputs of (C$_1$) and (C$_2$) were constrained by the combined convex input constraints, which precisely matched the feasible domain $\Theta_{\mysat}$ depicted in Fig.~\ref{fig:robot:sat}
In contrast, (C$_4$)'s control input was restricted independently by each component-wise constraint, limiting its ability to explore the full admissible input space, owing to the decoupled formulation of the auxiliary system.

\hfill

% zeta
The detail of the constraint handling during the time interval from $29.4$ s to $31.9$ s, is now investigated.
For (C$_4$), the auxiliary state $\zeta_1$ was generated in response to the control input saturation of the first link actuator, as shown in Fig.~\ref{fig:ctrl:result:scope:aux}.
Since the weights were adapted to reduce both the filtered error $\fe$ and the auxiliary state $\mv{\zeta}$, the control input saturation was subsequently reduced, as evident in Fig.\ref{fig:ctrl:result:scope:control}.
Once the saturation was resolved at approximately $31$ seconds (see, Fig.~\ref{fig:ctrl:result:tau:norm}), the auxiliary state $\zeta_1$ converged to zero due to the stable state matrix $\mm{A}_\zeta$ in the auxiliary dynamics, given by $\ddtt\mv{\zeta} = \mm{A}_\zeta \mv{\zeta} + \mm{B}_\zeta \Delta\rbu$ with $\Delta\rbu=\mv{0}_{2\times1}$, as shown again in Fig.~\ref{fig:ctrl:result:scope:aux}.

% beta + zeta
In the cases of (C$_1$) and (C$_2$), the Lagrange multipliers $\lambda_j\in\forall j\in\mathcal{I}$ directed the adaptation of the weights $\estwth_i,\ \forall i\in\{0,1,2\}$, to reduce violations of the constraints $c_j$. 
The multipliers $\lambda_j$ were generated dynamically in response to constraint violations.
% comparison; big, small beta
The effect of Lagrange multiplier $\lambda_j$ on the adaptation direction of $\estwth_2$ is illustrated in Fig.~\ref{fig:lambda_effect}, where two different magnitudes of $\lambda_j$ are considered. 
Notably, under Assumptions \ref{assum:convex} and \ref{assum:LICQ}, the constraint $c_j$ is convex in the $\estwth_2$-space (see, Lemma~\ref{lem:convex:angle}).
Large and small Lagrange multipliers correspond to the configurations of (C$_1$) and (C$_2$), respectively, because the update rates in (C$_1$) were set $10^2$ times larger than those in (C$_2$).
Consequently, in (C$_1$), $\lambda_j$ increases more rapidly in response to constraint violations (see, \eqref{eq:adap:lbd}). 
This results in a more aggressive adaptation of $\estwth_2$, leading to faster constraint satisfaction—but at the cost of oscillatory behavior in the learning process.

Following this reasoning, the real-time implementation results of (C$_1$) and (C$_2$) are compared in Fig.~\ref{fig:ctrl:result:scope:control} and Fig.~\ref{fig:ctrl:result:scope:beta}. 
In these figures, $\lambda_{\rbu}$ for (C$_1$) repeatedly grows and resets, indicating recurring violations and satisfactions of the $c_{\rbu}$ constraint in discrete time.
On the other hand, (C$_2$), with smaller $\lambda_{\rbu}$, updates $\estwth_2$ more conservatively. 
This slower response leads to delayed constraint satisfaction, which is reflected by the persistent positive value of $\lambda_{\rbu}$ in Fig.\ref{fig:ctrl:result:scope:beta}. 
However, this allows (C$_2$) to achieve smoother control input behavior, as shown in Fig.~\ref{fig:ctrl:result:scope:control}.

\hfill

Finally, the handling of the weight norm constraints $c_{\theta_i},\ \forall i\in\{0,1,2\}$, is investigated. 
Among all controllers, the weight norm constraint was only activated for the outer layer weights $c_{\theta_2}$ in (C$_3$), as shown in Fig.~\ref{fig:ctrl:result:weight}. 
This is because the weights in (C$_1$), (C$_2$), and (C$_4$) were implicitly suppressed by the control input constraints, whereas (C$_3$) did not incorporate such constraints.
% , \ie by limiting the control input, the weight norms were also limited.
% \MSRY{필요하다면 왜 제어입력 제약하는게 가중치 제약도 되는지 추가 가능}
Furthermore, the constraints on the inner and hidden layer weights were never violated; that is, $\norm{\estwth_0}$ and $\norm{\estwth_1}$ consistently remained below $\norm{\estwth_2}$. 
This is attributed to the fact that the backpropagated error was scaled by the gradient of the activation function, which is bounded between $-1$ and $1$; see, \eqref{eq:gradient:phi}.

Additional details on the constraint handling process are provided in Fig.~\ref{fig:ctrl:result:scope}. 
In particular, Fig.~\ref{fig:ctrl:result:scope:beta} shows that the Lagrange multipliers $\lambda_{\theta_2}$ increased when $\estwth_2$ exceeded the prescribed norm constraint $\overline\theta_2$, prompting an adjustment in the adaptation direction of $\estwth_2$ to mitigate the violation of the constraint $c_{\theta_2}$.

\KHCH{
    이부분은 "worth to note"하지는 않아보입니다. 그리고 FIg 5에서 이 것을 확인하기도 어렵습니다. 굳이 이내용을 유지하려면 Fig 6에서 multiplier가 발생했다 사리지는 현상을 표현할 때 사용해주세요.
}
\MSRY{
    부드럽게 제약조건만족이 안되는 것에 대한 변명(?)으로 포함하고 싶었습니다. 
    그렇기에 제약조건 파트 제일 마직막으로 뺐습니다.
}

In addition, it is worth noting that the control input constraints were repeatedly violated even after the constraint handling mechanisms were applied, as shown in Fig.~\ref{fig:ctrl:result:scope:control}. 
This occurred because the controllers were implemented in discrete time with a relatively low sampling rate of $250$ Hz.

% ------------------------------------
% Karsh-Kuhn-Tucker (KKT) conditions}
% ------------------------------------
\hfill

\subsubsection{Satisfaction of Karsh-Kuhn-Tucker (KKT) conditions}

Since the proposed CONAC was implemented in discrete time, a direct demonstration of the KKT conditions at steady state is not straightforward. 
However, the near-constant values of the weights and Lagrange multipliers during constraint violation suggest an implicit satisfaction of the KKT conditions. 
Specifically, the Lagrange multipliers corresponding to active constraints were positive, while those associated with inactive constraints remained zero.

% reference 자체가 계속 움직이니까 엄밀하게는 steady state를 보장할 수 없습니다. rough 하게는 theta norm이 어느정도 수렴하므로 이 것을 기반으로 주장해볼 수도 있습니다. theta_dot이 평균적으로는 0에 가까운 것 같서요

% ------------------------------------
% Computational Time
% ------------------------------------
\hfill

\subsubsection{Effectiveness of CONAC in Real-time Applications}

The effectiveness of the proposed CONAC in real-time applications was evaluated by measuring the computational time using CPU clock ticks on the OpenCR1.0 board. 
As shown in Fig.~\ref{fig:ctrl:real:result:cmp:time}, the computational times of all controllers remained below $4$ ms, demonstrating that the proposed CONAC is suitable for real-time implementation at a sampling rate of $250$ Hz.

\begin{figure}[t]
    \centering
        \includegraphics[width=\figSizeOneCol\linewidth]
        {
            src/measurement/figures/compare/Fig10.eps
        }%
    \caption{
        Computational time of (C$_1$) \protect\colorLine{blue}{solid}, (C$_2$) \protect\colorLine{cyan}{solid}, (C$_3$) \protect\colorLine{gray}{solid}, and (C$_4$) \protect\colorLine{magenta}{solid}.
    }
    \label{fig:ctrl:real:result:cmp:time}
  \end{figure}

%  SECTION CONCLUSION ======================================
\section{Conclusion}\label{sec:conclusion}

This paper proposed a constrained optimization-based neuro-adaptive controller (CONAC) for uncertain Euler-Lagrange systems, addressing both weight norm and input constraints through a rigorous optimization framework. 
The stability of the proposed controller was analyzed via Lyapunov theory, demonstrating bounded tracking and estimation errors under real-time adaptation.

The controller effectively incorporated both convex input constraints and weight norm constraints, ensuring that actuator limitations and neural network weights remained within predefined bounds. 
By embedding these constraints into the optimization process, CONAC ensured convergence of the weights in accordance with the Karush-Kuhn-Tucker (KKT) conditions, thereby achieving both stability and optimality.

A real-time implementation validated the superior performance of CONAC compared to conventional methods. 
The proposed controller successfully handled complex input constraints while rigorously managing weight norms, resulting in improved tracking accuracy and stable performance without significant oscillations.

Future work may extend this approach to include constraints on both control inputs and system states, further enhancing the flexibility and robustness of neuro-adaptive control systems within constrained optimization frameworks.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
\appendix

\subsection{Input Constraint Candidates}\label{sec:appen:cstr}

This section introduces potential weight and input constraints that can be incorporated into the proposed CONAC framework.

\subsubsection{Input Bound Constraint}\label{sec:appen:cstr:input:bound}

Most physical systems are subject to control input limitations due to inherent electrical and mechanical constraints.
These are expressed as $\mv{c}_{\overline \tau}:= [c_{\overline \tau_i}]_{i\in\{1,\cdots,n\}}$ and $\mv{c}_{\underline\tau}:= [c_{\underline\tau_i}]_{i\in\{1,\cdots,n\}}$, where
\begin{equation}
    \begin{aligned}
        c_{\overline \tau_i}=\tau_{(i)} - {\tau_{\overline \tau_i}}
        ,
        \quad
        c_{\underline\tau_i}={\tau_{\underline\tau_i}}-\tau_{(i)}
        ,
    \end{aligned}
    \label{eq:cstr:input:bound}
\end{equation}
with $\tau_{\overline \tau_i}\in\R_{>0}$ and $\tau_{\underline\tau_i}\in\R_{>0}$ representing the maximum and minimum control input bounds, respectively.
The gradients of $\mv{c}_{\overline \tau}$ and $\mv{c}_{\underline\tau}$ with respect to $\estwth$ are given by
\begin{equation}
    \begin{aligned}
        \pptfrac{\mv c_{\overline \tau}}{\estwth}
        =& 
        \left[
            \pptfrac{c_{\overline \tau_i}}{\estwth}^\top
        \right]_{i\in\{1,\cdots,n\}}
        % \begin{bmatrix}
        %     \pptfrac{c_{\overline \tau_1}}{\estwth}^\top \\
        %     \vdots \\
        %     \pptfrac{c_{\overline \tau_n}}{\estwth}^\top
        % \end{bmatrix}
        = 
            +\pptfrac{\estNN}{\estwth}
            \\
        =&
        =
        +
        \begin{bmatrix}
            (\mm{I}_{l_{k+1}}\otimes \estact_{k}^\top)&
            \!\cdots\! &
            (
                \cdot
            )
        \end{bmatrix} 
        \in
        \mathbb R^{n\times \Xi}
        , 
        \\
        \pptfrac{\mv c_{\underline\tau}}{\estwth}         
        =
        & 
        \left[
            \pptfrac{c_{\underline\tau_i}}{\estwth}^\top
        \right]_{i\in\{1,\cdots,n\}}
        % \begin{bmatrix}
        %     \pptfrac{c_{\underline\tau_1}}{\estwth}^\top \\
        %     \vdots \\
        %     \pptfrac{c_{\underline\tau_n}}{\estwth}^\top
        % \end{bmatrix}
        = 
        -\pptfrac{\estNN}{\estwth}
        \\
        =&
        =
        -\begin{bmatrix}
            (\mm{I}_{l_{k+1}}\otimes \estact_{k}^\top)&
            % V_k^\top \phi_{k}' (I_{l_{k}}\otimes  \phi_{k-1}^\top)&
            \!\cdots\! &
            (
                \cdot
            )
        \end{bmatrix} 
        \in
        \mathbb R^{n\times \Xi}
        .
    \end{aligned}
    \label{eq:cstr:input:bound:grad}
\end{equation}

\subsubsection{Input Norm Constraint}\label{sec:appen:cstr:input:norm}

Consider the control input $\rbu$ as the torque of each actuator corresponding to its generalized coordinate. 
Since torque is typically linearly proportional to current, actuators that share a common power source are often subject to total current limitations. This can be captured by the following inequality constraint: 
\begin{equation}
    c_{\rbu}
    =
    \norm{
        \mv{\tau}
    }^2 
    -
    \overline\tau^2
    ,
    \label{eq:cstr:input:norm}
\end{equation}
with $\overline\tau\in\R_{>0}$ denoting the maximum allowable control input magnitude. This input norm constraint is also commonly applied in current and torque control problems for electric motors \cite{Choi:2024aa}.
% The corresponding Lagrangian multiplier is $\lambda_{u_b}$. 
The gradients of $c_{\rbu}$ with respect to $\estwth$ are given by
\begin{equation}
    \pptfrac{\mv c_{\rbu}}{\estwth}
    \!=\! 
    \textstyle\sum_{i=1}^n 2\tau_{i} 
    \left(
        \myrow_i
        \left(
            \pptfrac{\estNN}{\estwth}
        \right)
    \right)^\top  
    \!=\! 
    {\pptfrac{\estNN}{\estwth}}^\top
    \mv{\tau}
    % \tau^\top (I_{l_{k+1}}\otimes \estNN_k^\top)
    \in \mathbb R^{\Xi}.
    \label{eq:cstr:input:norm:grad}
\end{equation}0
It should be noted that constraints \eqref{eq:cstr:input:bound} and \eqref{eq:cstr:input:norm} can be imposed simultaneously, as their gradients \eqref{eq:cstr:input:bound:grad} and \eqref{eq:cstr:input:norm:grad} are linearly independent, satisfying Assumption \ref{assum:LICQ}, \ie the LICQ condition.

\bibliographystyle{IEEEtran}
\bibliography{\template/refs}

\begin{IEEEbiography}[{
    \includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{\template/people/msRyu.jpg}}]{Myeongseok Ryu}
    received the B.S. degree in mechanical engineering from Incheon National University, South Korea, in 2023, and the M.S. degree in mechanical engineering from Gwangju Institute of Science and Technology (GIST), South Korea, in 2025. 
    He is currently serving as a researcher at the Cho Chun Shik Graduate School of Mobility, Korea Advanced Institute of Science and Technology (KAIST), South Korea.
    His research interests include adaptive control, neural-networks, and constrained optimization.
\end{IEEEbiography}

\begin{IEEEbiography}[{
    \includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{\template/people/dhHong.jpg}}]{Donghwa Hong}
    received his B.S. degree in Physics and Engineering Physics from Yonsei University, Wonju, South Korea, in 2023. He is currently pursuing the M.S. degree in Mechanical Engineering from GIST, Gwangju, South Korea.  
His research interests include robot control, neural-networks, and model identification. 
\end{IEEEbiography}

\begin{IEEEbiography}[{
    \includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{\template/people/khChoi.jpg}}]{Kyunghwan Choi}
    received his B.S., M.S., and Ph.D. degrees in Mechanical Engineering from KAIST, Daejeon, South Korea, in 2014, 2016, and 2020, respectively. Following his Ph.D., he worked at the Center for Eco-Friendly \& Smart Vehicles at KAIST as a Postdoctoral Fellow and later as a Research Assistant Professor. In 2022, he joined GIST as an Assistant Professor in the Department of Mechanical and Robotics Engineering. Currently, he serves as an Assistant Professor at the Cho Chun Shik Graduate School of Mobility at KAIST, where he is also the Director of the Mobility Intelligence and Control Laboratory. His research focuses on optimal and learning-based control for connected, automated, and electrified vehicles (CAEVs).
\end{IEEEbiography}

\end{document}