### 2025-06-10 presentation/presentation.tex
ppt final update for security check

% https://www.overleaf.com/learn/latex/Beamer#Using_a_colorthemev

% Inbuilt themes in beamer
\documentclass[8pt, aspectratio=169]{beamer}

\input{../template/macros/macros_general.tex}
\input{../template/macros/macros_math.tex}
\input{../template/symbols/symbols_NN.tex}
\input{../template/symbols/symbols_robot.tex}

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Block options: block, alertblock, exampleblock
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% % Theme choice:
% \usetheme{Antibes}
% % \usecolortheme{beaver}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Use roboto Font (recommended)
% \usepackage[sfdefault]{roboto}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Define where theme files are located. ('/styles')
\usepackage{styles/fluxmacros}
\usefolder{styles}
% Use Flux theme v0.1 beta
% Available style: asphalt, blue, red, green, gray 
\usetheme[style=blue]{flux}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Extra packages for the demo:
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{ragged2e}
\usepackage{schemabloc}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\usepackage{multirow}

% \usepackage{calc}
% \usepackage{svg}
% \usepackage{media9}             %pdflatex, latex+dvips+ps2pdf, xelatex
% \usepackage{multimedia}
% \usepackage{animate}

% Title page details: 
\title{
    Imposing a Weight Norm Constraint for Neuro-Adaptive Control 
} 
\subtitle{
    \textit{IEEE European Control Conference (ECC) 2025}\\
}
\author{
  \textbf{Myeongseok Ryu}\inst{1}, Jiyun Kim\inst{2}, and Kyunghwan Choi\inst{1}
  }
\date{2025-06-25}
\institute{%
    \begin{minipage}[c]{\linewidth}
        \centering
        \inst{1}%
        Department of Mechanical and Robotics Engineering\\
        Gwangju Institute of Science and Technology
        \and
        \inst{2}%
        AI Graduate School\\
        Gwangju Institute of Science and Technology
  \end{minipage}
}
\titlegraphic{assets/MIC_Lab_logo_white.png}
% make institute font size smaller
\setbeamerfont{institute}{size=\normalsize}

% \logo{
%   \vspace{-.7cm}%
%   \includegraphics[height=0.3cm]{assets/kaist_logo.png}
%   \includegraphics[height=0.3cm]{assets/MIC_Lab_logo.png}
%   \hspace{.6cm}%
%   }

\AtBeginSection[]{%
  \frame<beamer>{ 
    \frametitle{Outline}   
    \tableofcontents[currentsection] 
  }
}

% \AtBeginSubsection[]{%
%   \begin{frame}
%   \vfill
%   \centering
%     \insertsectionhead
%     \\
%     \large\textbf{\insertsubsectionhead}
%   \vfill
%   \end{frame}
% }

\begin{document}

% Title page frame
% \begin{frame}
\titlepage 
% \end{frame}

% Outline frame
\begin{frame}{Outline}
    \tableofcontents
\end{frame}

% Lists frame
\section{Background and Contributions}

\subsection{Introduction to Neuro-Adaptive Control}

\begin{frame}{\insertsubsectionhead}{What is Neuro-Adaptive Control?}

  \textbf{Neuro-Adaptive Control}
  \small{
    \begin{itemize}
      \item \textbf{Neuro-adaptive control} (NAC) is a control strategy that combines \textbf{neural networks (NNs)} with \textbf{adaptive control} \cite{Farrell:2006aa}.
      \item Features of both \textbf{NNs} and \textbf{adaptive control} can be found in NAC.
    \end{itemize}
  }

  \begin{figure}
    \includegraphics[width=0.65\textwidth]{figures/NAC.drawio.pdf}
  \end{figure}

\end{frame}

\begin{frame}{\insertsubsectionhead}{What is Neuro-Adaptive Control?}

  \textbf{Advantages of Neuro-Adaptive Control}
  \small{
    \begin{itemize}
      \item \textbf{Adaptability}: NAC adapts to changing environments and system dynamics.
      \item \textbf{Stability Guarantee}: The closed-loop stability is ensured using \textit{Lyapunov stability theory}.
      \item \textbf{Online Learning Capability}: NAC adapts in \textit{real-time} to new data with stability guarantees.
      \item \textbf{Robustness}: NAC handles \textit{uncertainties and disturbances} effectively with adaptive control techniques.
    \end{itemize}
  }

\begin{figure}
  \label{fig:general_framework}
  \includegraphics[width=0.65\textwidth]{figures/conv_nac.drawio.pdf}
  \caption{General framework of neuro-adaptive control (NAC).}
\end{figure}

\end{frame}

\begin{frame}{\insertsubsectionhead}{Existing Challenges in NAC}

  \textbf{Challenges}
  \small{
    \begin{itemize}
      \item \textbf{Optimality}: In general, the adaptation laws are driven with respect to the tracking error, which may not guarantee optimal performance.
      \item \textbf{Unpredictable Amplitude of Control input}: 
        \begin{itemize}
          \item The maximum amplitude of the NN weights is not predictable.
          \item This can result in unpredictable amplitude of the control input, which may lead to control saturation.
        \end{itemize}
      \item \textbf{Parameter Dependency}: 
      \begin{itemize}
        \item The performance of the NAC is highly dependent on the choice of parameters, such as learning rate and NN architecture.
      \end{itemize}
    \end{itemize}
  }

\end{frame}

% \subsection{Literature Review}

% \begin{frame}{\insertsubsectionhead}{Boundedness of NN Weights}

%   In general, the NN outputs are bounded by limiting the maximum amplitude of the NN weights.

%   \begin{enumerate}
%     \begin{columns}[T,onlytextwidth]
%         \column{0.49\textwidth}
%           \item \textbf{Projection Operator}

%           \begin{itemize}
%             \item Cats 
%             \item Dogs 
%             \item Birds
%           \end{itemize}

%         \column{0.49\textwidth}
%           \begin{figure}
%             \includegraphics[width=0.4\textwidth]{figures/KAIST-hi.png}
%           \end{figure}
%       \end{columns}

%       \begin{columns}[T,onlytextwidth]
%         \column{0.49\textwidth}
%           \item \textbf{$\epsilon$-modification, and $\sigma$-modification}

%           \begin{itemize}
%             \item Cats 
%             \item Dogs 
%             \item Birds
%           \end{itemize}

%         \column{0.49\textwidth}
%           \begin{figure}
%             \includegraphics[width=0.4\textwidth]{figures/KAIST-hi.png}
%           \end{figure}
%       \end{columns}
%   \end{enumerate}
% \end{frame}

% \begin{frame}{\insertsubsectionhead}{Input Saturation}

%   The input saturation is generally handled using an auxiliary system.

%   \begin{enumerate}
%     \begin{columns}[T,onlytextwidth]
%         \column{0.49\textwidth}
%           \item \textbf{Auxiliary Systems}

%           \begin{itemize}
%             \item Cats 
%             \item Dogs 
%             \item Birds
%           \end{itemize}

%         \column{0.49\textwidth}
%           \begin{figure}
%             \includegraphics[width=0.4\textwidth]{figures/KAIST-hi.png}
%           \end{figure}
%       \end{columns}

%       \begin{columns}[T,onlytextwidth]
%         \column{0.49\textwidth}
%           \item \textbf{Second?}

%           \begin{itemize}
%             \item Cats 
%             \item Dogs 
%             \item Birds
%           \end{itemize}

%         \column{0.49\textwidth}
%           \begin{figure}
%             \includegraphics[width=0.4\textwidth]{figures/KAIST-hi.png}
%           \end{figure}
%       \end{columns}
%   \end{enumerate}
% \end{frame}

\subsection{Research Objectives}

\begin{frame}{\insertsubsectionhead}

  \textbf{Objective 1: Optimality}
  \begin{itemize}
    \item Formulate a constrained optimization problem to minimize the tracking error.
    \item Guarantee the stability of the system and the NN weights.
  \end{itemize}

  \textbf{Objective 2: Stability}
  \begin{itemize}
    \item Derive an adaptation law that guarantees the stability of the system.
    \item Ensure that the NN weights remain bounded during operation.
  \end{itemize}

  \textbf{Objective 3: Boundedness of }
  \begin{itemize}
    \item Ensure the controller is robust to uncertainties and disturbances.
    \item Validate the proposed method through numerical simulations.
  \end{itemize}

\end{frame}

\section{Proposed Method}

\subsection{Architecture of the Proposed Method}

\begin{frame}{\insertsubsectionhead}

  \begin{columns}

    \column{0.5\textwidth}

      \textbf{Target Two-link Robotic Manipulator System}:
      \begin{itemize}
        \item Control input saturation function $\mysat(\cdot)$.
        \item Desired trajectory $\mv{q}_d$ is given.
      \end{itemize}
        \begin{equation}\label{eq:sys}
          \mm{M}\ddq + \mm{V}_m\dq + \mv{F} + \mv{G} + \mv{\tau}_d
          =
          \mysat(\mv{\tau})
        \end{equation}

      \textbf{Control Input}:
      \begin{itemize}
        \item NN's output $\mv{\Phi}$ is used as the control input.
        \item Consists of the estimated NN weights $\estwth$.
      \end{itemize}
      \begin{equation}\label{eq:input}
        \mv{\tau} := \mv{\Phi}(\mv{q}_n; \estwth)
      \end{equation}

      \textbf{Deep Neural Network (DNN)}:
      \begin{itemize}
        \item $k$ layers with $\estwth_i:=\myvec(\estwV_i)$.
        \item Activation function: $\phi(\cdot):=tanh(\cdot)$.
        \begin{equation}\label{eq:NN}
          \mv{\Phi}(\mv{q}_n; \estwth)
          :=
          \begin{cases}
              \estwV_i^\top \act_i(\estNN_{i-1}), 
              &
              i\in\{1,\dots ,k\},
              \\
              \estwV_0^\top {\q}_n,
              &
              i=0
              ,
          \end{cases}
        \end{equation}
      \end{itemize}

    \column{0.5\textwidth}

      \begin{figure}
        \includegraphics[width=0.8\textwidth]{figures/Controller.drawio.pdf}
        \caption{Architecture of the proposed method.}
      \end{figure}

      \begin{figure}
        \includegraphics[width=0.7\textwidth]{figures/DNN.drawio.pdf}
        \caption{Architecture of the DNN.}
      \end{figure}

  \end{columns}

    \let\thefootnote\relax\footnote{
      \textit{Notations}: 
        $\mv{q}\in\R^n$: Joint position, $\mm{M}$: Inertia matrix, $\mm{C}$: Coriolis matrix, $\mm{G}$: Gravity vector, $\mv{\tau}$: Control input, $\estwth$: Estimated NN weights.
        $\estwth$: Estimated NN weights, $\mv{e}$: Tracking error, $\mv{\tau}_d$: Disturbance, $\mysat(\cdot)$: Saturation function.
      }

\end{frame}

\subsection{Problem Formulation}

\begin{frame}{\insertsubsectionhead}{Optimization Problem}

\begin{columns}

  \column{0.45\textwidth}

    \textbf{Optimization Problem Statement}:

    \begin{itemize}
      \item \textbf{Find} NN weights $\estwth$,
      \item That \textbf{minimize} objective function $J(\cdot)$,
        \begin{equation}\label{eq:obj}
          J(\mv{r};\estwth)
          := 
          \frac{1}{2} \mv{r}^\top \mv{r}.
        \end{equation}
        \begin{itemize}
          \item where $\mv{r} := \ddt\mv{e} + \Lambda\mv{e}$ is the filtered tracking error,
        \end{itemize}
      \item while satisfying the following \textbf{constraints}:
        \begin{itemize}
          \item Boundedness of the NN weights $\estwth$.
          \item Saturation of the control input $\mv{\tau}$.
        \end{itemize}
    \end{itemize}

  \column{0.5\textwidth}

    \begin{block}{Considered Constraints}

      \begin{itemize}
        \item \textbf{Weight Boundedness for Each Layer}: 
          \begin{equation}\label{eq:cstr:weight}
            c_{\theta_i}(\estwth)
            :=
            \norm{\estwth_i}^2 - \overline{\theta_i}^2 \le 0
            , 
            \forall i\in\{0,\ldots,k\}
          \end{equation}

        \item \textbf{Convex control Input Saturation}: 
          \begin{itemize}
            \item Input bound constraint:
            \begin{equation}\label{eq:cstr:input_bound}
              c_{\overline{\tau}_i}(\estwth)
              :=
              \tau_i - \overline{\tau_i}
              \le 
              0
              ,
              \quad
              c_{\underline{\tau}_i}(\estwth)
              :=
              \underline{\tau_i} - \tau_i
              \le 
              0
            \end{equation}
            \item Input norm constraint:
            \begin{equation}\label{eq:cstr:input_norm}
              c_{\mv{\tau}}(\estwth)
              :=
              \norm{\mv{\tau}}^2 - \overline{\tau}^2
              \le
              0
            \end{equation}
        \end{itemize}
      \end{itemize}

    \end{block}

  \end{columns}

  \let\thefootnote\relax\footnote{
    \textit{Notations}: 
    $\Lambda\in\R_{>0}^{n\times n}$: filtering matrix
  }

\end{frame}

\begin{frame}{\insertsubsectionhead}{Optimization Problem}

  \textbf{Original Optimization Problem}

  \begin{itemize}
    \item Constrained optimization problem to minimize the tracking error.
    \item Inequality constraints $c_j(\estwth) \le 0$ for $j\in\mathcal{I}$.
  \end{itemize}

    \begin{equation}\label{eq:opt_problem}
      \begin{matrix}
        \min_{\estwth} J(\mv{r};\estwth)
        \\\ \\
        \text{s.t. } c_{j}(\estwth) 
        \le 
        0
        ,
        \forall j\in\mathcal{I}
      \end{matrix}
    \end{equation}

  \textbf{Define Lagrangian Function}

  \begin{equation}\label{eq:lagrangian}
      L(\mv{r},\estwth,[\lambda_j]_{j\in\mathcal{I}})
      :=
      J(\mv{r};\estwth)
      +
      \sum_{j\in\mathcal{I}} \lambda_j c_j(\estwth)
  \end{equation}

  \textbf{Dual Problem}
  \begin{itemize}
    \item The dual problem is to minimize the Lagrangian function with respect to the NN weights $\estwth$, while maximizing the Lagrange multipliers $\lambda_j$.
    \item The Lagrange multipliers $\lambda_j$ are non-negative, \ie $\lambda_j \ge 0$.
  \end{itemize}

  \begin{equation}\label{eq:dual_problem}
    \min_{\estwth}\max_{[\lambda_j]_{j\in\mathcal{I}}} 
    L(\mv{r},\estwth,[\lambda_j]_{j\in\mathcal{I}})
  \end{equation}

\end{frame}

\subsection{Adaptation Law Derivation}

\begin{frame}{\insertsubsectionhead}{Gradient Descent/Ascent Method}

  For the dual problem, 
  \begin{equation}
    \min_{\estwth}\max_{[\lambda_j]_{j\in\mathcal{I}}} 
    L(\mv{r},\estwth,[\lambda_j]_{j\in\mathcal{I}})
    ,
  \end{equation}
  the first-order gradient descent/ascent method is used to derive the adaptation law.

  \centering
  \begin{minipage}{0.6\textwidth}%

    \begin{block}{Adaptation Law}%

    \textbf{Gradient Descent Method for $\estwth$}:
    \begin{equation}\label{eq:adaptation_law:1}
      \ddt {{\estwth}}
      =
      -\alpha 
      \ppfrac{L}{{\estwth}}
      =-\alpha 
      \left(
          \ppfrac{J}{{\estwth}}
          +
          \sum_{j\in\mathcal{I}}
          \lambda_j 
          \ppfrac{c_j}{\estwth}
      \right),
    \end{equation}

    \textbf{Gradient Ascent Method for $\lambda_j, \forall j\in\mathcal I$}:
    \begin{equation}\label{eq:adaptation_law:2}
      \ddt\lambda_j
      = 
      \beta_j
      \ppfrac{L}{\lambda_j} 
      = 
      \beta_j c_j ,
    \end{equation}

    For non-negativity of the Lagrange multipliers,
    \begin{equation}\label{eq:adaptation_law:3}
      \lambda_j
      \leftarrow
      \max(\lambda_j,0)
      .
    \end{equation}

    % \begin{subequations}
    %     \begin{align}
    %       \ddt {{\estwth}}
    %       &
    %       =
    %       -\alpha 
    %       \ppfrac{L}{{\estwth}}
    %       =-\alpha 
    %       \left(
    %           \ppfrac{J}{{\estwth}}
    %           +
    %           \sum_{j\in\mathcal{I}}
    %           \lambda_j 
    %           \ppfrac{c_j}{\estwth}
    %       \right),
    %       \\
    %       \ddt\lambda_j
    %       & 
    %       = 
    %       \beta_j
    %       \ppfrac{L}{\lambda_j} 
    %       = 
    %       \beta_j c_j ,
    %       \quad\quad\quad\quad      \      
    %       \forall j\in\mathcal I,
    %       \\
    %       \lambda_j
    %       &
    %       \leftarrow
    %       \max(\lambda_j,0)
    %     \end{align}
    % \end{subequations}

    \end{block}
  \end{minipage}
\end{frame}

\subsection{Stability Analysis}

\begin{frame}{\insertsubsectionhead}{Lyapunov Stability Analysis}

  \begin{block}{Theorem 1 \cite{Ryu:2025aa}}
    For the dynamical system described in \eqref{eq:sys}, the neuro-adaptive controller in \eqref{eq:input} with the weight adaptation laws in \eqref{eq:adaptation_law:1}, \eqref{eq:adaptation_law:2} and \eqref{eq:adaptation_law:3} ensure the boundedness of the filtered error $\mv{r}$ and the weight estimate $\estwth$, under the control input constraints satisfying Assumption 1 and 2. This holds under the weight norm constraint \eqref{eq:cstr:weight}.
  \end{block}

  \begin{exampleblock}{Assumption 1 (Convex Input Constraint)}
    The constraint functions $c_j(\estwth),\forall j\in\mathcal{I}$, are convex in the $\tau$-space and satisfy $c_j(\estwth)\le0$ and $c_j(\idealwth)\le0$.
  \end{exampleblock}

  \begin{exampleblock}{Assumption 2, Linear Independence Constraint Qualification (LICQ)}
    The selected constraints satisfy the Linear Independence Constraint Qualification (LICQ) \cite[Chap. 12 Def. 12.1]{Nocedal:2006aa}.
  \end{exampleblock}

  Proof of Theorem 1 is omitted due to space limitations. The detailed proof can be found in \cite{Ryu:2025aa}.

\end{frame}

\section{Experimental Validation}

\subsection{Simulation Setup}

\begin{frame}{\insertsubsectionhead}{Two-Link Robotic Manipulator}

    \begin{columns}

      \column{.45\textwidth}

        \textbf{Target System}:

        \begin{equation*}
          \mm{M}\ddq + \mm{V}_m\dq + \mv{F} + \mv{G} + \mv{\tau}_d
          =
          \mv{\tau}
        \end{equation*}

        \begin{figure}
          \centering
          \includegraphics[width=.99\textwidth]{figures/RobotModel.drawio.png}
          \caption{Two-link robotic manipulator model.}
        \end{figure}

      \column{.55\textwidth}

        \textbf{Desired Trajectory}:

        \begin{equation}
          \qd
          =
          \begin{bmatrix}
              {q_d}_1\\
              {q_d}_2
          \end{bmatrix}
          = 
          \begin{bmatrix}
              +\cos(
                  \tfrac{\pi}{2}t
              ) + 1 \\
              -\cos(
                  \tfrac{\pi}{2}t
              ) - 1 
          \end{bmatrix}
          .
      \end{equation}

      \textbf{System Model Parameters}:

      \begin{table}
        \renewcommand{\arraystretch}{1.3}
        \caption{System model parameters.}
        \centering
        \begin{tabular}{c m{5em} c c c }
        \hline
        \textbf{Symbol} & \textbf{Description} & \textbf{Link 1} & \textbf{Link 2} \\
        \hline
        \hline 
        $m_p$ & Mass & 23.902 kg & 3.88 kg \\
        \hline
        $l_p$  & Length & 0.45 m & 0.45 m \\
        \hline
        ${l_c}_p$ & COM & 0.091 m & 0.048 m \\
        \hline
        $b_p$   & Viscous coef. &  2.288 Nms & 0.172 Nms \\
        \hline
        ${f_c}_p$  & Friction coef. & 7.17 Nm & 1.734 Nm \\
        \hline
        \end{tabular}
        \label{table: system parameters}
      \end{table}

    \end{columns}

\end{frame}

\begin{frame}{\insertsubsectionhead}{Controllers for Comparative Study}

    \begin{itemize}
      \item NAC-CO denotes the proposed controller based on constrained optimization.
      \item For NAC-L2 and NAC-eMod, the stabilizing terms $-\lambda\estwth$ and $\rho\norm{\tilde{\mv{z}}}\estwth$ ensures the boundedness of the NN weights
    \end{itemize}

    \begin{table}
      \renewcommand{\arraystretch}{1.5}
      % \caption{Controller Properties.}
      \centering
      \begin{tabular}{c m{20em} c c c }
      \hline
      \textbf{Name} & \textbf{Description} & \textbf{Adaptation Law} \\
      \hline
      \hline 
        \multirow{2}{*}{NAC-CO} & Constrained Optimization-based NAC & 
      $
        \ddt\estwth = -\alpha \left(\ppfrac{L}{\estwth}+\sum_{j\in\mathcal{I}}\lambda_j\ppfrac{c_j}{\estwth}\right)
      $ 
      \\
        (proposed) & ($\beta_j$ determines $\lambda_j$ adaptation speed) &
      $
        \ddt\lambda_j = \beta_j c_j, \lambda_j \leftarrow \max(\lambda_j,0)
      $
      \\
      \hline
      \multirow{2}{*}{NAC-L2} & NAC with $L_2$-regularization & \multirow{2}{*}
      {$
        \ddt\estwth = -\alpha \left(\ppfrac{J}{\estwth}+\lambda\estwth\right)
      $}
      \\
        & ($\lambda\in\R_{>0}$ stabilizes $\estwth$ towards origin) &
      \\
      \hline
      \multirow{2}{*}{NAC-eMod} & NAC with $\epsilon$-modification & \multirow{2}{*}
      {$
        \ddt\estwth = -\alpha \left(\ppfrac{J}{\estwth}+\rho\norm{\tilde{\mv{z}}}\estwth\right)
      $}
      \\
      & ($\rho$ stabilizes proportionally to the error $\tilde{\mv{z}}$) &
      \\
      \hline
      \end{tabular}
      \label{table:sys:param}
    \end{table}

    \centering
    \begin{minipage}{0.75\textwidth}
      \begin{block}{Simulation Objective}
        By varying the parameters, \ie $\beta_j$, $\lambda$, and $\rho$, the parameter dependencies will be investigated.
      \end{block}
    \end{minipage}

\end{frame}

\subsection{Simulation Results}

% \begin{frame}{Representative Simulation Results}

  % % \movie[width=3cm,height=2cm,poster]{}{overall.mp4}
  % \begin{figure}
  %   \centering
  %   \href{https://youtu.be/U5uDbJQkGY8}{
  %     \includegraphics[width=.1\textwidth]{figures/KAIST-hi.png}
  %   }
  % \end{figure}

  % \animategraphics[loop,width=4cm]{gifs/overall-}
  % \animategraphics[loop,width=4cm]{10}{gifs/overall-}{0}{10}

% \end{frame}

\begin{frame}{\insertsubsectionhead}{Box-and-Whisker Plots}

  \begin{columns}
    \column{0.55\textwidth}

      \textbf{Parameter Dependencies Investigation}:

      \begin{itemize}
        \item The parameters ranged from 0.001 to 1 across 10 samples.
        \item NAC-CO (proposed) shows the best performance and low variance.
        \item NAC-L2 shows the worst performance with high variance.
      \end{itemize}

    \column{0.5\textwidth}

      \begin{figure}
        \includegraphics[width=.99\textwidth]{figures/BoxWhisker.drawio.png}
        \caption{Parameter dependencies of the proposed method.}
      \end{figure}

  \end{columns}

    \begin{table}[!t]
      \renewcommand{\arraystretch}{1.1}
      % \caption{Quantitative comparison of square root of tracking ISE.}
      \centering
      \begin{tabular}{c c c c }
      \hline
      & \textbf{NAC-L2}\!&\!\textbf{NAC-eMod}\!&\!\textbf{NAC-CO} (proposed) 
      \\
      \hline
      \hline 
        Maximum\!&\!$11.1753\!\times\!10^{-3}$\!&\!$0.5603\!\times\!10^{-3}$\!&\!$0.3439\!\times\!10^{-3}$\!\\
      \hline
        Median\!&\!$0.5898\!\times\!10^{-3}$\!&\!$0.5519\!\times\!10^{-3}$\!&\!$0.3240\!\times\!10^{-3}$\!\\
      \hline
        Minimum\!&\!$0.5434\!\times\!10^{-3}$\!&\!$0.5434\!\times\!10^{-3}$\!&\!$0.3235\!\times\!10^{-3}$\!\\
      \hline
      \end{tabular}
      \label{table: error norm}
    \end{table}

    \let\thefootnote\relax\footnote{
      Squared root of the tracking error ISE (Integral of Squared Error) is used, \ie $\sqrt{\int_0^T \norm{\mv{r}}^2\,\der t}$, where $T$ denotes a simulation termination time.
    }

\end{frame}

\begin{frame}{\insertsubsectionhead}{Weight Norms}

  \begin{columns}

    \column{0.33\textwidth}

      \begin{figure}
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig8.eps}
        \caption{Weight norms of NAC-CO}
      \end{figure}

    \column{0.33\textwidth}

      \begin{figure}      
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig9.eps}
        \caption{Weight norms of NAC-L2}
      \end{figure}

    \column{0.33\textwidth}

      \begin{figure}
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig10.eps}
        \caption{Weight norms of NAC-eMod}
      \end{figure}

  \end{columns}

  \begin{itemize}
    \item NAC-CO (proposed) showed the weight norms are bounded under pre-defined constraint $\overline{\theta}=20$.
    \item NAC-L2 and NAC-eMod showed the bounded weight norms, but they depended on the parameters $\lambda$ and $\rho$, respectively.
    \item As the parameters $\lambda$ and $\rho$ increase, the weight norms of NAC-L2 and NAC-eMod biased towards the origin, which may lead to suboptimal performance.
    \item In addition, NAC-CO tracked the desired trajectory with a smaller weight norm than NAC-L2 and NAC-eMod.
  \end{itemize}

\end{frame}

\begin{frame}{\insertsubsectionhead}{Tracking Performance}

  \begin{columns}

    \column{0.33\textwidth}

      \begin{figure}
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig5.eps}
        \caption{Tracking error of NAC-CO}
      \end{figure}

    \column{0.33\textwidth}

      \begin{figure}      
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig6.eps}
        \caption{Tracking error of NAC-L2}
      \end{figure}

    \column{0.33\textwidth}

      \begin{figure}
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig7.eps}
        \caption{Tracking error of NAC-eMod}
      \end{figure}

  \end{columns}

  \begin{itemize}
    \item NAC-CO (proposed) outperformed NAC-L2 and NAC-eMod in terms of tracking performance.
    \item As the weights are biased towards the origin, the tracking performance of NAC-L2 and NAC-eMod deteriorated.
  \end{itemize}

\end{frame}

% \subsection{Simulation Setup}

% \begin{frame}{\insertsubsectionhead}{Two-Link Robotic Manipulator}

%     \begin{columns}

%       \column{.55\textwidth}
%         \begin{figure}
%           \centering
%           \includegraphics[width=.85\textwidth]{figures/RobotModel.drawio.png}
%           \caption{Two-Link Robotic Manipulator}
%         \end{figure}

%       \column{.5\textwidth}

%         \begin{table}[t]
%             \renewcommand{\arraystretch}{1.3}
%             \caption{Two-link robotic manipulator's parameters.}
%             \centering
%             \begin{tabular}{c m{9.5em} c c c }
%             % \begin{tabular}{c c c c c}
%             \hline
%             \textbf{Symbol} & \textbf{Description} & \textbf{Value} \\
%             \hline
%             \hline 
%             $m_1,m_2$ & Mass & 2.465 kg \\
%             \hline
%             $l_1,l_2$  & Length & 0.2 m \\  
%             \hline
%             ${l_c}_1,{l_c}_2$ & Center of mass & 0.139 m \\
%             \hline
%             $I_1,I_2$  & Inertia & 0.069 kgm\textsuperscript{2} \\
%             \hline
%             \end{tabular}
%         \end{table}

%     \end{columns}

% \end{frame}

\section{Conclusion}

\subsection{Conclusion and Future Work}

\begin{frame}{Conclusion}

  \textbf{Summary of Contributions}
  \begin{itemize}
    \item Proposed a novel constrained optimization-based neuro-adaptive control (CONAC) method.
    \item Adaptation laws are derived using constrained optimization method.
    \item The proposed method guarantees the stability of the system and the boundedness of the NN weights.
    \item Feasibility of the proposed method is validated through numerical simulations and real-time experiments.
  \end{itemize}

  \textbf{Future Work}
  \begin{itemize}
    \item Extend the proposed method to state constraints.
    \item Enhance the robustness and flexibility of the proposed method for various systems.
  \end{itemize}

\end{frame}

\section{}
\begin{frame}{}
    \centering \Large
    \emph{Thank you for your attention!}
\end{frame}

\begin{frame}[allowframebreaks]{References}

  \bibliography{localRefs}
  \bibliographystyle{IEEEtran}

\end{frame}

\end{document}