### 2025-06-18 presentation/presentation.tex
ppt final peer review 1

% https://www.overleaf.com/learn/latex/Beamer#Using_a_colorthemev

% Inbuilt themes in beamer
% \documentclass[8pt, aspectratio=169, handout]{beamer}
\documentclass[8pt, aspectratio=169]{beamer}
% \setbeamertemplate{caption}[numbered]

\input{../template/macros/macros_general.tex}
\input{../template/macros/macros_math.tex}
\input{../template/symbols/symbols_NN.tex}
\input{../template/symbols/symbols_robot.tex}

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Block options: block, alertblock, exampleblock
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% % Theme choice:
% \usetheme{Antibes}
% % \usecolortheme{beaver}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Use roboto Font (recommended)
% \usepackage[sfdefault]{roboto}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Define where theme files are located. ('/styles')
\usepackage{styles/fluxmacros}
\usefolder{styles}
% Use Flux theme v0.1 beta
% Available style: asphalt, blue, red, green, gray 
\usetheme[style=dding]{flux}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Extra packages for the demo:
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{ragged2e}
\usepackage{schemabloc}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\usepackage{multirow}
\usepackage{media9}

% https://latexcolor.com % color list
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\definecolor{awesome}{rgb}{1.0, 0.13, 0.32}
\newcommand{\ctxt}[2]{\color{#1}{#2}\color{black} }
% \newcommand{\ctxt}[2]{\color{#1}\mathbf{#2}\color{black} }

% Title page details: 
\title{
    Imposing a Weight Norm Constraint for Neuro-Adaptive Control 
} 
\subtitle{
    \textit{IEEE European Control Conference (ECC) 2025}\\
}
\author{
  \textbf{\ctxt{airforceblue}{Myeongseok Ryu}}\inst{1}, Jiyun Kim\inst{2}, and Kyunghwan Choi\inst{1}
  }
\date{2025-06-25}
\institute{%
    \begin{minipage}[c]{\linewidth}
        \centering
        \inst{1}%
        Mobility Intelligence and Control Laboratory (MIC Lab) \\
        % Department of Mechanical and Robotics Engineering\\
        % Gwangju Institute of Science and Technology
        CCS Graduate School of Mobility \\
        Korea Advanced Institute of Science and Technology (KAIST)
        \and
        \inst{2}%
        AI Graduate School\\
        Gwangju Institute of Science and Technology (GIST)
  \end{minipage}
}

% \titlegraphic{assets/MIC_Lab_logo.png}
\titlegraphic{assets/MIC_Lab_logo_white.png}

% make institute font size smaller
\setbeamerfont{institute}{size=\normalsize}

\AtBeginSection[]{%
  \frame<beamer>{ 
    \frametitle{Outline}   
    \tableofcontents[currentsection] 
  }
}

% \AtBeginSubsection[]{%
%   \begin{frame}
%   \vfill
%   \centering
%     \insertsectionhead
%     \\
%     \large\textbf{\insertsubsectionhead}
%   \vfill
%   \end{frame}
% }

\begin{document}

% Title page frame
\titlepage

% Outline frame
\begin{frame}{Outline}
    \tableofcontents
\end{frame}

% ╔═══════════════════════════════════════════════╗
% ║ Section: Background and Contributions         ║
% ╚═══════════════════════════════════════════════╝
\section{Background and Contributions}

\subsection{Introduction to Neuro-Adaptive Control}

\begin{frame}{\insertsubsectionhead}{What is Neuro-Adaptive Control?}

  \textbf{Neuro-Adaptive Control}
  \small{
    \begin{itemize}
      \item \textbf{Neuro-adaptive control} (NAC) is a control strategy that combines \ctxt{airforceblue}{neural networks (NNs) } with \ctxt{awesome}{adaptive control } \cite{Farrell:2006aa}.
      \item Features of both \ctxt{airforceblue}{NNs } and \ctxt{awesome}{adaptive control } can be found in NAC.
    \end{itemize}
  }

  \begin{figure}
    \includegraphics[width=0.65\textwidth]{figures/NAC.drawio.pdf}
  \end{figure}

\end{frame}

\begin{frame}{\insertsubsectionhead}{What is Neuro-Adaptive Control?}

  \textbf{Advantages of Neuro-Adaptive Control}
  \small{
    \begin{itemize}
      \item <2-> \textbf{Adaptability}: NAC adapts \ctxt{airforceblue}{NN weights }to changing environments and system dynamics.
      \item <3-> \textbf{Stability Guarantee}: The closed-loop stability is ensured using \ctxt{airforceblue}{Lyapunov stability theory}.
      \item <4-> \textbf{Online Learning Capability}: NAC adapts in \ctxt{airforceblue}{real-time } to new data with stability guarantees.
      \item <5-> \textbf{Robustness}: NAC handles \ctxt{airforceblue}{uncertainties and disturbances }effectively with adaptive control techniques.
    \end{itemize}
  }

\begin{figure}
  \label{fig:general_framework}
  \includegraphics[width=0.65\textwidth]{figures/conv_nac.drawio.pdf}
  \caption{General framework of neuro-adaptive control (NAC).}
\end{figure}

\end{frame}

\begin{frame}{\insertsubsectionhead}{What is Neuro-Adaptive Control?}

  \textbf{Existing Challenges in NAC}

  \begin{enumerate}

    \onslide<2->
    {
      \begin{columns}[T,onlytextwidth]

        \column{0.49\textwidth}

          \small{
            \item \textbf{Weight Boundedness:}

            \begin{itemize}
              \item Generally, NN weights are adapted by \ctxt{airforceblue}{gradient descent method}.
              \begin{itemize}
                \item Objective function typically consists of the control error.
              \end{itemize}
              \item Hence, the NN weights can grow \ctxt{awesome}{unbounded}, leading to instability (also known as \textit{parameter drift}).
              \item Unbounded weights can cause the NN to produce \ctxt{airforceblue}{large control inputs}, which may lead to following challenges.
            \end{itemize}
          }

          \column{0.49\textwidth}
            \centering

              \begin{figure}
                \includegraphics[width=0.7\textwidth]{figures/unbounded.drawio.pdf}
                \caption{Divergence of NN weights even with convergent control error.}
              \end{figure}
        \end{columns}
      }

      \onslide<3->
      {
        \begin{columns}[T,onlytextwidth]
          \column{0.49\textwidth}

          \small{
            \item \textbf{Unpredictable Amplitude of Control Input:}

            \begin{itemize}
              \item Typical issue of control problem in physical systems.
              \item The NN outputs are \ctxt{awesome}{unpredictable } and \ctxt{awesome}{not interpretable}.
              \item These features—unbounded NN weights and unpredictable amplitudes—can lead to \ctxt{airforceblue}{input saturation}.
            \end{itemize}
          }

          \column{0.49\textwidth}
            \centering
            \begin{figure}
              \includegraphics[width=0.8\textwidth]{figures/unpredictable.drawio.pdf}
              \caption{Unpredictable amplitude of NN outputs.}
            \end{figure}
        \end{columns}
      }

  \end{enumerate}

\end{frame}

\subsection{Literature Review}

\begin{frame}{\insertsubsectionhead}

  \begin{enumerate}

    \begin{columns}[T]

        \column{0.33\textwidth}
          \item \textbf{Projection Operator} \small{\textit{for weight boundednss}}

          \onslide<2->
          {
            \begin{itemize}
              \item Projects the NN weights onto a convex set.
              \item Ensures that the weights \ctxt{airforceblue}{remain }within a predefined bound.
            \end{itemize}

            \begin{equation}
              \estwth \leftarrow \myproj_{\overline{\theta}}(\estwth)
            \end{equation}
          }

        \column{0.33\textwidth}
          \item \textbf{$\sigma$-modification, and $\epsilon$-modification} \small{\textit{for weight boundednss}}

          \onslide<3->
          {
            \begin{itemize}
              \item Add a \ctxt{airforceblue}{stabilizing term }(\eg $ -\lambda \estwth $) to adaptation law.
              \item Construct a \ctxt{airforceblue}{invariant set }of the NN weights.
            \end{itemize}

            \begin{equation}
              \ddtt{\estwth} \leftarrow \ddtt{\estwth} - \ctxt{awesome}{\lambda \estwth}
            \end{equation}
          }

        \column{0.33\textwidth}
          \item \textbf{Additional Control Inputs} \small{\textit{for control saturation}}

          \onslide<4->
          {
            \begin{itemize}
              \item \ctxt{airforceblue}{Conventional controllers }are used to address control input saturation. 
                    \begin{itemize}
                      \item Barrier Lyapunov function or auxiliary system-based control inputs.
                    \end{itemize}
              \item In general, \ctxt{awesome}{nominal models }are required.
            \end{itemize}
          }

      \end{columns}

  \end{enumerate}

  \vspace{0.5cm}

  \begin{columns}[T]

      \column{0.33\textwidth}

        \onslide<2->
        {
          \centering
          \begin{figure}
            \includegraphics[width=0.8\textwidth]{figures/existing_proj.drawio.pdf}
            \caption{Projection of NN weights on a convex set.}
          \end{figure}
        }

      \column{0.33\textwidth}

        \onslide<3->
        {
          \centering
          \begin{figure}
            \includegraphics[width=0.8\textwidth]{figures/existing_modi.drawio.pdf}
            \caption{Adaptation result with stabilizing function (\eg $\sigma$-modification).}
          \end{figure}
        }

      \column{0.33\textwidth}

        \onslide<4->
        {
          \centering
          \begin{figure}
            \includegraphics[width=0.999\textwidth]{figures/existing_conv.drawio.pdf}
            \caption{Control input saturation handling with additional control inputs.}
          \end{figure}
        }

    \end{columns}

\end{frame}

\begin{frame}{\insertsubsectionhead}{Limitations of Existing Methods}

  \onslide<1->
  {
    \textbf{Limitation 1: Lack of Optimality}
  }
  \begin{itemize}
    \item<2-> The existing methods do not guarantee the \ctxt{airforceblue}{optimality } of the control input.
    \item<3-> Projection operator:
    \begin{itemize}
      \item The projection operator \ctxt{awesome}{simply projects } the NN weights onto a convex set, regardless of the imposed \ctxt{airforceblue}{constraints } (\eg weight boundedness or input saturation).
      \item Moreover, if the convex set is conservatively defined, the weights may be limited to a \ctxt{awesome}{suboptimal region}.
    \end{itemize}
    \item<4-> $\sigma$- and $\epsilon$-modification:
    \begin{itemize}
      \item The stabilizing term \ctxt{awesome}{biases } the NN weights towards the origin.
      \item Therefore, the weights converge toward a \ctxt{awesome}{suboptimal point}.
    \end{itemize}
  \end{itemize}

  \onslide<1->
  {
    \textbf{Limitation 2: Disruption of Learning Process by Additional Control Inputs}
  }
  \begin{itemize}
    \item<5-> Feedback tracking error for learning is \ctxt{awesome}{disrupted } by additional control inputs.
      \begin{itemize}
        \item The feedback error \ctxt{airforceblue}{does not reflect } the error induced by the NN, directly.
        \item The additional control inputs may exceeds the \ctxt{awesome}{input saturation limits }, already.
    \end{itemize}  
  \end{itemize}

\end{frame}

\subsection{Contributions}

\begin{frame}{\insertsubsectionhead}

  \textbf{Contribution 1: Unified Optimization Framework} 
  \onslide<2->
  {
    \begin{itemize}
      \item Trajectory tracking and constraint handling are formulated as a \ctxt{airforceblue}{unified constrained optimization } problem.
      \item The \ctxt{awesome}{conventional controllers } does not required.
      \begin{itemize}
        \item Nominal model knowledge is not required for the conventional controllers.
      \end{itemize}
    \end{itemize}
  }

  \textbf{Contribution 2: Online Learning Capability (Stability Guarantees)}
  \onslide<3->
  {
    \begin{itemize}
      \item Stability are rigorously proven using \ctxt{airforceblue}{Lyapunov stability theory}.
      \item Hence, \ctxt{airforceblue}{online learning} with \ctxt{awesome}{no prior system knowledge } is possible.
    \end{itemize}
  }

  \textbf{Contribution 3: Weight and Control Input Constraint Handling}
  \onslide<4->
  {
    \begin{itemize}
      \item Weight and control input \ctxt{awesome}{constraints } are \ctxt{airforceblue}{explicitly considered } in the optimization problem.
      \item \ctxt{airforceblue}{Any combination } of convex input constraints can be handled.
    \end{itemize}
  }

\end{frame}

% ╔═══════════════════════════════════════════════╗
% ║ Section: Proposed Method                      ║
% ╚═══════════════════════════════════════════════╝
\section{Proposed Method}

\subsection{Architecture of the Proposed Method}

\begin{frame}{\insertsubsectionhead}

  \begin{columns}

    \column{0.5\textwidth}

    \onslide<2->
    {
      \textbf{Target Two-link Robotic Manipulator System}:
      \small{
        \begin{itemize}
          \item Control input \ctxt{awesome}{saturation function } $\mysat(\cdot)$.
          \item Desired trajectory $\mv{q}_d$ is given.
        \end{itemize}
          \begin{equation}\label{eq:sys}
            \mm{M}\ddq + \mm{V}_m\dq + \mv{F} + \mv{G} + \mv{\tau}_d
            =
            \mysat(\mv{\tau})
          \end{equation}    
      }
    }

    \onslide<3->
    {
      \textbf{Control Input}:
      \small{
        \begin{itemize}
          \item NN's \ctxt{airforceblue}{output } $\mv{\Phi}$ is used as the control input.
          \item Consists of the \ctxt{awesome}{estimated NN weights } $\estwth$.
        \end{itemize}
        \begin{equation}\label{eq:input}
          \mv{\tau} := \mv{\Phi}(\mv{q}_n; \estwth)
        \end{equation}
      }
    }

    \onslide<4->
    {
      \textbf{Deep Neural Network (DNN)}:
      \small{
        \begin{itemize}
          \item $k$ layers with weights $\estwth_i:=\myvec(\estwV_i)$.
          \item Activation function: $\phi(\cdot):=tanh(\cdot)$.
          \begin{equation}\label{eq:NN}
            \mv{\Phi}(\mv{q}_n; \estwth)
            :=
            \begin{cases}
                \estwV_i^\top \act_i(\estNN_{i-1}), 
                &
                i\in\{1,\dots ,k\},
                \\
                \estwV_0^\top {\q}_n,
                &
                i=0
                ,
            \end{cases}
          \end{equation}
        \end{itemize}
      }
    }

    \column{0.5\textwidth}

      \begin{figure}
        \includegraphics[width=0.8\textwidth]{figures/Controller.drawio.pdf}
        \caption{Architecture of the proposed method.}
      \end{figure}

    \onslide<4->
    {
      \begin{figure}
        \includegraphics[width=0.7\textwidth]{figures/DNN.drawio.pdf}
        \caption{Architecture of the DNN.}
      \end{figure}
    }

  \end{columns}

    \let\thefootnote\relax\footnote{
      \textit{Notations}: 
        $\mv{q}\in\R^n$: Joint position, $\mm{M}$: Inertia matrix, $\mm{C}$: Coriolis matrix, $\mm{G}$: Gravity vector, $\mv{\tau}$: Control input, $\mv{\tau}_d$: Disturbance.
      }

\end{frame}

\subsection{Problem Formulation}

\begin{frame}{\insertsubsectionhead}{Optimization Problem}

\begin{columns}

  \column{0.45\textwidth}

    \textbf{Optimization Problem Statement}:

    \begin{itemize}
      \item \ctxt{awesome}{Find } NN weights $\estwth$,
      \item That \ctxt{airforceblue}{minimize } objective function $J(\cdot)$,
        \begin{equation}\label{eq:obj}
          J(\mv{r};\estwth)
          := 
          \frac{1}{2} \mv{r}^\top \mv{r}.
        \end{equation}
        \begin{itemize}
          \item where $\mv{r} := \ddt\mv{e} + \Lambda\mv{e}$ is \ctxt{awesome}{filtered tracking error},
        \end{itemize}
      \item while satisfying the following \ctxt{awesome}{constraints}:
        \begin{itemize}
          \item \ctxt{airforceblue}{Boundedness } of the NN weights $\estwth$.
          \item \ctxt{airforceblue}{Saturation } of the control input $\mv{\tau}$.
        \end{itemize}
    \end{itemize}

  \column{0.5\textwidth}

  \onslide<2->
  {
    \begin{block}{Considered Constraints}

      \begin{itemize}
        \item \textbf{Weight Boundedness for Each Layer}: 
          \begin{equation}\label{eq:cstr:weight}
            c_{\theta_i}(\estwth)
            :=
            \norm{\estwth_i}^2 - \overline{\theta_i}^2 \le 0
            , 
            \forall i\in\{0,\ldots,k\}
          \end{equation}

        \item \textbf{Convex control Input Saturation}: 
          \begin{itemize}
            \item Input bound constraint \textit{for each control input}:
            \begin{equation}\label{eq:cstr:input_bound}
              c_{\overline{\tau}_i}(\estwth)
              :=
              \tau_i - \overline{\tau_i}
              \le 
              0
              ,
              \quad
              c_{\underline{\tau}_i}(\estwth)
              :=
              \underline{\tau_i} - \tau_i
              \le 
              0
            \end{equation}
            \item Input norm constraint:
            \begin{equation}\label{eq:cstr:input_norm}
              c_{\mv{\tau}}(\estwth)
              :=
              \norm{\mv{\tau}}^2 - \overline{\tau}^2
              \le
              0
            \end{equation}
        \end{itemize}
      \end{itemize}

    \end{block}
  }

  \end{columns}

  \let\thefootnote\relax\footnote{
    \textit{Notations}: 
    $\Lambda\in\R_{>0}^{n\times n}$: filtering matrix
  }

\end{frame}

\begin{frame}{\insertsubsectionhead}{Optimization Problem}

  \textbf{Original Optimization Problem}

  \begin{itemize}
    \item Constrained optimization problem to minimize the tracking error.
    \item Inequality constraints $c_j(\estwth) \le 0$ for $j\in\mathcal{I}$.
  \end{itemize}

    \begin{equation}\label{eq:opt_problem}
      \begin{matrix}
        \min_{\estwth} J(\mv{r};\estwth)
        \\\ \\
        \text{s.t. } c_{j}(\estwth) 
        \le 
        0
        ,
        \forall j\in\mathcal{I}
      \end{matrix}
    \end{equation}

  \onslide<2->
  {
    \textbf{Define Lagrangian Function}

    \begin{equation}\label{eq:lagrangian}
        L(\mv{r},\estwth,[\lambda_j]_{j\in\mathcal{I}})
        :=
        J(\mv{r};\estwth)
        +
        \sum_{j\in\mathcal{I}} \lambda_j c_j(\estwth)
    \end{equation}
  }

  \onslide<3->
  {
    \textbf{Dual Problem}
    \begin{itemize}
      \item The dual problem is to \ctxt{airforceblue}{minimize } the Lagrangian function with respect to the \ctxt{airforceblue}{NN weights } $\estwth$, while \ctxt{awesome}{maximizing } with respect to the \ctxt{awesome}{Lagrange multipliers } $\lambda_j$.
      \item The Lagrange multipliers $\lambda_j$ are non-negative, \ie $\lambda_j \ge 0$.
    \end{itemize}

    \begin{equation}\label{eq:dual_problem}
      \min_{\estwth}\max_{[\lambda_j]_{j\in\mathcal{I}}} 
      L(\mv{r},\estwth,[\lambda_j]_{j\in\mathcal{I}})
    \end{equation}
  }

\end{frame}

\subsection{Adaptation Law Derivation}

\begin{frame}{\insertsubsectionhead}{Gradient Descent/Ascent Method}

  To solve the dual problem, 
  \begin{equation}
    \min_{\estwth}\max_{[\lambda_j]_{j\in\mathcal{I}}} 
    L(\mv{r},\estwth,[\lambda_j]_{j\in\mathcal{I}})
    ,
  \end{equation}
  the first-order gradient descent/ascent method is used to derive the adaptation law.

  \centering
  \begin{minipage}{0.6\textwidth}%

    \begin{block}{Adaptation Law}%

    \onslide<2->
    {
      \textbf{Gradient \ctxt{airforceblue}{Descent } Method for weights $\estwth$}:
      \begin{equation}\label{eq:adaptation_law:1}
        \ddt {{\estwth}}
        =
        -\alpha 
        \ppfrac{L}{{\estwth}}
        =-\alpha 
        \left(
            \ppfrac{J}{{\estwth}}
            +
            \sum_{j\in\mathcal{I}}
            \lambda_j 
            \ppfrac{c_j}{\estwth}
        \right),
      \end{equation}
    }

    \onslide<3->
    {
      \textbf{Gradient \ctxt{awesome}{Ascent } Method for Lagrange multipliers $\lambda_j, \forall j\in\mathcal I$}:
      \begin{equation}\label{eq:adaptation_law:2}
        \ddt\lambda_j
        = 
        \beta_j
        \ppfrac{L}{\lambda_j} 
        = 
        \beta_j c_j ,
      \end{equation}
    }

    \onslide<4->
    {
      For non-negativity of the Lagrange multipliers,
      \begin{equation}\label{eq:adaptation_law:3}
        \lambda_j
        \leftarrow
        \max(\lambda_j,0)
        .
      \end{equation}
    }

    \end{block}
  \end{minipage}
\end{frame}

\subsection{Stability Analysis}

\begin{frame}{\insertsubsectionhead}{Lyapunov Stability Analysis}

  \centering
  \begin{minipage}{.9\textwidth}

    \begin{block}{Theorem 1 \cite{Ryu:2025aa}}
      For the \ctxt{airforceblue}{dynamical system } described in \eqref{eq:sys}, the \ctxt{awesome}{neuro-adaptive controller } in \eqref{eq:input} with the \ctxt{airforceblue}{weight adaptation laws } in \eqref{eq:adaptation_law:1}, \eqref{eq:adaptation_law:2} and \eqref{eq:adaptation_law:3} ensure the \ctxt{airforceblue}{boundedness } of the \ctxt{awesome}{filtered error } $\mv{r}$ and the \ctxt{awesome}{weight estimate } $\estwth$, under the control input constraints satisfying Assumption 1 and 2. This holds under the \ctxt{awesome}{weight norm constraint } \eqref{eq:cstr:weight}.
    \end{block}

    \begin{exampleblock}{Assumption 1 (Convex Input Constraint)}
      The constraint functions $c_j(\estwth),\forall j\in\mathcal{I}$, are convex in the $\tau$-space and satisfy $c_j(\estwth)\le0$ and $c_j(\idealwth)\le0$.
    \end{exampleblock}

    \begin{exampleblock}{Assumption 2, Linear Independence Constraint Qualification (LICQ)}
      The selected constraints satisfy the Linear Independence Constraint Qualification (LICQ) \cite[Chap. 12 Def. 12.1]{Nocedal:2006aa}.
    \end{exampleblock}

  \end{minipage}

  \vspace{.2cm}

  Proof of Theorem 1 is \textbf{omitted } due to space limitations. The detailed proof can be found in \cite{Ryu:2025aa}.

\end{frame}

% ╔═══════════════════════════════════════════════╗
% ║ Section: Experimental Validation              ║
% ╚═══════════════════════════════════════════════╝
\section{Experimental Validation}

\subsection{Simulation Setup}

\begin{frame}{\insertsubsectionhead}{Two-Link Robotic Manipulator}

    \begin{columns}

      \column{.45\textwidth}

        \textbf{Target System}:

        \begin{equation*}
          \mm{M}\ddq + \mm{V}_m\dq + \mv{F} + \mv{G} + \mv{\tau}_d
          =
          \mv{\tau}
        \end{equation*}

        \begin{figure}
          \centering
          \includegraphics[width=.99\textwidth]{figures/RobotModel.drawio.png}
          \caption{Two-link robotic manipulator model.}
        \end{figure}

      \column{.55\textwidth}

        \textbf{Desired Trajectory}:

        \begin{equation}
          \qd
          =
          \begin{pmatrix}
              {q_d}_1\\
              {q_d}_2
          \end{pmatrix}
          = 
          \begin{pmatrix}
              +\cos(
                  \tfrac{\pi}{2}t
              ) + 1 \\
              -\cos(
                  \tfrac{\pi}{2}t
              ) - 1 
          \end{pmatrix}
          .
      \end{equation}

      \textbf{System Model Parameters}:

      \begin{table}
        \renewcommand{\arraystretch}{1.3}
        \caption{System model parameters.}
        \centering
        \begin{tabular}{c m{5em} c c c }
        \hline
        \textbf{Symbol} & \textbf{Description} & \textbf{Link 1} & \textbf{Link 2} \\
        \hline
        \hline 
        $m_p$ & Mass & 23.902 kg & 3.88 kg \\
        \hline
        $l_p$  & Length & 0.45 m & 0.45 m \\
        \hline
        ${l_c}_p$ & COM & 0.091 m & 0.048 m \\
        \hline
        $b_p$   & Viscous coef. &  2.288 Nms & 0.172 Nms \\
        \hline
        ${f_c}_p$  & Friction coef. & 7.17 Nm & 1.734 Nm \\
        \hline
        \end{tabular}
        \label{table: system parameters}
      \end{table}

    \end{columns}

\end{frame}

\begin{frame}{\insertsubsectionhead}{Controllers Setting}

  \begin{itemize}
    \item Only \ctxt{awesome}{weight norm constraint } is considered in this presentation (for \textbf{ECC}).
    \item Input saturation constraints are not considered.
  \end{itemize}

  \centering
  \begin{figure}
    \includegraphics[width=0.8\textwidth]{figures/Controller.drawio.pdf}
    \caption{Architecture of the controllers.}
  \end{figure}

\end{frame}

\begin{frame}{\insertsubsectionhead}{Controllers for Comparative Study}

    \begin{itemize}
      \item NAC-CO denotes the proposed controller based on \ctxt{airforceblue}{constrained optimization }.
      \item For NAC-L2 and NAC-eMod, the \ctxt{awesome}{stabilizing terms } $-\lambda\estwth$ and $\rho\norm{\mv{r}}\estwth$ ensures the weights boundedness, respectively.
    \end{itemize}

    \begin{table}
      \renewcommand{\arraystretch}{1.5}
      % \caption{Controller Properties.}
      \centering
      \begin{tabular}{c m{20em} c c c }
      \hline
      \textbf{Name} & \textbf{Description} & \textbf{Adaptation Law} \\
      \hline
      \hline 
        \multirow{2}{*}{NAC-L2} & NAC with $L_2$-regularization \tiny{(equal to $\sigma$-modification)} & \multirow{2}{*}
        {$
          \ddt\estwth = -\alpha \left(\ppfrac{J}{\estwth}+\ctxt{awesome}{\lambda}\estwth\right)
        $}
        \\
          & ($\ctxt{awesome}{\lambda}$ stabilizes $\estwth$ towards origin) &
        \\
      \hline
        \multirow{2}{*}{NAC-eMod} & NAC with $\epsilon$-modification & \multirow{2}{*}
        {$
          \ddt\estwth = -\alpha \left(\ppfrac{J}{\estwth}+\ctxt{awesome}{\rho}\norm{\tilde{\mv{z}}}\estwth\right)
        $}
        \\
        & ($\ctxt{awesome}{\rho}$ stabilizes proportionally to filtered error $\mv{r}$) &
        \\
      \hline
        \multirow{2}{*}{NAC-CO} & Constrained Optimization-based NAC & 
        $
          \ddt\estwth = -\alpha \left(\ppfrac{J}{\estwth}+\sum_{j\in\mathcal{I}}\lambda_j\ppfrac{c_j}{\estwth}\right)
        $ 
        \\
          (proposed) & ($\ctxt{awesome}{\beta_j}$ determines $\lambda_j$ adaptation speed) &
        $
          \ddt\lambda_j = \ctxt{awesome}{\beta_j} c_j$, and $\lambda_j \leftarrow \max(\lambda_j,0)
        $
        \\
      \hline
      \end{tabular}
      \label{table:sys:param}
    \end{table}

    \onslide<2->
    {
      \centering
      \begin{minipage}{0.75\textwidth}
        \begin{block}{Simulation Objective}
          By \ctxt{airforceblue}{varying the parameters }, \ie $\beta_j$, $\lambda$, and $\rho$, the parameter dependencies will be investigated.
        \end{block}
      \end{minipage}
    }

\end{frame}

\subsection{Simulation Results}

\begin{frame}{\insertsubsectionhead}{Box-and-Whisker Plots}

  \begin{columns}
    \column{0.6\textwidth}

      \textbf{Parameter Dependencies Investigation}:

      \small
      {
        \begin{itemize}
          \item<+-> The parameters ranged from $0.001$ to $1$ across 10 samples.
          \item<+-> NAC-L2 shows the \ctxt{awesome}{worst performance } with high variance.
          \item<+-> NAC-CO (proposed) shows the \ctxt{airforceblue}{best performance } and \ctxt{airforceblue}{lowest variance}.
          \item<+-> This result is because, 
            \begin{itemize}
              \item<+-> NAC-L2 and NAC-eMod are \ctxt{airforceblue}{biased towards the origin}. $\ddtt\estwth=-\alpha(\pptfrac{J}{\estwth}+\ctxt{awesome}{\lambda\estwth})$ (NAC-L2) or $+\ctxt{awesome}{\rho\norm{\mv{r}}\estwth}$ (NAC-eMod), proportionally to $\lambda$ and $\rho$, respectively.
              % \item<+-> $\beta_j$ in NAC-CO (proposed) only determines the \ctxt{airforceblue}{update speed } of the Lagrange multipliers $\lambda_j$, \ie $\ddtt\lambda_j=\beta_j c_j$.
              % \begin{itemize}
                \item<+-> $\ctxt{awesome}{-\lambda_j\pptfrac{c_j}{\estwth}}$ in NAC-CO (proposed) (\ie $\ddtt\estwth=-\alpha(\pptfrac{J}{\estwth}+\ctxt{awesome}{\lambda_j\pptfrac{c_j}{\estwth}})$) \ctxt{airforceblue}{disappears } when constraints are inactive (\ie $c_j<0$, and $\lambda=\beta_jc_j$ and $\lambda_j\leftarrow\max(\lambda_j,0)$). %
              % \end{itemize}
            \end{itemize}
        \end{itemize}
      }

    \column{0.4\textwidth}

      \begin{figure}
        \includegraphics[width=.89\textwidth]{figures/BoxWhisker.drawio.png}
        \caption{Box-and-whisker plots of the tracking error ISE.}
      \end{figure}

  \end{columns}

    \begin{table}[!t]
      \renewcommand{\arraystretch}{1.1}
      % \caption{Quantitative comparison of square root of tracking ISE.}
      \centering
      \begin{tabular}{c c c c }
      \hline
      & \textbf{NAC-L2}\!&\!\textbf{NAC-eMod}\!&\!\textbf{NAC-CO} (proposed) 
      \\
      \hline
      \hline 
        Maximum & \ctxt{awesome}{$11.1753 \times\!10^{-3}$} & $0.5603 \times 10^{-3}$ & \ctxt{airforceblue}{$0.3439 \times 10^{-3}$}  \\
      \hline
        Median & \ctxt{awesome}{$0.5898\times\!10^{-3}$}  & $0.5519 \times 10^{-3}$ & \ctxt{airforceblue}{$0.3240 \times 10^{-3}$}  \\
      \hline
        Minimum & \ctxt{awesome}{$0.5434\times\!10^{-3}$}  & $0.5434 \times 10^{-3}$ & \ctxt{airforceblue}{$0.3235 \times 10^{-3}$}  \\
      \hline
      \end{tabular}
      \label{table: error norm}
    \end{table}

    \let\thefootnote\relax\footnote{
      Squared root of the tracking error ISE (Integral of Squared Error), \ie $\sqrt{\int_0^T \norm{\mv{r}}^2\,\der t}$, where $T$ denotes a simulation time.
    }

\end{frame}

\begin{frame}{\insertsubsectionhead}{Weight Norms}

  \begin{columns}

    \column{0.33\textwidth}

      \begin{figure}      
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig9.eps}
        \caption{Weight norms of NAC-L2}
      \end{figure}

    \column{0.33\textwidth}

      \begin{figure}
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig10.eps}
        \caption{Weight norms of NAC-eMod}
      \end{figure}

    \column{0.33\textwidth}

      \begin{figure}
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig8.eps}
        \caption{Weight norms of NAC-CO (proposed)}
      \end{figure}

  \end{columns}

  \begin{itemize}
    \item<+-> NAC-CO (proposed) showed the \ctxt{airforceblue}{weight norms } are bounded \ctxt{awesome}{under pre-defined constraint } $\overline{\theta}=20$.
    \item<+-> NAC-L2 and NAC-eMod showed the bounded weight norms, but they \ctxt{awesome}{depended } on the parameters $\lambda$ and $\rho$, respectively.
    % \item<+-> As the parameters $\lambda$ and $\rho$ increase, the weight norms of NAC-L2 and NAC-eMod biased towards the origin, which may lead to suboptimal performance.
    \item<+-> In other words, NAC-CO tracked the desired trajectory with a \ctxt{airforceblue}{smaller weight } norm than NAC-L2 and NAC-eMod.
  \end{itemize}

\end{frame}

\begin{frame}{\insertsubsectionhead}{Tracking Performance}

  \begin{columns}

    \column{0.33\textwidth}

      \begin{figure}      
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig6.eps}
        \caption{Tracking error of NAC-L2}
      \end{figure}

    \column{0.33\textwidth}

      \begin{figure}
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig7.eps}
        \caption{Tracking error of NAC-eMod}
      \end{figure}

    \column{0.33\textwidth}

      \begin{figure}
        \includegraphics[width=0.99\textwidth]{figures/ECC/fig5.eps}
        \caption{Tracking error of NAC-CO (proposed)}
      \end{figure}

  \end{columns}

  \begin{itemize}
    \item NAC-CO (proposed) \ctxt{airforceblue}{outperformed } NAC-L2 and NAC-eMod in terms of \ctxt{awesome}{tracking performance}.
    \item As the weights are \ctxt{awesome}{biased }towards the origin proportionally to the parameters $\lambda$ and $\rho$ in NAC-L2 and NAC-eMod, respectively, the tracking performance of NAC-L2 and NAC-eMod deteriorated, as approaching toward \ctxt{airforceblue}{suboptimal points}.
  \end{itemize}

\end{frame}

\begin{frame}{Real-time Implementation Result}

  \begin{itemize}
    \item This video demonstrates:
    \begin{itemize}
      \item \ctxt{airforceblue}{Applicability }of the proposed method to real-time control (under 4 ms sampling time).
      \item \ctxt{awesome}{Convex input constraints } handling.
    \end{itemize}
  \end{itemize}

  \centering
  \includemedia[
      width=0.64\linewidth,
      height=0.36\linewidth,
      activate=onclick,
      addresource=animation.mp4,
      flashvars={
        source=animation.mp4
        &autoPlay=true
        &loop=true
      }
    ]{}{VPlayer.swf}

\end{frame}

% ╔═══════════════════════════════════════════════╗ 
% ║ Section: Conclusion and Future Work           ║
% ╚═══════════════════════════════════════════════╝
\section{Conclusion}

\subsection{Conclusion and Future Work}

\begin{frame}{Conclusion}

  \textbf{Summary of Contributions}
  \begin{itemize}
    \item Proposed a novel constrained optimization-based neuro-adaptive control (CONAC) method.
    \item Adaptation laws are derived using \ctxt{airforceblue}{constrained optimization method}.
    \item The proposed method guarantees the \ctxt{awesome}{stability }of the system and the boundedness of the NN weights.
    \item Feasibility of the proposed method is validated through numerical simulations.
  \end{itemize}

  \textbf{Future Work}
  \begin{itemize}
    \item Extend the proposed method to \ctxt{airforceblue}{state constraints}.
    \item Enhance the \ctxt{awesome}{robustness }and \ctxt{awesome}{flexibility }of the proposed method for \ctxt{airforceblue}{various systems}.
  \end{itemize}

\end{frame}

\section{}
\begin{frame}{}
    \centering \Large
    \emph{Thank you for your attention!}
\end{frame}

\begin{frame}[allowframebreaks]{References}

  \bibliography{localRefs}
  \bibliographystyle{IEEEtran}

\end{frame}

\end{document}